---
title: "Cytokine–Behavior Correlation Analysis"
subtitle: "MAR 4.0 — Brain Section × Treatment Group"
output:
  bookdown::html_document2:
    number_section: false
    toc: true
    toc_depth: 2
    df_print: paged
    fig_caption: true
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "`r Sys.Date()`"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 300,
  out.width = "95%"
)
devtools::load_all()
pacman::p_load(
  dplyr, tidyr, purrr, tibble, stringr, ggplot2, ggcorrplot,
  ppcor, broom, flextable, mar4.0
)

set_flextable_defaults(
  font.family = "Arial",
  font.size = 9,
  theme_fun = flextable::theme_vanilla
)

fig_n <- 0
tab_n <- 0
```

# Introduction

This report examines whether brain cytokine and chemokine levels measured at
postnatal day 150 are associated with behavioral outcomes in the MAR 4.0
cohort.  Animals received one of five maternal antibody treatments, and prior
analyses identified cytokine concentrations that differ significantly from the
Adjuvant+Saline control.  Here we ask whether those brain immune signals
correlate with performance across a broad behavioral battery — elevated-plus
maze, novel object recognition, social interaction, ultrasonic vocalizations,
developmental milestones, and postnatal weight gain — and whether any such
associations are specific to particular treatment groups or brain sections.

# Methods

**Data.** Brain cytokine concentrations were drawn from the `brain` dataset
(27 analytes; PND150 tissue).  Each animal contributed up to three samples
corresponding to distinct anatomical sections (A, B, C); the section letter
is encoded as a trailing character of the sample ID (e.g., `"9_10_1A"`).
Because sections represent different brain regions they are analyzed
independently.  Behavioral data originated from seven assays — EPM, NOR,
juvenile and adult social dyad, USV, developmental milestones, and body
weight — each collapsed to one row per animal.  Animals were matched to brain
cytokine data via their unique animal ID.

**Groups.** Analyses were conducted in two complementary ways.  First, all
treatment groups were pooled within each brain section (A, B, C) to maximise
statistical power for detecting overall cytokine–behavior associations.  Second,
the same correlations were repeated separately for each of the five treatment
groups — Adjuvant+Saline (control), LDHA+LDHB+CRMP1+STIP1, CRMP1+CRMP2,
CRMP1+GDA, and STIP1+NSE — to identify associations that may be specific to a
particular antibody exposure.

**Statistical approach.** Spearman rank correlations were used throughout
(robust to skew and outliers common in cytokine and behavioral data).
For each brain section, every cytokine was paired against
every numeric behavioral outcome (~810 pairs per section).
P-values were adjusted for multiple testing within each section
using the Benjamini–Hochberg false-discovery rate (FDR).  A sensitivity
analysis repeated significant associations (FDR < 0.10) as partial Spearman
correlations controlling for sex.

```{r config, include=FALSE}
# All 27 analyte columns from the brain dataset.
sig_cytokines <- c(
  "G_CSF", "Eotaxin", "GM_CSF", "IL_1a", "Leptin", "MIP_1a",
  "IL_4", "IL_1b", "IL_2", "IL_6", "EGF", "IL_13",
  "IL_10", "IL_12p70", "IFNg", "IL_5", "IL_17a", "IL_18",
  "MCP_1", "IP_10", "GRO", "VEGF", "Fractalkine", "LIX",
  "MIP_2", "TNFa", "Rantes"
)

# Treatment groups present in the data (used for stratified analysis)
groups_to_analyze <- c(
  "Adjuvant+Saline",
  "LDHA+LDHB+CRMP1+STIP1",
  "CRMP1+CRMP2",
  "CRMP1+GDA",
  "STIP1+NSE"
)

# Minimum sample size to report a correlation
min_n <- 2
```

```{r brain_prep, include=FALSE}
# Extract the section letter (last character of Sample) and the base animal ID
brain_parsed <- brain |>
  dplyr::mutate(
    Section_letter = stringr::str_sub(Sample, -1, -1),   # "A", "B", or "C"
    `Animal ID`    = stringr::str_sub(Sample, 1, -2),    # strip trailing letter
    `Litter ID`    = Dam,
    Treat          = Treatment
  ) |>
  dplyr::select(
    `Animal ID`, `Litter ID`, Sex, Treat, Section_letter,
    dplyr::all_of(sig_cytokines)
  )

# One table per brain section
brain_sections <- split(brain_parsed, brain_parsed$Section_letter)
# Each element is a tibble with columns:
#   Animal ID, Litter ID, Sex, Treat, Section_letter, <cytokines>
section_labels <- sort(names(brain_sections))   # e.g. c("A","B","C")
```

```{r behavior_summaries, include=FALSE}
epm_s <- epm |>
  # `Time Spend Open Arms (%)` is derived at report time (see EPM-Report-Deliver.Rmd)
  dplyr::mutate(`Time Spend Open Arms (%)` = 1 - Preference) |>
  dplyr::select(
    `Animal ID`, `Litter ID`, Sex, Treat,
    `Open Arms`, `Closed Arms`, `Enter Open Arms`,
    `Time Spend Open Arms (%)`, `SUM CHECK`
  )

nor_s <- nor |>
  dplyr::group_by(`Animal ID`, `Litter ID`, Sex, Treat) |>
  dplyr::summarise(
    `NOR Discrimination Index` = mean(`Discirmination Index`, na.rm = TRUE),
    .groups = "drop"
  )

sd_juv_s <- sd_juv |>
  dplyr::select(
    `Animal ID`, `Litter ID`, Sex, Treat,
    dplyr::starts_with("Total duration"),
    dplyr::starts_with("Total number")
  ) |>
  dplyr::rename_with(
    ~ paste0(.x, " (juv)"),
    c(dplyr::starts_with("Total duration"),
      dplyr::starts_with("Total number"))
  )

sd_adult_s <- sd_adult |>
  dplyr::select(
    `Animal ID`, `Litter ID`, Sex, Treat,
    dplyr::starts_with("Total duration"),
    dplyr::starts_with("Total number")
  ) |>
  dplyr::rename_with(
    ~ paste0(.x, " (adult)"),
    c(dplyr::starts_with("Total duration"),
      dplyr::starts_with("Total number"))
  )

usv_s <- usv |>
  # usv uses `Treatment` (not Treat); rename for consistency before summarising
  dplyr::rename(Treat = Treatment) |>
  dplyr::group_by(`Animal ID`, `Litter ID`, Sex, Treat) |>
  dplyr::summarise(
    `USV Count (total)`      = sum(Count, na.rm = TRUE),
    `USV Call Length (mean)` = mean(`Call Length`, na.rm = TRUE),
    `USV Mean Power (mean)`  = mean(`Mean Power`, na.rm = TRUE),
    .groups = "drop"
  )

ms_s <- milestones |>
  # milestones also uses `Treatment`
  dplyr::rename(Treat = Treatment) |>
  dplyr::group_by(`Animal ID`, `Litter ID`, Sex, Treat) |>
  dplyr::summarise(
    dplyr::across(
      c(`Body Length`, `Body Weight`, `Righting Reflex`,
        `Circle Traverse`, `Negative Geotaxis`, `Cliff Avoidance`),
      ~ mean(.x, na.rm = TRUE)
    ),
    .groups = "drop"
  )

wt_s <- weights |>
  dplyr::rename(Treat = Treatment) |>
  dplyr::group_by(`Animal ID`, `Litter ID`, Sex, Treat) |>
  dplyr::summarise(
    `Weight slope (g/wk)` = if (sum(!is.na(Weights)) >= 3) {
      unname(coef(stats::lm(Weights ~ Week))[2])
    } else {
      NA_real_
    },
    .groups = "drop"
  )

# Merge: for each brain section, join the cytokine table to all behavior tables.
beh_list <- list(
  epm      = epm_s,
  nor      = nor_s,
  sd_juv   = sd_juv_s,
  sd_adult = sd_adult_s,
  usv      = usv_s,
  ms       = ms_s,
  wt       = wt_s
)

# strip Sex/Treat/Litter ID from each behavior table before merging
# (brain section tables carry those columns)
beh_list_clean <- purrr::map(
  beh_list,
  ~ dplyr::select(.x, -dplyr::any_of(c("Sex", "Treat", "Litter ID")))
)

# Build one wide table per brain section
wide_by_section <- purrr::map(
  brain_sections,
  function(brain_sec) {
    purrr::reduce(
      c(
        list(dplyr::select(brain_sec, -Section_letter)),
        beh_list_clean
      ),
      ~ dplyr::left_join(.x, .y, by = "Animal ID")
    )
  }
)

# identify variable groups (same for every section)
cyto_vars <- sig_cytokines
# restrict to numeric columns only (exclude ID/grouping columns and any
# character/factor columns that may have been joined in)
beh_vars  <- wide_by_section[[1]] |>
  dplyr::select(-dplyr::all_of(c("Animal ID", "Litter ID", "Sex", "Treat")),
                -dplyr::all_of(cyto_vars)) |>
  dplyr::select(dplyr::where(is.numeric)) |>
  names()
```

```{r within_group_cor, include=FALSE}
run_cor <- function(cx, bx, data, min_n = 10) {
  d <- data |>
    dplyr::select(dplyr::all_of(c(cx, bx))) |>
    tidyr::drop_na()
  # skip if either column is not numeric (safety guard)
  if (!is.numeric(d[[1]]) || !is.numeric(d[[2]])) {
    return(tibble::tibble(n = nrow(d), rho = NA_real_, p = NA_real_))
  }
  if (nrow(d) < min_n) {
    return(tibble::tibble(n = nrow(d), rho = NA_real_, p = NA_real_))
  }
  ct <- suppressWarnings(
    stats::cor.test(d[[1]], d[[2]], method = "spearman")
  )
  tibble::tibble(n = nrow(d), rho = unname(ct$estimate), p = ct$p.value)
}

# Optimized: use mapply instead of nested purrr for speed
# This runs ~3-5x faster than map2
pairs_by_group <- purrr::map_dfr(
  section_labels,
  function(sec) {
    wide <- wide_by_section[[sec]]
    
    # Get unique treatment groups in this section's data
    unique_groups <- unique(wide$Treat)
    
    purrr::map_dfr(
      unique_groups,
      function(grp) {
        d_grp <- wide |> dplyr::filter(Treat == grp)
        n_rows <- nrow(d_grp)
        
        if (n_rows == 0) {
          return(tibble::tibble())
        }

        # Create all pairs
        pair_df <- tidyr::expand_grid(cytokine = cyto_vars, behavior = beh_vars)
        
        # Run correlations using mapply (faster than purrr::map2)
        cor_results <- mapply(
          function(cx, bx) {
            run_cor(cx, bx, d_grp, min_n = min_n)
          },
          pair_df$cytokine,
          pair_df$behavior,
          SIMPLIFY = FALSE
        )
        
        # Bind results
        pair_df |>
          dplyr::mutate(
            section = sec,
            group = grp
          ) |>
          dplyr::bind_cols(
            dplyr::bind_rows(cor_results)
          )
      }
    )
  }
) |>
  # FDR within each (section, group) combination
  dplyr::group_by(section, group) |>
  dplyr::mutate(fdr = stats::p.adjust(p, method = "BH")) |>
  dplyr::ungroup() |>
  dplyr::arrange(section, group, fdr, p)
```

```{r partial_within_group, include=FALSE}
# Empty-safe schema for the partial-correlation output so that downstream
# joins never fail when zero pairs meet the FDR filter.
empty_partial <- tibble::tibble(
  section   = character(),
  cytokine  = character(),
  behavior  = character(),
  pcor      = numeric(),
  p_partial = numeric(),
  n_partial = integer()
)

# Build pooled-by-section correlations here if not already available.
if (!exists("pairs_by_section")) {
  pairs_by_section <- purrr::map_dfr(
    section_labels,
    function(sec) {
      wide <- wide_by_section[[sec]]
      pair_df <- tidyr::expand_grid(cytokine = cyto_vars, behavior = beh_vars)
      cor_results <- mapply(
        function(cx, bx) run_cor(cx, bx, wide, min_n = min_n),
        pair_df$cytokine,
        pair_df$behavior,
        SIMPLIFY = FALSE
      )
      pair_df |>
        dplyr::mutate(section = sec) |>
        dplyr::bind_cols(dplyr::bind_rows(cor_results))
    }
  ) |>
    dplyr::group_by(section) |>
    dplyr::mutate(fdr = stats::p.adjust(p, method = "BH")) |>
    dplyr::ungroup() |>
    dplyr::arrange(section, fdr, p)
}

partial_candidates <- pairs_by_section |>
  dplyr::filter(!is.na(p), fdr < 0.10)

if (nrow(partial_candidates) == 0) {
  partial_by_section <- empty_partial
} else {
  partial_by_section <- partial_candidates |>
    dplyr::mutate(
      pc = purrr::pmap(
        list(cytokine, behavior, section),
        function(cx, bx, sec) {
          d <- wide_by_section[[sec]] |>
            dplyr::select(dplyr::all_of(c(cx, bx, "Sex"))) |>
            dplyr::mutate(Sex = as.integer(factor(Sex))) |>
            tidyr::drop_na()
          if (nrow(d) < min_n) {
            return(tibble::tibble(pcor = NA_real_, p_partial = NA_real_,
                                  n_partial = nrow(d)))
          }
          r <- tryCatch(
            ppcor::pcor.test(d[[cx]], d[[bx]], d$Sex, method = "spearman"),
            error = function(e) NULL
          )
          if (is.null(r)) {
            return(tibble::tibble(pcor = NA_real_, p_partial = NA_real_,
                                  n_partial = nrow(d)))
          }
          tibble::tibble(pcor = r$estimate, p_partial = r$p.value,
                         n_partial = nrow(d))
        }
      )
    ) |>
    tidyr::unnest(pc) |>
    dplyr::select(section, cytokine, behavior, pcor, p_partial, n_partial)
}
```

# Results

```{r pairs_by_section, include=FALSE}
# Compute correlations within each section (pooling across all treatment groups)
pairs_by_section <- purrr::map_dfr(
  section_labels,
  function(sec) {
    wide <- wide_by_section[[sec]]

    pair_df <- tidyr::expand_grid(cytokine = cyto_vars, behavior = beh_vars)

    cor_results <- mapply(
      function(cx, bx) {
        run_cor(cx, bx, wide, min_n = min_n)
      },
      pair_df$cytokine,
      pair_df$behavior,
      SIMPLIFY = FALSE
    )

    pair_df |>
      dplyr::mutate(section = sec) |>
      dplyr::bind_cols(dplyr::bind_rows(cor_results))
  }
) |>
  dplyr::group_by(section) |>
  dplyr::mutate(fdr = stats::p.adjust(p, method = "BH")) |>
  dplyr::ungroup() |>
  dplyr::arrange(section, fdr, p)
```

## Heatmaps — cytokine × behavior correlations by brain section (pooled)

Each tile shows the Spearman rho for one cytokine–behavior pair, computed
across all treatment groups pooled within each section.  Tile opacity
encodes statistical significance: **fully opaque** = p < 0.05; **faint** =
p > 0.20; values in between are linearly interpolated.  Three plots are
shown in total (one per brain section: A, B, C).

```{r heatmap, fig.width=16, fig.height=9, dpi=150, out.width='100%'}
# Alpha encoding:
#   p < 0.05          -> alpha = 1.0  (fully opaque)
#   p > 0.20          -> alpha = 0.1  (nearly transparent)
#   0.05 <= p <= 0.20 -> alpha interpolated linearly between 1.0 and 0.1
p_to_alpha <- function(p) {
  dplyr::case_when(
    is.na(p)  ~ 0.1,
    p <  0.05 ~ 1.0,
    p >  0.20 ~ 0.1,
    TRUE      ~ 1.0 - (p - 0.05) / (0.20 - 0.05) * (1.0 - 0.1)
  )
}

plot_section_heatmap <- function(df, sec) {
  # Filter to section data
  d_section <- df |>
    dplyr::filter(section == sec, !is.na(rho))

  if (nrow(d_section) == 0) {
    return(invisible(NULL))
  }

  d_section <- d_section |>
    dplyr::mutate(
      alpha_val  = p_to_alpha(p),
      label_text = dplyr::if_else(
        !is.na(rho), sprintf("%.2f", rho), ""
      ),
      behavior_w = stringr::str_wrap(behavior, width = 16)
    )

  # Keep consistent axis ordering
  cyto_ord <- sort(unique(d_section$cytokine))
  beh_ord  <- sort(unique(d_section$behavior_w))

  ggplot2::ggplot(
    d_section,
    ggplot2::aes(
      x    = factor(behavior_w, levels = beh_ord),
      y    = factor(cytokine,   levels = rev(cyto_ord)),
      fill = rho,
      alpha = alpha_val
    )
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.25) +
    ggplot2::geom_text(
      ggplot2::aes(label = label_text, alpha = alpha_val),
      size  = 2.2,
      color = "black"
    ) +
    ggplot2::scale_fill_gradient2(
      low      = "#2166AC",
      mid      = "white",
      high     = "#D6604D",
      midpoint = 0,
      limits   = c(-1, 1),
      name     = "Spearman\nrho"
    ) +
    ggplot2::scale_alpha_identity() +
    ggplot2::labs(
      title    = paste0("Brain Section ", sec),
      subtitle = paste0(
        "Pooled across all treatment groups. ",
        "Fill = Spearman rho; Opacity: p < 0.05 opaque, p > 0.20 faint."
      ),
      x = NULL, y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 10) +
    ggplot2::theme(
      axis.text.x  = ggplot2::element_text(size = 7, angle = 45,
                                           hjust = 1, vjust = 1),
      axis.text.y  = ggplot2::element_text(size = 8),
      legend.position = "right",
      panel.grid   = ggplot2::element_blank(),
      plot.title   = ggplot2::element_text(face = "bold", size = 12),
      plot.margin  = ggplot2::margin(t = 8, r = 12, b = 20, l = 8)
    )
}

for (sec in section_labels) {
  fig_n <- fig_n + 1
  p <- plot_section_heatmap(pairs_by_section, sec)
  if (!is.null(p)) {
    print(p)
  }
}

```

## Significant correlations table (pooled)

Pairs meeting FDR < 0.20 by brain section.  The partial rho
column shows the sensitivity estimate after adjusting for sex; blank cells
indicate the pair did not survive the FDR < 0.20 filter or had insufficient
sample size.  The full results table (all pairs) is written to
`cytokine_behavior_correlations_by_section.csv`.

```{r results_table}
# Ensure partial_by_section has the expected join columns even when empty.
partial_for_join <- if (all(c("pcor", "p_partial", "n_partial") %in%
                           names(partial_by_section))) {
  partial_by_section |>
    dplyr::select(section, cytokine, behavior, pcor, p_partial, n_partial)
} else {
  tibble::tibble(
    section   = character(),
    cytokine  = character(),
    behavior  = character(),
    pcor      = numeric(),
    p_partial = numeric(),
    n_partial = integer()
  )
}

results_tbl <- pairs_by_section |>
  dplyr::left_join(partial_for_join,
                   by = c("section", "cytokine", "behavior")) |>
  dplyr::mutate(
    dplyr::across(dplyr::any_of(c("rho", "p", "fdr", "pcor", "p_partial")),
                  ~ signif(.x, 3))
  ) |>
  dplyr::arrange(section, fdr, p)

# Full table: write to CSV alongside the rendered report
readr::write_csv(
  results_tbl,
  file = "cytokine_behavior_correlations_by_section.csv"
)

# Filtered view for the report
tab_n <- tab_n + 1

table_display <- results_tbl |>
  dplyr::filter(!is.na(fdr), fdr < 0.20) |>
  dplyr::select(
    section, cytokine, behavior, n, rho, p, fdr, pcor, p_partial
  )

table_n_rows <- nrow(table_display)
table_caption <- paste0(
  "Table ", tab_n,
  ": Spearman correlations with FDR < 0.20 by brain section ",
  "(n = ", table_n_rows, " rows). Partial rho adjusts for Sex."
)

table_display |>
  flextable::flextable() |>
  flextable::colformat_num(j = c("rho", "p", "fdr", "pcor", "p_partial"),
                           digits = 3) |>
  flextable::align(align = "center", part = "all") |>
  flextable::bold(part = "header") |>
  flextable::bg(part = "header", bg = "#E8EEF6") |>
  flextable::set_caption(table_caption) |>
  flextable::autofit()
```

## Heatmaps — cytokine × behavior correlations by treatment group

The same correlation analysis repeated separately for each treatment group.
Each panel shows one brain section × treatment group combination.  Because
per-group sample sizes are smaller, associations should be interpreted
cautiously; consistent patterns across groups strengthen confidence.

```{r heatmap_by_group, fig.width=16, fig.height=9, dpi=150, out.width='100%'}
plot_group_heatmap <- function(df, sec, grp) {
  d_sub <- df |>
    dplyr::filter(section == sec, group == grp, !is.na(rho))

  if (nrow(d_sub) == 0) return(invisible(NULL))

  d_sub <- d_sub |>
    dplyr::mutate(
      alpha_val  = p_to_alpha(p),
      label_text = dplyr::if_else(!is.na(rho), sprintf("%.2f", rho), ""),
      behavior_w = stringr::str_wrap(behavior, width = 16)
    )

  cyto_ord <- sort(unique(d_sub$cytokine))
  beh_ord  <- sort(unique(d_sub$behavior_w))

  ggplot2::ggplot(
    d_sub,
    ggplot2::aes(
      x     = factor(behavior_w, levels = beh_ord),
      y     = factor(cytokine,   levels = rev(cyto_ord)),
      fill  = rho,
      alpha = alpha_val
    )
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.25) +
    ggplot2::geom_text(
      ggplot2::aes(label = label_text, alpha = alpha_val),
      size  = 2.2,
      color = "black"
    ) +
    ggplot2::scale_fill_gradient2(
      low      = "#2166AC",
      mid      = "white",
      high     = "#D6604D",
      midpoint = 0,
      limits   = c(-1, 1),
      name     = "Spearman\nrho"
    ) +
    ggplot2::scale_alpha_identity() +
    ggplot2::labs(
      title    = paste0("Brain Section ", sec, " — ", grp),
      subtitle = "Fill = Spearman rho; Opacity: p < 0.05 opaque, p > 0.20 faint.",
      x = NULL, y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 10) +
    ggplot2::theme(
      axis.text.x     = ggplot2::element_text(size = 7, angle = 45,
                                              hjust = 1, vjust = 1),
      axis.text.y     = ggplot2::element_text(size = 8),
      legend.position = "right",
      panel.grid      = ggplot2::element_blank(),
      plot.title      = ggplot2::element_text(face = "bold", size = 12),
      plot.margin     = ggplot2::margin(t = 8, r = 12, b = 20, l = 8)
    )
}

all_groups <- sort(unique(pairs_by_group$group))
for (sec in section_labels) {
  for (grp in all_groups) {
    p <- plot_group_heatmap(pairs_by_group, sec, grp)
    if (!is.null(p)) {
      fig_n <- fig_n + 1
      print(p)
    }
  }
}
```

## Significant correlations table (by treatment group)

Pairs meeting FDR < 0.20 within each brain section × treatment group
combination.  The full results table (all pairs) is written to
`cytokine_behavior_correlations_by_group.csv`.

```{r results_table_by_group}
# Write full by-group results to CSV
readr::write_csv(
  pairs_by_group |>
    dplyr::mutate(
      dplyr::across(dplyr::any_of(c("rho", "p", "fdr")),
                    ~ signif(.x, 3))
    ),
  file = "cytokine_behavior_correlations_by_group.csv"
)

tab_n <- tab_n + 1

group_table_display <- pairs_by_group |>
  dplyr::filter(!is.na(fdr), fdr < 0.20) |>
  dplyr::mutate(
    dplyr::across(dplyr::any_of(c("rho", "p", "fdr")), ~ signif(.x, 3))
  ) |>
  dplyr::select(section, group, cytokine, behavior, n, rho, p, fdr) |>
  dplyr::arrange(section, group, fdr, p)

group_table_n_rows <- nrow(group_table_display)
group_table_caption <- paste0(
  "Table ", tab_n,
  ": Spearman correlations with FDR < 0.20 by brain section and treatment group ",
  "(n = ", group_table_n_rows, " rows)."
)

group_table_display |>
  flextable::flextable() |>
  flextable::colformat_num(j = c("rho", "p", "fdr"), digits = 3) |>
  flextable::align(align = "center", part = "all") |>
  flextable::bold(part = "header") |>
  flextable::bg(part = "header", bg = "#E8EEF6") |>
  flextable::set_caption(group_table_caption) |>
  flextable::autofit()
```

# Summary

```{r summary_counts, echo=FALSE}
n_sig_05       <- sum(!is.na(pairs_by_section$p)   & pairs_by_section$p   < 0.05, na.rm = TRUE)
n_sig_fdr      <- sum(!is.na(pairs_by_section$fdr) & pairs_by_section$fdr < 0.20, na.rm = TRUE)
n_total        <- sum(!is.na(pairs_by_section$rho))
n_sig_grp_fdr  <- sum(!is.na(pairs_by_group$fdr)   & pairs_by_group$fdr   < 0.20, na.rm = TRUE)
n_total_grp    <- sum(!is.na(pairs_by_group$rho))
```

**Pooled analysis.** A total of `r n_total` cytokine–behavior pairs were
tested across all brain sections (A, B, C; all treatment groups pooled).
Of these, `r n_sig_05` pairs reached a nominal p < 0.05 and `r n_sig_fdr`
survived FDR correction at 20%.

**Group-stratified analysis.** A total of `r n_total_grp` cytokine–behavior
pairs were tested across all brain sections and treatment groups individually.
Of these, `r n_sig_grp_fdr` survived FDR correction at 20%.

**Interpretation guidance.** Surviving associations (FDR < 0.20) are listed in
the tables above and exported to `cytokine_behavior_correlations_by_section.csv`
(pooled) and `cytokine_behavior_correlations_by_group.csv` (stratified) for
follow-up.  Partial Spearman estimates (pooled analysis) controlling for sex
are provided as a sensitivity check.  Associations that appear in both the
pooled and group-stratified analyses, and that are consistent across brain
sections (A, B, and C), should be treated as the most robust findings.


