---
title: "MAR 4.0 - Elevated Plus Maze"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2025-05-08"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE , warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, mar4.0)
fig_n = 0
tab_n = 0

data <- epm |>
  rename(Treatment = Treat) |>
  mutate(
    `Time Spend Open Arms (%)` = 1 - Preference,
    `Time Spend Open Arms (%) - ADJ` = (`Time Spend Open Arms (%)` * (n() - 1) + 0.0001) / n(),
    `Log2 Open Arms` = log2(`Open Arms` + min(`Open Arms`[`Open Arms` > 0]) / 2)
  )
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glmmTMB::glmmTMB(as.formula(formula_str), ziformula = ~ 1, family = glmmTMB::tweedie(link = "log"), data = data))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}

# Function to summarize and format model results
format_summary_lmer <- function(model_summary, column_name) {
  formatted <- apply(model_summary$coefficients[, c(1,2,4)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

format_summary_glmmTMB <- function(model_summary, column_name) {
  formatted <- apply(model_summary$coefficients$cond[, c(1,2,4)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(..., summary_full) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  for (i in 2:length(summaries)) {
    merged_data <- merge(merged_data, summaries[[i]], by = "row.names", all = TRUE)
    rownames(merged_data) <- merged_data$Row.names
    merged_data <- merged_data[ , !(colnames(merged_data) %in% "Row.names")]
  }
  ordered_data <- merged_data[order(match(rownames(merged_data), rownames(summary_full))), ]
  ordered_data[is.na(ordered_data)] = ""
  return(ordered_data)
}

# Function to perform ANOVA
perform_anova <- function(model, type = "III") {
  return(Anova(model, type = type)[, 3])
}
```




```{r}
tx_tbl <- data |>
  dplyr::select(
    Treatment, 
    `Enter Open Arms`, 
    `SUM CHECK`, 
    `Open Arms`, 
    `Closed Arms`,
    `Time Spend Open Arms (%)`
  ) |>
  gtsummary::tbl_summary(
    by = Treatment,
    type = list(
      `Enter Open Arms` ~ "continuous2",
      `SUM CHECK` ~ "continuous2",
      `Open Arms` ~ "continuous2",
      `Closed Arms` ~ "continuous2",
      `Time Spend Open Arms (%)` ~ "continuous2"
    ),
    statistic = list(
      gtsummary::all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} [{p25}, {p75}]",
        "{min}, {max}"
      )
    )
  ) |>
  gtsummary::add_overall()


tx_sex_tbl <- data |>
  dplyr::select(
    Treatment, Sex,
    `Enter Open Arms`, 
    `SUM CHECK`, 
    `Open Arms`, 
    `Closed Arms`,
    `Time Spend Open Arms (%)`
  ) |>
  gtsummary::tbl_strata(
    strata = Sex,
    .tbl_fun = 
      ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = list(
            `Enter Open Arms` ~ "continuous2",
            `SUM CHECK` ~ "continuous2",
            `Open Arms` ~ "continuous2",
            `Closed Arms` ~ "continuous2",
            `Time Spend Open Arms (%)` ~ "continuous2"
          ),
          statistic = list(
            gtsummary::all_continuous2() ~ c(
              "{mean} ({sd})",
              "{median} [{p25}, {p75}]",
              "{min}, {max}"
            )
          )
        ) |>
        gtsummary::add_overall() |>
        gtsummary::modify_header(
          gtsummary::all_stat_cols() ~ "**{level}**"
        ),
    .combine_with = 'tbl_stack'
  )
```
    
```{r warning=FALSE}
# Comment Code to examine the distribution and model fitting
# var(data$`Enter Open Arms`) 
# hist(data$`Enter Open Arms`)
# hist(data$`SUM CHECK`)
# hist(data$`Open Arms`,breaks = 20)
# hist(data$`Log2 Open Arms`,breaks = 20)
# hist(data$`Closed Arms`,breaks = 20)
# hist(data$`Time Spend Open Arms (%)`,breaks = 20) 

vars = c("Enter Open Arms",  # overdispersion where mean = 6 < var = 9.6
         "SUM CHECK",       # overdispersion where mean = 13 < var = 30
         "Log2 Open Arms",       # Use log 2 transform
         "Closed Arms",     # OLS
         "Time Spend Open Arms (%) - ADJ")    # Beta (0 - 1)

usv_anova_tb <- data.frame(Variables = c("Interaction - Treatment",
                                         "Interaction - Sex",
                                         "Interaction - Treatment:Sex", 
                                         "No Interaction - Treatment",
                                         "No Interaction - Sex"))
merged_summary_list <- list()
post_hoc_h1_list <- list()
post_hoc_h2_list <- list()

#1 Enter Open Arms
formula1 <-  paste0("`", vars[1], "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex` + (1 | `Litter ID`)")
formula2 <-  paste0("`", vars[1], "` ~ `Treatment` + `Sex` + (1 | `Litter ID`)")
model1 <- glmmTMB::glmmTMB(as.formula(formula1), family = glmmTMB::nbinom2(link = "log"), data = data)
model2 <- glmmTMB::glmmTMB(as.formula(formula2), family = glmmTMB::nbinom2(link = "log"), data = data)
summary1 <- format_summary_glmmTMB(summary(model1), "Three Interac")
summary2 <- format_summary_glmmTMB(summary(model2), "Two - Full")
merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
merged_summary_list[[1]] <- merged_summary_df
usv_anova_tb[1:3, vars[1]] <- perform_anova(model1)[2:4]
usv_anova_tb[4:5, vars[1]] <- perform_anova(model2)[2:3]
p_list_3 <- emmeans::contrast(emmeans::emmeans(model1, ~ Treatment), method = "pairwise", adjust = "none") |>
  as.data.frame() |> 
  mutate(PND = 4) |>
  select(`contrast`, `p.value`, `PND`)
p_list_2 <- emmeans::contrast(emmeans::emmeans(model2, ~ Treatment), method = "pairwise", adjust = "none") |>
  as.data.frame() |> 
  mutate(PND = 4) |>
  select(`contrast`, `p.value`, `PND`)
post_hoc_h1_list[[1]] <- run_posthoc_test_num_h1(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[1], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)
post_hoc_h2_list[[1]] <- run_posthoc_test_num_h2(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[1], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)
  
  
#2 SUM CHECK
formula1 <-  paste0("`", vars[2], "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex` + (1 | `Litter ID`)")
formula2 <-  paste0("`", vars[2], "` ~ `Treatment` + `Sex` + (1 | `Litter ID`)")
model1 <- glmmTMB::glmmTMB(as.formula(formula1), family = glmmTMB::nbinom2(link = "log"), data = data)
model2 <- glmmTMB::glmmTMB(as.formula(formula2), family = glmmTMB::nbinom2(link = "log"), data = data)
summary1 <- format_summary_glmmTMB(summary(model1), "Three Interac")
summary2 <- format_summary_glmmTMB(summary(model2), "Two - Full")
merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
merged_summary_list[[2]] <- merged_summary_df
usv_anova_tb[1:3, vars[2]] <- perform_anova(model1)[2:4]
usv_anova_tb[4:5, vars[2]] <- perform_anova(model2)[2:3]
p_list_3 <- emmeans::contrast(emmeans::emmeans(model1, ~ Treatment), method = "pairwise", adjust = "none") |>
  as.data.frame() |> 
  mutate(PND = 4) |>
  select(`contrast`, `p.value`, `PND`)
p_list_2 <- emmeans::contrast(emmeans::emmeans(model2, ~ Treatment), method = "pairwise", adjust = "none") |>
  as.data.frame() |> 
  mutate(PND = 4) |>
  select(`contrast`, `p.value`, `PND`)
post_hoc_h1_list[[2]] <- run_posthoc_test_num_h1(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[2], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)
post_hoc_h2_list[[2]] <- run_posthoc_test_num_h2(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[2], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)
  
#3 Log2 Open Arms
formula1 <-  paste0("`", vars[3], "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex` + (1 | `Litter ID`)")
formula2 <-  paste0("`", vars[3], "` ~ `Treatment` + `Sex` + (1 | `Litter ID`)")
model1 <- lmerTest::lmer(as.formula(formula1), data = data)
model2 <- lmerTest::lmer(as.formula(formula2), data = data)
summary1 <- format_summary_lmer(summary(model1), "Three Interac")
summary2 <- format_summary_lmer(summary(model2), "Two - Full")
merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
merged_summary_list[[3]] <- merged_summary_df
usv_anova_tb[1:3, vars[[3]]] <- perform_anova(model1)[2:4]
usv_anova_tb[4:5, vars[[3]]] <- perform_anova(model2)[2:3]
p_list_3 <- emmeans::contrast(emmeans::emmeans(model1, ~ Treatment), method = "pairwise", adjust = "none") |>
  as.data.frame() |> 
  mutate(PND = 4) |>
  select(`contrast`, `p.value`, `PND`)
p_list_2 <- emmeans::contrast(emmeans::emmeans(model2, ~ Treatment), method = "pairwise", adjust = "none") |>
  as.data.frame() |> 
  mutate(PND = 4) |>
  select(`contrast`, `p.value`, `PND`)
post_hoc_h1_list[[3]] <- run_posthoc_test_num_h1(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[3], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)
post_hoc_h2_list[[3]] <- run_posthoc_test_num_h2(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[3], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)

#4 Closed Arms
formula1 <-  paste0("`", vars[4], "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex` + (1 | `Litter ID`)")
formula2 <-  paste0("`", vars[4], "` ~ `Treatment` + `Sex` + (1 | `Litter ID`)")
model1 <- lmerTest::lmer(as.formula(formula1), data = data)
model2 <- lmerTest::lmer(as.formula(formula2), data = data)
summary1 <- format_summary_lmer(summary(model1), "Three Interac")
summary2 <- format_summary_lmer(summary(model2), "Two - Full")
merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
merged_summary_list[[4]] <- merged_summary_df
usv_anova_tb[1:3, vars[[4]]] <- perform_anova(model1)[2:4]
usv_anova_tb[4:5, vars[[4]]] <- perform_anova(model2)[2:3]
post_hoc_h1_list[[4]] <- run_posthoc_test_num_h1(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[4], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)
post_hoc_h2_list[[4]] <- run_posthoc_test_num_h2(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[4], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)

#5 Time Spend Open Arms (%)
formula1 <-  paste0("`", vars[5], "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex` + (1 | `Litter ID`)")
formula2 <-  paste0("`", vars[5], "` ~ `Treatment` + `Sex` + (1 | `Litter ID`)")
model1 <- glmmTMB::glmmTMB(as.formula(formula1), family = glmmTMB::beta_family(link = "logit"), data = data)
model2 <- glmmTMB::glmmTMB(as.formula(formula2), family = glmmTMB::beta_family(link = "logit"), data = data)
summary1 <- format_summary_glmmTMB(summary(model1), "Three Interac")
summary2 <- format_summary_glmmTMB(summary(model2), "Two - Full")
merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
merged_summary_list[[5]] <- merged_summary_df
usv_anova_tb[1:3, vars[5]] <- perform_anova(model1)[2:4]
usv_anova_tb[4:5, vars[5]] <- perform_anova(model2)[2:3]
post_hoc_h1_list[[5]] <- run_posthoc_test_num_h1(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[5], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)
post_hoc_h2_list[[5]] <- run_posthoc_test_num_h2(data = data |> mutate(PND = 4, Animal = as.factor(seq_len(n()))), 
                                                 var = vars[5], 
                                                 p_list_3 = p_list_3,
                                                 p_list_2 = p_list_2) |>
select(-PND)

usv_anova_tb = usv_anova_tb |> 
  column_to_rownames(var = "Variables") |>
  t() |>
  round(3)
```

# Background

In MAR 4.0 project, this section we will wants to know analysis result on the	Elevated Plus Maze data. 

We are going to run the analysis use R-package structure within github. R 4.4.3 is used for the analysis, and we are wrapping all the materials into a R package.

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with a final sample size of `r epm |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. 

The interested variables at EPM data are: `r paste(vars, collapse = ", ")`.

To prepare the data, we adjusted the proportion of time spent in open arms to ensure all values lie strictly within (0, 1):

$$
\text{Adjusted Time Spend Open Arms (\%)} = \frac{\text{Time Spend Open Arms (\%)} \times (n - 1) + 0.0001}{n}
$$

We also log-transformed the `Open Arms` variable, adding half of the smallest non-zero value to avoid issues with zero:

$$
\log_2(\text{Open Arms} + \frac{1}{2} \cdot \min(\text{Open Arms} > 0))
$$

# Summary statistics

The all interested variables were summarized by treatment group overall (Table \@ref(tab:summary-tbl)), and by treatment groups stratified by sex (Table \@ref(tab:summary-tbl-sex)).

```{r summary-tbl}
tx_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of USV by Treatment Groups."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 1.16
  )
```

\

```{r summary-tbl-sex}
tx_sex_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of USV by Treatment Groups, stratified",
      "by Sex."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 0.7
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```

# Model Results

         
The relationship between treatment groups and variables of interest was assessed using linear mixed-effects models (Open Arms and Closed Arms), "Open Arms" takes a log2 transformation () to meet the normality assumption. Negative binomial regression is used to fitting counting variables (Enter Open Arms and SUM CHECK). Beta distribution is used to model the proportion variable "Time Spend Open Arms (%)", and this variable took a transformation to scale up to avoid 0's in the proportion.

The regression fitting process followed these steps:

- Fit the interaction: Begin with a model interactions.

- Evaluate the interaction: If the interaction is not significant, proceed with a model without interaction.

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report's delivery format.

## Post Hoc Analysis and P-Value Reporting

To evaluate pairwise differences between treatment groups across postnatal day (PND) levels, we performed post hoc tests tailored to the data type. The resulting tables report p-values from both the interaction model (inter) and the additive model (add). For each model, both raw and adjusted p-values were provided, with adjustment performed using the Hochberg method. The p-value adjustment was applied separately within each time period (PND = 4, 8, 12) and within each hypothesis type (primary or secondary).

For categorical outcomes, the proportion difference between groups was computed along with its standard error using the pooled standard error formula for two proportions. A 95% confidence interval for the difference was calculated using the normal approximation.

A custom helper function was used to format and annotate p-values: values less than 0.05 were marked with an asterisk (*), and all values were rounded to three decimal places to enhance readability.

**Note:** Post hoc analysis p-values are derived from the `emmeans` package, which calculates *estimated marginal means* (least-squares means) based on the fitted model. These p-values reflect pairwise comparisons between adjusted group means, accounting for other covariates or factors included in the model. Therefore, they may differ from the p-values reported in the main model summary (e.g., from ANOVA tables or survival model estimates), which test overall effects (e.g., for the entire Treatment factor) rather than specific comparisons. A non-significant main effect does not rule out significant pairwise comparisons between individual groups, and vice versa. Post hoc p-values should be interpreted accordingly, in the context of the specific contrasts being tested.


## ANOVA Table

Type III ANOVA table (Table \@ref(tab:anova-tb)) display the anova p-values for each variables at each models, and below is the explanation of each model:

Two-way interaction model (last column) show the p-value on the addition two-way interaction: y = trt + sex + *trt x sex*

Final additive model (last column) show the ANOVA p-value on the models: y = trt + sex


```{r anova-tb}
usv_anova_tb |>
  as.data.frame() |>
  tibble::rownames_to_column("Variable") |>
  flextable::flextable() |>
  flextable::set_caption(
    caption = paste0(
      "ANOVA P-Value Table of USV Object in the Three Different Model ",
      "(Interaction, Additive, and Basic)"
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 1.5
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```

## Estimation Table

The overall estimation tables also displayed here for all the models for an easier comparison. The no interaction model is also displayed for examine purpose.


### EXAMPLE - How to interpret the ANOVA table

The main effect of Treatment has a p-value of 0.001 in both the 3-way and 2-way models, indicating a highly statistically significant effect. This means that, regardless of Sex or PND (Postnatal Day), the outcome variable differs significantly across the treatment groups. In other words, the choice of treatment alone has a meaningful impact on the outcome, and this finding is robust across both models. 

### EXAMPLE - How to interpret the Posthoc table

Post hoc pairwise comparisons were conducted to examine differences between treatment groups at each postnatal day (PND) using chi-squared or Fisher's exact tests, depending on cell sizes. The table reports the proportion differences, standard errors, and 95% confidence intervals for each comparison. For example, at PND 4, the comparison between CRMP1+CRMP2 and the control group (Adjuvant+Saline) yielded a proportion difference of 0.215, with a 95% confidence interval of [0.048, 0.382] and an adjusted p-value of 0.028*, indicating a statistically significant increase in the response for the treatment group. P-values were adjusted using the Bonferroni method to control for multiple comparisons within each time point and hypothesis set. Asterisks next to p-values denote statistical significance at the 0.05 level.

```{r results='asis'}
# Loop to generate flextables
for (i in 1:length(vars)) {
  cat("\n\n### ---", vars[i], "---\n\n")
  merged_summary_list[[i]] |>
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    flextable::flextable() |>
    flextable::width(
      width = 1.15
    ) |>
    flextable::width(
      j = 1,
      width = 1.5
    ) |>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, 
             ": Model Estimation Result in the Model - ", 
             vars[i] ,"\n"))
  
  post_hoc_h1_list[[i]]|>
    as.data.frame(row.names = NULL) |>
    flextable::flextable() |>
    flextable::set_table_properties(layout = "autofit") |>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, 
             ": Post-hoc test for comparing the control null group (Adjuvant+Saline) to treatments with a control 95% family-wise confidence level - ", 
             vars[i] ,"\n"))

  post_hoc_h2_list[[i]]|>
    as.data.frame(row.names = NULL) |>
    flextable::flextable() |>
    flextable::set_table_properties(layout = "autofit") |>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, ": Post-hoc test for comparing across treatments with a control 95% family-wise confidence level - ", vars[i] ,"\n\n"))
  cat(paste0("\n- - - - - - - - - - - - - - - - - - - - - -\n"))
  cat(paste0("\n- - - - - - - - - - - - - - - - - - - - - -\n"))
}
```
