---
title: "DRAFT -- TECHNICAL MEMORANDUM"
subtitle: "<b>TO: </b> Kyoungmi Kim <br> <b> FROM: </b> Matt Ponzini <br> <b> SUBJECT:</b> MAR 4.0 Longitudinal Analyses <br> <b>DATE:</b> `r paste0(' ', Sys.Date())`"
output: 
  bookdown::html_document2:
    number_sections: TRUE
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{=html}
<style type="text/css">

h1.title {
  font-size: 24px;
  font-weight: bolder;
}

h1 { /* Header 1 */
  font-size: 16px;
  font-weight: bold;
}

h2 { /* Header 2 */
  font-size: 14px;
  font-weight: bold;
}

h3 { /* Header 3 */
  font-size: 13px;
  font-weight: bold;
}

h3.subtitle { /* Header 3 Subtitle */
  font-size: 20px;
  font-weight: normal;
  line-height: 150%;
}

</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)

# load packages
pacman::p_load(dplyr, tibble, tidyr, ggplot2, gtsummary, mar4.0, lmerTest)

# update 'asis' chunk to allow inline code
knitr::knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knitr::knit_child(text = options$code)
})
  

internal <- FALSE
```

<hr style="height:1px;color:black;background-color:black">

```{r sysinfo, include=internal}
# displays R version, platform, base packages, attached packages, and packages
# loaded via a namespace
sessionInfo()
```

# STUDY DESCRIPTION/OBJECTIVE

Behavior differences between Control and MAR rodents. MAR rodents are further classified by the auto-antibody cocktail used to induce MAR.

# DATA PROCESSING

```{r data-import, include=internal}
# if data import file(s) is (are) large, consider caching the data
#   see https://bookdown.org/yihui/rmarkdown-cookbook/cache.html for more 
#   information


```

```{r data-processing, include=internal}
milestones2 <- milestones |>
  dplyr::mutate(
    PND = factor(PND, levels = c('4', '8', '12'))
  )
```

-   Cross-fostered animals are excluded from analysis.  
-   Data are transformed to long format.  
-   Time (PND) is treated as continuous and categorical for comparison.

# STATISTICAL ANALYSIS

```{r data-analysis, include=internal}
# if analysis is computationally intensive there are a few options.
#   1) run analysis and save all results to load here
#   2) run analysis here and set chunk option cache=TRUE. A cached chunk will
#      automatically load output and objects from a previous run.
#         see https://bookdown.org/yihui/rmarkdown-cookbook/cache.html for more 
#         information

fit1 <- lmerTest::lmer(
  `Body Length` ~ PND*Treatment*Sex + (1|`Litter ID`) + (1|`Animal ID`),
  data = milestones
)

fit1_cat <- lmerTest::lmer(
  `Body Length` ~ PND*Treatment*Sex + (1|`Litter ID`) + (1|`Animal ID`),
  data = milestones2
)


fit1_summary <- broom.mixed::tidy(fit1) |>
  dplyr::select(-c(effect, group, statistic, df)) |>
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  tidyr::unite(col = `Estimate (SE)`, estimate, std.error, sep = " (") |>
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ")"),
    p.value = dplyr::case_when(
      p.value == "0.000" ~ "<0.001",
      .default = p.value
    )
  )

fit1_cat_summary <- broom.mixed::tidy(fit1_cat) |>
  dplyr::select(-c(effect, group, statistic, df)) |>
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  tidyr::unite(col = `Estimate (SE)`, estimate, std.error, sep = " (") |>
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ")"),
    p.value = dplyr::case_when(
      p.value == "0.000" ~ "<0.001",
      .default = p.value
    )
  )


fit1_anova <- car::Anova(fit1, type = 3)

fit1_cat_anova <- car::Anova(fit1_cat, type = 3)

# fit with PND:TX interaction only
fit2 <- lmerTest::lmer(
  `Body Length` ~ PND + Treatment + Sex + PND*Treatment + (1|`Litter ID`) + (1|`Animal ID`),
  data = milestones
)

fit2_cat <- lmerTest::lmer(
  `Body Length` ~ PND + Treatment + Sex + PND*Treatment + (1|`Litter ID`) + (1|`Animal ID`),
  data = milestones2
)


fit2_summary <- broom.mixed::tidy(fit2) |>
  dplyr::select(-c(effect, group, statistic, df)) |>
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  tidyr::unite(col = `Estimate (SE)`, estimate, std.error, sep = " (") |>
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ")"),
    p.value = dplyr::case_when(
      p.value == "0.000" ~ "<0.001",
      .default = p.value
    )
  )

fit2_cat_summary <- broom.mixed::tidy(fit2_cat) |>
  dplyr::select(-c(effect, group, statistic, df)) |>
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  tidyr::unite(col = `Estimate (SE)`, estimate, std.error, sep = " (") |>
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ")"),
    p.value = dplyr::case_when(
      p.value == "0.000" ~ "<0.001",
      .default = p.value
    )
  )


fit2_anova <- car::Anova(fit2, type = 3)

fit2_cat_anova <- car::Anova(fit2_cat, type = 3)

# fit main effects only
fit3 <- lmerTest::lmer(
  `Body Length` ~ PND + Treatment + Sex + (1|`Litter ID`) + (1|`Animal ID`),
  data = milestones
)

fit3_summary <- broom.mixed::tidy(fit3) |>
  dplyr::select(-c(effect, group, statistic, df)) |>
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  tidyr::unite(col = `Estimate (SE)`, estimate, std.error, sep = " (") |>
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ")"),
    p.value = dplyr::case_when(
      p.value == "0.000" ~ "<0.001",
      .default = p.value
    )
  )

fit3_anova <- car::Anova(fit3, type = 3)

# fit without sex main effect
fit4 <- lmerTest::lmer(
  `Body Length` ~ PND + Treatment + PND*Treatment + (1|`Litter ID`) + (1|`Animal ID`),
  data = milestones
)

fit4_summary <- broom.mixed::tidy(fit4) |>
  dplyr::select(-c(effect, group, statistic, df)) |>
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  tidyr::unite(col = `Estimate (SE)`, estimate, std.error, sep = " (") |>
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ")"),
    p.value = dplyr::case_when(
      p.value == "0.000" ~ "<0.001",
      .default = p.value
    )
  )

fit4_anova <- car::Anova(fit4, type = 3)

# fit with sex and time*sex
fit5 <- lmerTest::lmer(
  `Body Length` ~ PND + Treatment + Sex + PND*Sex + (1|`Litter ID`) + (1|`Animal ID`),
  data = milestones
)

fit5_summary <- broom.mixed::tidy(fit5) |>
  dplyr::select(-c(effect, group, statistic, df)) |>
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  tidyr::unite(col = `Estimate (SE)`, estimate, std.error, sep = " (") |>
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ")"),
    p.value = dplyr::case_when(
      p.value == "0.000" ~ "<0.001",
      .default = p.value
    )
  )

fit5_anova <- car::Anova(fit5, type = 3)
```

Generalized linear mixed effects models (with a identity or logit link function) were fit with fixed effects for time (PND), sex, and treatment; and random effects for the individual and the litter. To explore the potential for sex differences in the treatment groups overtime, a three-way interaction was included.



# RESULTS

## Time as Continuous

When modeling time as continuous, the association between treatment and body length is not statistically significant.

```{r, message=FALSE, warning=FALSE}
dplyr::bind_cols(fit1_anova, 
                 fit2_anova |> 
                   dplyr::add_row() |> 
                   dplyr::add_row() |> 
                   dplyr::add_row(), 
                 fit3_anova |> 
                   dplyr::add_row() |> 
                   dplyr::add_row() |> 
                   dplyr::add_row() |>
                   dplyr::add_row()
                ) |>
  suppressMessages() |>
  kableExtra::kbl(
    col.names = c("Term", rep(c("Chisq", "df", "p-value"), 3))
  ) |>
  kableExtra::kable_styling()
```

## Time as Categorical

When modeling time as a factor, there is a significant association between tail length and treatment over time.

```{r}
dplyr::bind_cols(fit1_cat_anova, 
                 fit2_cat_anova |> 
                   dplyr::add_row() |> 
                   dplyr::add_row() |> 
                   dplyr::add_row()
                ) |>
  suppressMessages() |>
  kableExtra::kbl(
    col.names = c("Term", rep(c("Chisq", "df", "p-value"), 2))
  ) |>
  kableExtra::kable_styling()
```


## Is the relationship linear over time?

Explore the relationship of outcome variables over time. Do the outcome variables change linearly?

```{r fig.cap="Change in Body Weight over time by Treatment Group."}
p1 <- ggplot(data = milestones) +
    aes(
      x = PND,
      y = `Body Weight`,
      color = Treatment,
      linetype = Treatment,
      shape = Treatment
    ) +
    geom_jitter() +
    geom_smooth(aes(group = Treatment), method = 'loess', se = FALSE) +
    # geom_smooth(method = 'lm', se = FALSE) +
    theme_bw() +
  theme(legend.direction = 'horizontal')

p_legend <- cowplot::get_legend(p1)
p1 <- p1 + ggplot2::theme(legend.position = 'none')

p2 <- ggplot(data = milestones) +
    aes(
      x = PND,
      y = `Body Weight`,
      color = Treatment,
      linetype = Treatment,
      shape = Treatment
    ) +
    geom_jitter() +
    # geom_smooth(method = 'loess', se = FALSE) +
    geom_smooth(method = 'lm', se = FALSE) +
    theme_bw() +
  theme(legend.position = 'none')

cowplot::plot_grid(
  cowplot::plot_grid(p1, p2, nrow = 1),
  p_legend,
  ncol = 1,
  rel_heights = c(0.7, 0.1)
)
```

<br>
<br>

```{r fig.cap="Change in Tail Length over time by Treatment Group."}
p1 <- ggplot(data = milestones) +
    aes(
      x = PND,
      y = `Tail Length`,
      color = Treatment,
      linetype = Treatment,
      shape = Treatment,
      group = `Animal ID`
    ) +
    geom_jitter() +
    geom_smooth(aes(group = Treatment), method = 'loess', se = FALSE) +
    # geom_smooth(method = 'lm', se = FALSE) +
    theme_bw() +
  theme(legend.direction = 'horizontal')

p_legend <- cowplot::get_legend(p1)
p1 <- p1 + ggplot2::theme(legend.position = 'none')

p2 <- ggplot(data = milestones) +
    aes(
      x = PND,
      y = `Tail Length`,
      color = Treatment,
      linetype = Treatment,
      shape = Treatment
    ) +
    geom_jitter() +
    # geom_smooth(method = 'loess', se = FALSE) +
    geom_smooth(method = 'lm', se = FALSE) +
    theme_bw() +
  theme(legend.position = 'none')

cowplot::plot_grid(
  cowplot::plot_grid(p1, p2, nrow = 1),
  p_legend,
  ncol = 1,
  rel_heights = c(0.7, 0.1)
)

```

<br>
<br>

```{r fig.cap="Change in Body Length over time by Treatment Group."}
p1 <- ggplot(data = milestones) +
    aes(
      x = PND,
      y = `Body Length`,
      color = Treatment,
      linetype = Treatment,
      shape = Treatment
    ) +
    geom_jitter() +
    geom_smooth(method = 'loess', se = FALSE) +
    # geom_smooth(method = 'lm', se = FALSE) +
    theme_bw() +
  theme(legend.direction = 'horizontal')

p_legend <- cowplot::get_legend(p1)
p1 <- p1 + ggplot2::theme(legend.position = 'none')

p2 <- ggplot(data = milestones) +
    aes(
      x = PND,
      y = `Body Length`,
      color = Treatment,
      linetype = Treatment,
      shape = Treatment
    ) +
    geom_jitter() +
    # geom_smooth(method = 'loess', se = FALSE) +
    geom_smooth(method = 'lm', se = FALSE) +
    theme_bw() +
  theme(legend.position = 'none')

cowplot::plot_grid(
  cowplot::plot_grid(p1, p2, nrow = 1),
  p_legend,
  ncol = 1,
  rel_heights = c(0.7, 0.1)
)

```


## Model Diagnostics

Compare model diagnostics for full models (three-way interaction between time, sex, and treatment).

```{r, fig.cap="Model Diagnostics with Time as Continuous."}
performance::check_model(
  fit1,
  check = c('linearity', 'normality', 'homogeneity', 'qq')
)
```

<br>
<br>

```{r, fig.cap = "Nodel Diagnostics with Time as Factor."}
performance::check_model(
  fit1_cat,
  check = c('linearity', 'normality', 'homogeneity', 'qq')
)
```
