---
title: "DRAFT -- TECHNICAL MEMORANDUM"
subtitle: "<b>TO: </b> Kyoungmi Kim <br> <b> FROM: </b> Matt Ponzini <br> <b> SUBJECT:</b> MAR 4.0 Longitudinal Analyses <br> <b>DATE:</b> `r paste0(' ', Sys.Date())`"
output: 
  bookdown::html_document2:
    number_sections: TRUE
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
---

<style type="text/css">

h1.title {
  font-size: 24px;
  font-weight: bolder;
}

h1 { /* Header 1 */
  font-size: 16px;
  font-weight: bold;
}

h2 { /* Header 2 */
  font-size: 14px;
  font-weight: bold;
}

h3 { /* Header 3 */
  font-size: 13px;
  font-weight: bold;
}

h3.subtitle { /* Header 3 Subtitle */
  font-size: 20px;
  font-weight: normal;
  line-height: 150%;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)

# load packages
pacman::p_load(dplyr, tibble, tidyr, ggplot2, gtsummary, mar4.0, lmerTest)

# update 'asis' chunk to allow inline code
knitr::knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knitr::knit_child(text = options$code)
})
  

internal <- FALSE
```

<hr style="height:1px;color:black;background-color:black">

# STUDY DESCRIPTION/OBJECTIVE

Behavior differences between Control and MAR rodents. MAR rodents are further classified by the auto-antibody cocktail used to induce MAR.

# DATA PROCESSING

-   Cross-fostered animals are excluded from analysis.  
-   Data are transformed to long format.  
-   Time (PND) is treated as categorical (4, 8, or 12).


# STATISTICAL ANALYSIS

Generalized linear mixed effects models (with a identity or logit link function) were fit with fixed effects for time (PND), sex, and treatment; and random effects for the individual and the litter. We begin with a full model with a three-way interaction between time, treatment group, and sex, and all nested terms. <b>If the three-way interaction is not significant...</b>

# Results


```{r}
fit1 <- lmerTest::lmer(`Body Length` ~ PND * Treatment * Sex + (1|`Animal ID`) + (1|`Litter ID`),
                       data = milestones)

gtsummary::tbl_regression(fit1) |>
  gtsummary::add_global_p()
```

```{r warning=FALSE, message=FALSE}
fit2 <- lmerTest::lmer(`Body Length` ~ PND + Treatment + Sex + PND * Treatment +
                         PND * Sex + Treatment * Sex +
                         (1|`Animal ID`) + (1|`Litter ID`),
                       data = milestones)

fit2_tbl <- gtsummary::tbl_regression(fit2) |>
  gtsummary::add_global_p()

fit3 <- lmerTest::lmer(`Body Length` ~ PND + Treatment + Sex + PND * Treatment + 
                         (1|`Animal ID`) + (1|`Litter ID`),
                       data = milestones)

fit3_tbl <- gtsummary::tbl_regression(fit3) |>
  gtsummary::add_global_p()

gtsummary::tbl_merge(tbls=  list(fit2_tbl, fit3_tbl))
```
