---
title: "MAR 4.0 - Milestones"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2025-02-28"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, rstatix, ggpubr, mar4.0)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
fig_n = 0
tab_n = 0
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glm(as.formula(formula_str), data = data, family = binomial(link = "logit")))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}

# Function to summarize and format model results
format_summary <- function(model_summary, coef_indices, column_name) {
  formatted <- apply(model_summary$coefficients[, coef_indices], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(summary_full, ...) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  
  for (i in 2:length(summaries)) {
    merged_data <- full_join(merged_data, summaries[[i]], by = "Effect")
  }
  
  # Convert numeric columns (p-values) to 4 decimal places (non-scientific)
  numeric_cols <- sapply(merged_data, is.numeric)
  merged_data[numeric_cols] <- lapply(merged_data[numeric_cols], function(x) {
    ifelse(is.na(x), "", as.character(format(round(x, 4), nsmall = 4, scientific = FALSE)))
  })
  
  return(merged_data)
}

# Function to return the gtsummary-table by input dataset and outcome
summary_tbls <- function(
    dataset, outcome
) {

  tbl_male <- dataset |>
    dplyr::filter(Sex == "M") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          ) 
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Males"
    )

  tbl_female <- dataset |>
    dplyr::filter(Sex == "F") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          )
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Females"
    )

  ans <- list(
    "males" = tbl_male,
    "females" = tbl_female
  )

  return(ans)
}
```

```{r}
###################### Run the model and create plot for numeric variables

# Ensure PND is a factor with reference level set to "4"
data = milestones |>
  mutate(Startle = ifelse(Startle=="N",0,1),
         `Forelimb Grasp` = ifelse(`Forelimb Grasp`=="N",0,1),
         PND = relevel(factor(PND), ref = "4"),
         `Circle Traverse 30s` = ifelse(`Circle Traverse`<30,1,0) |> as.factor(),
         `Negative Geotaxis 30s` = ifelse(`Negative Geotaxis`<30,1,0)|> as.factor(),
         `Cliff Avoidance 30s` = ifelse(`Cliff Avoidance`<30,1,0)|> as.factor(),
         `Log2 Righting Reflex` = log2(`Righting Reflex`))


# Variables to analyze, including Count
vars = names(data)[11:24]
vars_numeric = c(1:8,14)
vars_binary = c(9,11,12,13)

# Create ANOVA results table
merged_summary_list <- list()
plot_list <- list()
post_hoc_h1_list <- list()
post_hoc_h2_list <- list()

# Main loop over variables
for (i in vars_numeric) {
  var <- vars[i]
  
  # Data preparation (ensure complete cases)
  data_fct_compete <- data |>
    dplyr::mutate(
      Treatment = factor(Treatment, levels = c(
        "Adjuvant+Saline", "CRMP1+CRMP2", "CRMP1+GDA",
        "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE"
      )),
      Sex = factor(Sex, levels = c("M", "F")),
      Litter = factor(`Litter ID`),
      Animal = factor(`Animal ID`)
    ) |>
    dplyr::filter(dplyr::if_all(c(Animal, Litter, Treatment, Sex, PND, !!sym(var)), ~ !is.na(.x)))
  
  # Perform 3-way ANOVA (Treatment * Sex * PND)
  mix_aov_3way <- rstatix::anova_test(
    dv = !!sym(var), 
    wid = Animal, 
    between = c(Treatment, Sex), 
    within = PND, 
    data = data_fct_compete, 
    type = 3
  )

  # Perform 2-way ANOVA (Treatment * PND)
  mix_aov_2way <- rstatix::anova_test(
    dv = !!sym(var), 
    wid = Animal, 
    between = c(Treatment), 
    within = PND,
    covariate = Sex,
    data = data_fct_compete, 
    type = 3
  )

  # Extract p-values from ANOVA tests and merge them into one table
  p_values_3way <- mix_aov_3way$ANOVA[,c("Effect","p")] |> dplyr::rename(`P-Value (3-Way Model)` = p)  
  p_values_2way <- mix_aov_2way$ANOVA[,c("Effect","p")] |> dplyr::rename(`P-Value (2-Way Model)` = p) 
  combined_anova_results <- create_combined_estimate(p_values_3way, p_values_2way)
  
  ######## Store results in list format for further use
  merged_summary_list[[i]] <- combined_anova_results
  
  ######## Create Boxplots for numeric variable
  plot_list[[i]] <- ggboxplot(
  data_fct_compete, x = "Treatment", y = var,
  color = "Sex", fill = "Sex", palette = "Set1"
) +
  facet_wrap(~PND) +
  theme_minimal() +
  labs(title = paste0("Distribution of ", var, " by Treatment & Sex"),
       x = "Treatment",
       y = paste0(var, " Value")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
  
  ######## Running post-hoc test on continuous variable
  post_hoc_h1_list[[i]] <- run_posthoc_test_num_h1(data = data_fct_compete, vars = var)
  post_hoc_h2_list[[i]] <- run_posthoc_test_num_h1(data = data_fct_compete, vars = var)
}
```

```{r}
###################### Run the model and create plot for binary variables

merged_estimate_list = list()

for (i in vars_binary) {
  var <- vars[i]
  model_type = ifelse(i == 9, "glm","glmer")

  if (model_type == "glm"){
    # Fit models
    formula1 <-  paste0("`", var, "` ~ `Treatment` + `Sex` + `Treatment`*`Sex` ")
    formula2 <-  paste0("`", var, "` ~ `Treatment` + `Sex` ")
    model1 <- glm(as.formula(formula1), data = data, family = binomial(link = "logit"))
    model2 <- glm(as.formula(formula2), data = data, family = binomial(link = "logit"))
    
    # Summarize and format results
    coef_indices <- if (model_type == "glm") c(1, 2, 4) else c(1, 2, 5)
    summary1 <- format_summary(summary(model1), coef_indices, "Two Vars Model") |> tibble::rownames_to_column()
    names(summary1) <- c("Effect", names(summary1)[2])
    summary2 <- format_summary(summary(model2), coef_indices, "Main Effect Model")|> tibble::rownames_to_column()
    names(summary2) <- c("Effect", names(summary2)[2])
    merged_estimate <- create_combined_estimate(summary_full = summary1, summary2)
    merged_estimate_list[[i]] <- merged_estimate
    
    # Perform ANOVA
    anova_result_1 = Anova(model1, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_1 = anova_result_1[,c(1,4)]
    names(anova_result_1) <- c("Effect", "Two Vars Model")
    anova_result_2 = Anova(model2, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_2 = anova_result_2[,c(1,4)]
    names(anova_result_2) <- c("Effect", "Main Effect Model")
    combined_anova_results <- create_combined_estimate(summary_full = anova_result_1, anova_result_2)
    merged_summary_list[[i]] <- combined_anova_results
  } else {
    formula1 <-  paste0("`", var, "` ~ `Treatment` * Sex * `PND` + (1|`Animal ID`)")
    formula2 <-  paste0("`", var, "` ~ `Treatment` + Sex + PND + `Treatment`:PND + Sex:PND + (1|`Animal ID`)")
    model1 <- glmer(as.formula(formula1), data = data, family = binomial(link = "logit"),
                    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))
    model2 <- glmer(as.formula(formula2), data = data, family = binomial(link = "logit"),
                    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))
    anova_result_1 = Anova(model1, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_1 = anova_result_1[,c(1,4)]
    names(anova_result_1) <- c("Effect", "Two Vars Model")
    anova_result_2 = Anova(model2, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_2 = anova_result_2[,c(1,4)]
    names(anova_result_2) <- c("Effect", "Two Vars Model")
    combined_anova_results <- create_combined_estimate(summary_full = anova_result_1, anova_result_2)
    merged_summary_list[[i]] <- combined_anova_results
  }

  # Create and store bar plot
  plot_summary <- data |> 
    select(Treatment, PND, Sex, !!sym(var)) |> 
    mutate(Treatment = as.factor(Treatment),
           Sex = as.factor(Sex),
           PND = as.factor(PND),
           !!sym(var) :=  ifelse(!!sym(var) == 1,1,0)) |>
    filter(!is.na(!!sym(var))) |> 
    group_by(Treatment, Sex, PND) |> 
    summarise(Proportion_Yes = mean(!!sym(var)), .groups = "drop")

  plot_list[[i]] <- ggplot(plot_summary, aes(x = Treatment, y = Proportion_Yes, fill = Sex, color = Sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~PND) +
  scale_fill_brewer(palette = "Set1") +  
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  labs(title = paste0("Proportion of 'Yes' in ", var, " by Treatment & Sex"),
       x = "Treatment",
       y = paste0(var, " Value")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        panel.grid.major = element_line(color = "gray85"),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(), 
        strip.text = element_text(face = "bold")) 
  
  post_hoc_h1_list[[i]] <- run_posthoc_test_cat_h1(data = data_fct_compete, var = var)
  post_hoc_h2_list[[i]] <- run_posthoc_test_cat_h2(data = data_fct_compete, var = var)
}





```


# Background

In MAR 4.0 project, this section we will wants to know analysis result on the milestone data. 

We are going to run the analysis use R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with a total observation `r milestones |> dplyr::pull(dplyr::any_of('Animal ID')) |> length()`, and total number of animals `r milestones |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. Finally, the `Startle` and `Forelimb Grasp` variables were converted to binary for analysis.

We also dichotomized three variables (`Circle Traverse`, `Negative Geotaxis`, and `Cliff Avoidance`) to if the animal achieved the milestone within 30 seconds, and those three variables are evaluated in both numeric and binary variable version.

The interested variables at milestones data are: `r paste(vars, collapse = ", ")`.

# Summary statistics

The all interested variables were summarized by Treatment group overall, stratified by Time and Sex.

```{r summary-tbl, results='asis'}
for (i in seq_along(vars)) {
  summary_list <- summary_tbls(data, vars[i])
  
  purrr::walk2(summary_list, vars[i], function(tbl, var_name) {
    tbl |>
      gtsummary::as_flex_table() |>
      flextable::width(width = 1.15) |>
      flextable::width(j = 1, width = 1.16) |>
      flextable::flextable_to_rmd()
    
    tab_n <<- tab_n + 1  
    cat(paste0("\nTable ", tab_n, 
               ": Summary of ", 
               var_name, 
               " by Treatment Groups with Time Stratification (4, 8, and 12th week).\n\n"))
  })
}
```

# Model Results

To evaluate differences in mean milestone measures between treatments over time, we conducted repeated measures ANOVA (or logistic regression for a cross-sectional categorical milestone measure), incorporating animals as a random effect.

We fitted 2 models as follows in order: 

- Fit the 3-way interaction ANOVA model: Begin with a 3-way interaction model with the main effects of time, treatment (trt), and sex, their 2-way interactions and the three-way interaction of all three variables. This model assumes sex as an effect modifier. 

- If the 3-way interaction term in ANOVA is not statistically significant (i.e, if sex is not an effect modifier), we drop the 3-way interaction term and fit the 2-way interaction model that assumes sex as a confounding variable. 

Additionally, for binary outcomes (Startle), we use ordinary logistic regression to evaluate the effect on the 12th week, this is because no animal has Startle at the 4th and 8th week (this is non-repeated measure). Forelimb Grasp is no evaluated because no animal has Startle at any time point.

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report’s delivery format.

## ANOVA Table

Type III ANOVA table displays the anova p-values for each variable in each model, and below is the explanation of each continuous outcome model:

- Three interaction model shows the p-value on three way interaction: y = time(PND) * trt * sex

- Two-way interaction model shows the p-value without the three way interaction: y = time(PND) * trt + Sex + time(PND) : Sex


**OR** for binary outcome models:

- Three interaction model shows the p-value on three way interaction: y = trt * sex

- Two-way interaction model shows the p-value without the three way interaction: y = trt + sex


The models also account for within-subjects variation attributed to animals (i.e, inter-animal variability within the same treatment group), and all p-values are calculated through Type-III sum of squares of residuals, using the R “rstatix” package.

Additionally, we use boxplots for intuitive visualizations for cross-sectional pairwise group comparisons at each time point tested by t-test with a combination of Bonferroni adjustment method (more * means more significance of the pairwise t-test).

- *Some Righting Reflex are excluded (n = 4): At PND 4, 9_42_2 has 55.8, at PND 8, 9_42_7 has 27.4, and at PND 12, 9_15_8 and 9_45_1 have 6.06 and 5.95.*

## Post-hoc Tests

Post-hoc tests are multiple pairwise comparisons between groups to identify which groups are different. Based on our hypothesis, we designed two set of post-hoc tests for each of variable. The first set is all comparison between the control group and each treatment group, and the second set is across all treatment group. The p-values are adjusted by Bonferroni method on each set of comparison and on each time period.

```{r anova-tb, results='asis'}
set_flextable_defaults(
  font.family = "Arial",  # Optional: Set default font
  font.size = 10,         # Optional: Set default font size
  theme_fun = theme_vanilla # Optional: Set default theme
)

for (i in c(vars_numeric,vars_binary)) {
  # Create flextable summary table
  merged_summary_list[[i]] |>
    as.data.frame(row.names = NULL) |>
    flextable::flextable() |>
    flextable::set_table_properties(layout = "autofit") |>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, ": Model Estimation ANOVA Result - ", vars[i] ,"\n\n"))
  
  print(plot_list[[i]])
  fig_n = fig_n + 1
  cat(paste0("\nFigure ",fig_n, ": Summary Plot of Response Variable by Treatment, Time, and Sex\n\n"))
  
  post_hoc_h1_list[[i]]|>
    flextable::flextable_to_rmd()
    tab_n = tab_n + 1
    cat(paste0("\nTable ",tab_n, ": Post-hoc test for comparing baseline group to treatments with a control 95% family-wise confidence level - ", vars[i] ,"\n"))

  post_hoc_h2_list[[i]]|>
    flextable::flextable_to_rmd()
    tab_n = tab_n + 1
    cat(paste0("\nTable ",tab_n, ": Post-hoc test for comparing across treatments with a control 95% family-wise confidence level - ", vars[i] ,"\n\n"))
    cat(paste0("\n- - - - - - - - - - - - - - - - - - - - - -\n"))
}



```

