---
title: "MAR 4.0 - Milestones"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2024-10-29"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, mar4.0)
```

```{r}
tx_tbl <- milestones |>
  dplyr::select(
    Treatment, `Body Length`, `Tail Length`, `Belly Temp`, `Body Weight`, `Righting Reflex`, 
    `Circle Traverse`, `Negative Geotaxis`, `Cliff Avoidance`, Startle, `Forelimb Grasp`) |>
  gtsummary::tbl_summary(
    by = Treatment,
    statistic = list(
      gtsummary::all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} [{p25}, {p75}]",
        "{min}, {max}"
      )
    )
  ) |>
  gtsummary::add_overall()

tx_sex_tbl <- milestones |>
  dplyr::select(
    Treatment, Sex, `Body Length`, `Tail Length`, `Belly Temp`, `Body Weight`, `Righting Reflex`, 
    `Circle Traverse`, `Negative Geotaxis`, `Cliff Avoidance`, Startle, `Forelimb Grasp`) |>
  gtsummary::tbl_strata(
    strata = Sex,
    .tbl_fun = 
      ~ .x |>
      gtsummary::tbl_summary(
        by = Treatment,
        statistic = list(
          gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})",
            "{median} [{p25}, {p75}]",
            "{min}, {max}"
          )
        )
      ) |>
      gtsummary::add_overall() |>
      gtsummary::modify_header(
        gtsummary::all_stat_cols() ~ "**{level}**"
      ),
    .combine_with = 'tbl_stack'
  )

```


```{r}
data = milestones |>
  mutate(Startle = ifelse(Startle=="N",0,1),
         `Forelimb Grasp` = ifelse(`Forelimb Grasp`=="N",0,1))
vars = names(milestones)[11:20]

milestone_anova_tb <- data.frame(Variables = c("Interaction Model - Treatment","Interaction Model - Sex","Interaction Model - Interaction",
                                                 "Additive Model - Treatment", "Additive Model - Sex", "Basic Model - Treatment"))
merged_summary_list <- list()
# Loop through each variable
for (i in 1:length(vars)) {
  # Fit the mixed effects model
  if (vars[i] %in% c("Startle","Forelimb Grasp")){
    lmer_model1 <- glmer(as.formula(paste0("`", vars[i], "`", " ~ `Treatment` + (1 | `Litter ID`)")), 
                         data = data, family = binomial(link = "logit"))
    lmer_summary1 <- lmer_model1 |> summary()
    lmer_model2 <- glmer(as.formula(paste0("`",vars[i],"`", "~ `Treatment` + `Sex` + (1 | `Litter ID`)")), 
                         data = data, family = binomial(link = "logit"))
    lmer_summary2 <- lmer_model2 |> summary()
    lmer_model3 <- glmer(as.formula(paste0("`",vars[i],"`", "~ `Treatment` + `Sex` + `Treatment` * `Sex` + (1 | `Litter ID`)")), 
                         data = data, family = binomial(link = "logit"))
    lmer_summary3 <- lmer_model3 |> summary()
    formatted_summary3 <- apply(lmer_summary3$coefficients[, c(1, 2, 4)], 1, function(row) {
      estimate <- round(row[1], 3)
      se <- round(row[2], 3)
      p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
      formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
      return(formatted_row)
    })
    formatted_summary2 <- apply(lmer_summary2$coefficients[, c(1, 2, 4)], 1, function(row) {
      estimate <- round(row[1], 3)
      se <- round(row[2], 3)
      p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
      formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
      return(formatted_row)
    })
    formatted_summary1 <- apply(lmer_summary1$coefficients[, c(1, 2, 4)], 1, function(row) {
      estimate <- round(row[1], 3)
      se <- round(row[2], 3)
      p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
      formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
      return(formatted_row)
    })
  } else{
    lmer_model1 <- lmerTest::lmer(as.formula(paste0("`",vars[i],"`", "~ `Treatment` + (1 | `Litter ID`)")), 
                               data = data) 
    lmer_summary1 <- lmer_model1 |> summary()
    lmer_model2 <- lmerTest::lmer(as.formula(paste0("`",vars[i],"`", "~ `Treatment` + `Sex` + (1 | `Litter ID`)")), 
                                 data = data) 
    lmer_summary2 <- lmer_model2 |> summary()
    lmer_model3 <- lmerTest::lmer(as.formula(paste0("`",vars[i],"`", "~ `Treatment` + `Sex` + `Treatment` * `Sex` + (1 | `Litter ID`)")), 
                                 data = data) 
    lmer_summary3 <- lmer_model3 |> summary()
    formatted_summary3 <- apply(lmer_summary3$coefficients[, c(1, 2, 5)], 1, function(row) {
      estimate <- round(row[1], 3)
      se <- round(row[2], 3)
      p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
      formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
      return(formatted_row)
    })
    formatted_summary2 <- apply(lmer_summary2$coefficients[, c(1, 2, 5)], 1, function(row) {
      estimate <- round(row[1], 3)
      se <- round(row[2], 3)
      p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
      formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
      return(formatted_row)
    })
    formatted_summary1 <- apply(lmer_summary1$coefficients[, c(1, 2, 5)], 1, function(row) {
      estimate <- round(row[1], 3)
      se <- round(row[2], 3)
      p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
      formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
      return(formatted_row)
    })
  }
  formatted_summary3 <- as.data.frame(t(formatted_summary3))
  formatted_summary2 <- as.data.frame(t(formatted_summary2))
  formatted_summary1 <- as.data.frame(t(formatted_summary1))
  colnames(formatted_summary3) <- c(paste0("Est(SE)-Interaction"), paste0("P-Interaction"))
  colnames(formatted_summary2) <- c(paste0("Est(SE)-Additive"), paste0("P-Additive"))
  colnames(formatted_summary1) <- c(paste0("Est(SE)-Basic"), paste0("P-Basic"))
  # Convert the result into a data frame for clarity
  merged_summary_df = formatted_summary3
  merged_summary_df$`Est(SE)-Additive` = c(formatted_summary2[,1],rep("",4))
  merged_summary_df$`P-Additive` = c(formatted_summary2[,2] ,rep("",4))
  merged_summary_df$`Est(SE)-Basic` = c(formatted_summary1[,1],rep("",5))
  merged_summary_df$`P-Basic` = c(formatted_summary1[,2] ,rep("",5))
  
  merged_summary_list[[i]] = merged_summary_df
  # Extract the ANOVA Type III p-value
  milestone_anova_tb[1:3, vars[i]] = Anova(lmer_model3, type = "III")[2:4,3]
  milestone_anova_tb[4:5, vars[i]] = Anova(lmer_model2, type = "III")[2:2,3]
  milestone_anova_tb[6, vars[i]] = Anova(lmer_model1, type = "III")[2,3]
  
#   emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treatment`,adjust = "none")
#   milestone_tb_contrast = emm_s.t$contrasts %>%
#     as.data.frame() %>%
#     select(estimate, p.value) %>%
#     mutate(across(everything(), ~ round(.x, 3)))
#   rownames(milestone_tb_contrast) <- rownames(emm_s.t$contrasts)
}

milestone_anova_tb = milestone_anova_tb |> 
  column_to_rownames(var = "Variables") |>
  t() |>
  round(3)
```


# Background

In MAR 4.0 project, this section we will wants to know analysis result on the milestone data. 

We are going to run the analysis use R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with a final sample size of `r milestones |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. Finally, the `Startle` and `Forelimb Grasp` variables were converted to binary for analysis.

# Summary statistics

The discrimination index was summarized by Treatment group overall (Table \@ref(tab:summary-tbl)), and by Treatment groups stratified by sex (Table \@ref(tab:summary-tbl-sex)).

```{r summary-tbl}
tx_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of Milestone by Treatment Groups."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 1.16
  )
```

\

```{r summary-tbl-sex}
tx_sex_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of Milestone by Treatment Groups, stratified",
      "by Sex."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 0.7
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```

# Model Results

The relationship between treatment groups and interested variables were assessed using linear mixed effects models or logistic regression with a random litter effect. First we fit a model with treatment groups, sex, and the interaction between them to assess for potential sex differences across the groups.\



## ANOVA Table

Type III ANOVA table (Table \@ref(tab:anova-tb)) display the anova p-values for each variables at each models (treatment-sex interaction, sex additive, and treatment only). There was not sufficient evidence to conclude that there are sex differences between the groups to any of the variables.

```{r anova-tb}
milestone_anova_tb |>
  as.data.frame() |>
  tibble::rownames_to_column("Variable") |>
  flextable::flextable() |>
  flextable::set_caption(
    caption = paste0(
      "ANOVA Table of Milestone Object in the Three Different Model ",
      "(Interaction, Additive, and Basic)"
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 0.7
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```

## Estimation Table

The overall estimation tables also displayed here for all the models for an easier comparison.

```{r results='asis'}
# Loop to generate flextables
for (i in 1:length(vars)) {
  merged_summary_list[[i]] |>
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    flextable::flextable() |>
    flextable::set_caption(
      caption = paste0(
        "Model Estimation Result in the Model - ", vars[i]
      )
    ) |>
    flextable::width(
      width = 1.15
    ) |>
    flextable::width(
      j = 1,
      width = 1.5
    ) |>
    flextable::flextable_to_rmd()

}
```
