---
title: "MAR 4.0 - Milestones"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2025-02-28"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, rstatix, ggpubr, ggbiplot, factoextra, survival, 
               survminer, ggplot2, stringr, mar4.0, patchwork)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
fig_n = 0
tab_n = 0
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glm(as.formula(formula_str), data = data, family = binomial(link = "logit")))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}

# Function to summarize and format model results
format_summary <- function(model_summary, coef_indices, column_name) {
  formatted <- apply(model_summary$coefficients[, coef_indices], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(summary_full, ...) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  
  for (i in 2:length(summaries)) {
    merged_data <- full_join(merged_data, summaries[[i]], by = "Effect")
  }
  
  # Convert numeric columns (p-values) to 4 decimal places (non-scientific)
  numeric_cols <- sapply(merged_data, is.numeric)
  merged_data[numeric_cols] <- lapply(merged_data[numeric_cols], function(x) {
    ifelse(is.na(x), "", as.character(format(round(x, 4), nsmall = 4, scientific = FALSE)))
  })
  
  return(merged_data)
}

# Function to return the gtsummary-table by input dataset and outcome
summary_tbls <- function(
    dataset, outcome
) {

  tbl_male <- dataset |>
    dplyr::filter(Sex == "M") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          ) 
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Males"
    )

  tbl_female <- dataset |>
    dplyr::filter(Sex == "F") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          )
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Females"
    )

  ans <- list(
    "males" = tbl_male,
    "females" = tbl_female
  )

  return(ans)
}
```

```{r}
###################### Run the model and create plot for numeric variables

# Ensure PND is a factor with reference level set to "4"
data = milestones |>
  mutate(`Animal ID` = ifelse(is.na(`Animal ID`),paste0(`Litter ID`,"_99"),`Animal ID`),
          Startle = ifelse(Startle=="N",0,1),
         `Forelimb Grasp` = ifelse(`Forelimb Grasp`=="N",0,1),
         PND = relevel(factor(PND), ref = "4"),
         `Circle Traverse 30s` = ifelse(`Circle Traverse`<30,1,0) |> as.factor(),
         `Negative Geotaxis 30s` = ifelse(`Negative Geotaxis`<30,1,0)|> as.factor(),
         `Cliff Avoidance 30s` = ifelse(`Cliff Avoidance`<30,1,0)|> as.factor(),
         `Log2 Righting Reflex` = log2(`Righting Reflex`))


# Variables to analyze, including Count
vars = names(data)[11:24]
vars_numeric = c(1:5,14)
vars_survival = c(6:8)
vars_binary = c(9,11,12,13)

# Create ANOVA results table
merged_summary_list <- list()
plot_list <- list()
post_hoc_h1_list <- list()
post_hoc_h2_list <- list()

# Main loop over variables
for (i in vars_numeric) {
  var <- vars[i]
  # Data preparation (ensure complete cases)
  data_fct_compete <- data |>
    dplyr::mutate(
      Treatment = factor(Treatment, levels = c(
        "Adjuvant+Saline", "CRMP1+CRMP2", "CRMP1+GDA",
        "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE"
      )),
      Sex = factor(Sex, levels = c("M", "F")),
      Litter = factor(`Litter ID`),
      Animal = factor(`Animal ID`)
    ) |>
    dplyr::filter(dplyr::if_all(c(Animal, Litter, Treatment, Sex, PND, !!sym(var)), ~ !is.na(.x)))
  
  # If working on the variable with 30s max recording, trimming those equal to 30s
  # Also check low counting PND and exclude those
  if (var %in% c("Circle Traverse","Negative Geotaxis","Cliff Avoidance")){
    data_fct_compete <- data_fct_compete |>
      filter(!is.na(!!sym(var)) & !!sym(var) < 30)
    
    exclude_pnd <- data_fct_compete |>
      group_by(PND, Treatment, Sex) |>
      summarise(n = n(), .groups = "drop") |>
      filter(n <= 1) |>
      pull(PND) |>
      unique()

    data_fct_compete <- data_fct_compete |>
      filter(!PND %in% exclude_pnd)
  }

  
  # Perform 3-way ANOVA (Treatment * Sex * PND)
  mix_aov_3way <- rstatix::anova_test(
    dv = !!sym(var), 
    wid = Animal, 
    between = c(Treatment, Sex), 
    within = PND, 
    data = data_fct_compete, #|> filter(PND != "12"), 
    type = 3
  )

  # Perform 2-way ANOVA (Treatment * PND)
  mix_aov_2way <- rstatix::anova_test(
    dv = !!sym(var), 
    wid = Animal, 
    between = c(Treatment), 
    within = PND,
    covariate = Sex,
    data = data_fct_compete, 
    type = 3
  )

  # Extract p-values from ANOVA tests and merge them into one table
  if (!is.null(mix_aov_3way$ANOVA)) {
    p_values_3way <- mix_aov_3way$ANOVA[, c("Effect", "p")] |>
      dplyr::rename(`P-Value (3-Way Model)` = p)
    
    p_values_2way <- mix_aov_2way$ANOVA[, c("Effect", "p")] |>
      dplyr::rename(`P-Value (2-Way Model)` = p)
  } else {
    p_values_3way <- mix_aov_3way |>
      dplyr::as_tibble() |>
      select(`Effect`, `p`) |>
      dplyr::rename(`P-Value (3-Way Model)` = p) |>
      as.data.frame()
    
    p_values_2way <- mix_aov_2way |>
      dplyr::as_tibble() |>
      dplyr::select(Effect, p) |>
      dplyr::rename(`P-Value (2-Way Model)` = p) |>
      as.data.frame()
  }
  combined_anova_results <- create_combined_estimate(p_values_3way, p_values_2way)
  
  ######## Store results in list format for further use
  merged_summary_list[[i]] <- combined_anova_results
  
  ######## Create Boxplots for numeric variable
  # Compute summary statistics (Mean & SE)
  data_summary <- data_fct_compete %>%
    group_by(Treatment, Sex, PND) %>%
    summarise(
      Mean = mean(.data[[var]], na.rm = TRUE),
      SE = sd(.data[[var]], na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )

  # Create the plot with SE
  plot_list[[i]] <- ggboxplot(data_fct_compete, x = "Treatment", y = var, color = "Sex", fill = "Sex", palette = "Set1") +
    geom_errorbar(data = data_summary, 
                  aes(x = Treatment, y = Mean, ymin = Mean - 2*SE, ymax = Mean + 2*SE, group = interaction(Treatment, Sex)),  
                  position = position_dodge(width = 0.8), 
                  width = 0.2, 
                  color = "black") +  
    geom_point(data = data_summary, 
               aes(x = Treatment, y = Mean, color = Sex, group = interaction(Treatment, Sex)), 
               position = position_dodge(width = 0.8), 
               size = 1, 
               shape = 21) +  
    facet_wrap(~PND, scales = "free_x", nrow = 1, labeller = label_both) +
    theme_minimal() +
    labs(title = paste0("Distribution of ", var, " by Treatment & Sex"),
         x = "Treatment",
         y = paste0(var, " Value")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          strip.text = element_text(size = 12, face = "bold"))


  ######## Running post-hoc test on continuous variable
  post_hoc_h1_list[[i]] <- run_posthoc_test_num_h1(data = data_fct_compete, vars = var)
  post_hoc_h2_list[[i]] <- run_posthoc_test_num_h2(data = data_fct_compete, vars = var)
}
```

```{r}
###################### Run the model and create plot for binary variables
merged_estimate_list = list()

for (i in vars_binary) {
  var <- vars[i]
  model_type = ifelse(i == 9, "glm","glmer")

  if (model_type == "glm"){
    # Fit models
    formula1 <-  paste0("`", var, "` ~ `Treatment` + `Sex` + `Treatment`*`Sex` ")
    formula2 <-  paste0("`", var, "` ~ `Treatment` + `Sex` ")
    model1 <- glm(as.formula(formula1), data = data, family = binomial(link = "logit"))
    model2 <- glm(as.formula(formula2), data = data, family = binomial(link = "logit"))
    
    # Summarize and format results
    coef_indices <- if (model_type == "glm") c(1, 2, 4) else c(1, 2, 5)
    summary1 <- format_summary(summary(model1), coef_indices, "Two Vars Model") |> tibble::rownames_to_column()
    names(summary1) <- c("Effect", names(summary1)[2])
    summary2 <- format_summary(summary(model2), coef_indices, "Main Effect Model")|> tibble::rownames_to_column()
    names(summary2) <- c("Effect", names(summary2)[2])
    merged_estimate <- create_combined_estimate(summary_full = summary1, summary2)
    merged_estimate_list[[i]] <- merged_estimate
    
    # Perform ANOVA
    anova_result_1 = Anova(model1, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_1 = anova_result_1[,c(1,4)]
    names(anova_result_1) <- c("Effect", "Two Vars Model")
    anova_result_2 = Anova(model2, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_2 = anova_result_2[,c(1,4)]
    names(anova_result_2) <- c("Effect", "Main Effect Model")
    combined_anova_results <- create_combined_estimate(summary_full = anova_result_1, anova_result_2)
    merged_summary_list[[i]] <- combined_anova_results
  } else {
    formula1 <-  paste0("`", var, "` ~ `Treatment` * Sex * `PND` + (1|`Animal ID`)")
    formula2 <-  paste0("`", var, "` ~ `Treatment` + Sex + PND + `Treatment`:PND + Sex:PND + (1|`Animal ID`)")
    model1 <- glmer(as.formula(formula1), data = data, family = binomial(link = "logit"),
                    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))
    model2 <- glmer(as.formula(formula2), data = data, family = binomial(link = "logit"),
                    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))
    anova_result_1 = Anova(model1, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_1 = anova_result_1[,c(1,4)]
    names(anova_result_1) <- c("Effect", "P-Value (3-Way Model)")
    anova_result_2 = Anova(model2, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_2 = anova_result_2[,c(1,4)]
    names(anova_result_2) <- c("Effect", "P-Value (2-Way Model)")
    combined_anova_results <- create_combined_estimate(summary_full = anova_result_1, anova_result_2)
    merged_summary_list[[i]] <- combined_anova_results
  }

  # Create and store bar plot
  plot_summary <- data |> 
    select(Treatment, PND, Sex, !!sym(var)) |> 
    mutate(Treatment = as.factor(Treatment),
           Sex = as.factor(Sex),
           PND = as.factor(PND),
           !!sym(var) := ifelse(!!sym(var) == 1, 1, 0)) |> 
    filter(!is.na(!!sym(var))) |> 
    group_by(Treatment, Sex, PND) |> 
    summarise(
      Proportion_Yes = mean(!!sym(var)),
      n = n(),
      SE = sqrt((Proportion_Yes * (1 - Proportion_Yes)) / n),  # Standard Error for Proportions
      .groups = "drop"
    ) |> 
    mutate(
      Lower = pmax(Proportion_Yes - 2 * SE, 0),  # Ensure lower bound is not negative
      Upper = pmin(Proportion_Yes + 2 * SE, 1)   # Ensure upper bound does not exceed 1
    )

  # Create bar plot with 2SE error bars
  plot_list[[i]] <- ggplot(plot_summary, aes(x = Treatment, y = Proportion_Yes, fill = Sex, color = Sex)) +
    geom_bar(stat = "identity", position = position_dodge(0.8)) + 
    geom_errorbar(aes(ymin = Lower, ymax = Upper), color = "black",
                  position = position_dodge(0.8), width = 0.2) +  
    facet_wrap(~PND,labeller = label_both) +
    scale_fill_brewer(palette = "Set1") +  
    scale_color_brewer(palette = "Set1") +
    theme_minimal() +
    labs(title = paste0("Proportion of 'Yes' in ", var, " by Treatment & Sex"),
         x = "Treatment",
         y = paste0(var, " Proportion")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          panel.grid.major = element_line(color = "gray85"),
          panel.grid.minor = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(face = "bold"))
  
  post_hoc_h1_list[[i]] <- run_posthoc_test_cat_h1(data = data_fct_compete, var = var)
  post_hoc_h2_list[[i]] <- run_posthoc_test_cat_h2(data = data_fct_compete, var = var)
}

# Finalize the tables (updating the raw_pval to hypothesis level fdr)
fdr_values <- p.adjust(unlist(lapply(post_hoc_h1_list, function(df) if (!is.null(df)) df$P_Value_Raw)), method = "fdr") |> round(3)
fdr_adjusted <- ifelse(fdr_values < 0.05, paste0(fdr_values, "*"), as.character(fdr_values))
index <- 1
for (i in seq_along(post_hoc_h1_list)) {
  if (!is.null(post_hoc_h1_list[[i]])) {
    n <- nrow(post_hoc_h1_list[[i]])
    post_hoc_h1_list[[i]]$FDR <- fdr_adjusted[index:(index + n - 1)] 
    post_hoc_h1_list[[i]] <- post_hoc_h1_list[[i]] |>
      select(-P_Value_Raw) |> 
      flextable() |>  
      set_table_properties(layout = "autofit") 
    index <- index + n
  }
}

# Extract non-null P_Value_Raw values

fdr_values <- p.adjust(unlist(lapply(post_hoc_h2_list, function(df) if (!is.null(df)) df$P_Value_Raw)), method = "fdr") |> round(3)
fdr_adjusted <- ifelse(fdr_values < 0.05, paste0(fdr_values, "*"), as.character(fdr_values))
index <- 1
for (i in seq_along(post_hoc_h2_list)) {
  if (!is.null(post_hoc_h2_list[[i]])) {
    n <- nrow(post_hoc_h2_list[[i]])
    post_hoc_h2_list[[i]]$FDR <- fdr_adjusted[index:(index + n - 1)]
    post_hoc_h2_list[[i]] <- post_hoc_h2_list[[i]] %>%
      select(-P_Value_Raw) |>  
      flextable() |>  
      set_table_properties(layout = "autofit") 
    index <- index + n
  }
}
```


```{r}
###################### Run the model and create plot for survival variables
merged_forest_plot = list()
plot_list_km_curve = list()

for (i in vars_survival) {
  var <- vars[i]

  data_survival = data[,c("Animal ID","Litter ID","PND",var, "Treatment","Sex")] |>
    mutate(
      censor = ifelse(!!sym(var) == 30, 0, 1) |> as.character() |> as.numeric(),
      PND = factor(as.character(PND), levels = c("4", "8", "12")),
      Treatment = factor(Treatment,
                         levels = c("Adjuvant+Saline", "CRMP1+CRMP2",
                                    "CRMP1+GDA", "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE")),
      Sex = factor(Sex, levels = c("M", "F")),
      `Litter ID` = as.factor(`Litter ID`)
      ) |> 
    na.omit()

  # Run Cox model and extract tidy summary with exponentiation coefficients (hazard ratios)
  pnd_results <- list()
  pnd_km_plots <- list()
  for (pnd_val in c("4", "8", "12")) {
    data_subset <- data_survival |> filter(PND == pnd_val)

    if (nrow(data_subset) > 0) {  # Only fit if there's data
      formula_fit <- paste0("Surv(`", var, "`, censor) ~ Treatment + Sex + frailty(`Litter ID`)") |>
        as.formula()
      cox_fit <- coxph(formula_fit, data = data_subset)
      cox_tidy <- tidy(cox_fit, 
                       exponentiate = TRUE, 
                       conf.int = TRUE) |>
        filter(!str_detect(term, "Litter ID"))  # remove frailty term

      # Rename term nicely
      cox_tidy$term <- c("CRMP1+CRMP2 vs Adjuvant+Saline",
                         "CRMP1+GDA vs Adjuvant+Saline",
                         "LDHA+LDHB+CRMP1+STIP1 vs Adjuvant+Saline",
                         "STIP1+NSE vs Adjuvant+Saline",
                         "F vs M")

      # Add PND column
      cox_tidy <- cox_tidy |>
        transmute(
          PND = pnd_val,
          Term = term,
          HR = round(estimate, 3),
          `Std. Error` = round(std.error, 3),
          `P-value` = ifelse(p.value < 0.001, "<0.001", round(p.value, 3)),
          `95% CI Lower` = round(conf.low, 3),
          `95% CI Upper` = round(conf.high, 3)
        )

      pnd_results[[pnd_val]] <- cox_tidy
      
      # Step 4. Fit KM curve
      surv_object <- Surv(data_subset[[var]], data_subset$censor)
      km_fit <- survfit(surv_object ~ Treatment, data = data_subset)
      pnd_km_plots[[pnd_val]] <- ggsurvplot(
        fit = km_fit,
        data = data_subset |> as.data.frame(),
        conf.int = TRUE,
        pval = TRUE,
        risk.table = TRUE,
        xlab = "Time (Seconds)",
        ylab = "Turn-Over Probability",
        title = paste0("PND ", pnd_val),
        ggtheme = theme_minimal()
      )
    }
  }

  library(patchwork)
  
  # combine PND 4/8/12 results into one table
  plot_list_km_curve[[i]] <- pnd_km_plots[["4"]]$plot + 
                              pnd_km_plots[["8"]]$plot + 
                              pnd_km_plots[["12"]]$plot +
                              plot_layout(ncol = 1) +
                              plot_layout(guides = "collect") & 
                              theme(
                                legend.position = "right",
                                legend.direction = "vertical"
                              )

  merged_summary_list[[i]] <- bind_rows(pnd_results)

  # Draw forest plot ("Hazard Ratios with 95% Confidence Intervals")
  cox_plot_data <- merged_summary_list[[i]] |>
    mutate(lower = `95% CI Lower`,
           upper = `95% CI Upper`,
           PND_label = paste0("PND = ", PND))|>
    filter(Term != "F vs M")
  merged_forest_plot[[i]] = ggplot(cox_plot_data, aes(x = HR, y = reorder(Term, HR))) +
  geom_point(size = 3, color = "darkblue") +
  geom_errorbarh(aes(xmin = `95% CI Lower`, xmax = `95% CI Upper`), height = 0.2, color = "darkblue") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  scale_x_log10() +
  labs(
    x = "Hazard Ratio (log scale)",
    y = ""
  ) +
  facet_wrap(~ PND_label, ncol = 1, scales = "free_y") +  # <--- facet by PND vertically
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    strip.text = element_text(face = "bold", size = 14)
  )

  
  # Create and store bar plot
  plot_summary <- data_survival |>
    group_by(Treatment, Sex, PND) |> 
    summarise(
      Proportion_Yes = mean(censor),
      n = n(),
      SE = sqrt((Proportion_Yes * (1 - Proportion_Yes)) / n),  # Standard Error for Proportions
      .groups = "drop"
    ) |> 
    mutate(
      Lower = pmax(Proportion_Yes - 2 * SE, 0),  # Ensure lower bound is not negative
      Upper = pmin(Proportion_Yes + 2 * SE, 1)   # Ensure upper bound does not exceed 1
    )

  # Create bar plot with 2SE error bars
  plot_list[[i]] <- ggplot(plot_summary, aes(x = Treatment, y = Proportion_Yes, fill = Sex, color = Sex)) +
    geom_bar(stat = "identity", position = position_dodge(0.8)) + 
    geom_errorbar(aes(ymin = Lower, ymax = Upper), color = "black",
                  position = position_dodge(0.8), width = 0.2) +  
    facet_wrap(~PND,labeller = label_both) +
    scale_fill_brewer(palette = "Set1") +  
    scale_color_brewer(palette = "Set1") +
    theme_minimal() +
    labs(title = paste0("Proportion of Animals First Record Achieves 30s Mark in ", var, " by Treatment & Sex"),
         x = "Treatment",
         y = paste0(var, " Proportion")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          panel.grid.major = element_line(color = "gray85"),
          panel.grid.minor = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(face = "bold"))
  
  #post_hoc_h1_list[[i]] <- run_posthoc_test_cat_h1(data = data_fct_compete, var = var)
  #post_hoc_h2_list[[i]] <- run_posthoc_test_cat_h2(data = data_fct_compete, var = var)
}
```
# Background

In MAR 4.0 project, this section we will conduct analysis on the milestone data and report analysis results.

We are going to run the analysis using R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with total observations of `r milestones |> dplyr::pull(dplyr::any_of('Animal ID')) |> length()`, and total number of animals `r milestones |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. 

The interested variables at milestones data are: `r paste(vars, collapse = ", ")`.

The `Startle` and `Forelimb Grasp` variables were converted to binary (yes/no of performance) for analysis.

We also dichotomized three variables (`Circle Traverse`, `Negative Geotaxis`, and `Cliff Avoidance`) as if the animal achieved the milestone within 30 seconds. Those three variables are evaluated in both numeric and binary variable version.

# Summary statistics

The all interested variables were summarized by Treatment group overall, stratified by Time and Sex.

```{r summary-tbl, results='asis'}
for (i in seq_along(vars)) {
  cat("\n\n### ---", vars[i], "---\n\n")
  summary_list <- summary_tbls(data, vars[i])
  
  purrr::walk2(summary_list, vars[i], function(tbl, var_name) {
    tbl |>
      gtsummary::as_flex_table() |>
      flextable::width(width = 1.15) |>
      flextable::width(j = 1, width = 1.16) |>
      flextable::flextable_to_rmd()
    
    tab_n <<- tab_n + 1  
    cat(paste0("\nTable ", tab_n, 
               ": Summary of ", 
               var_name, 
               " by Treatment Groups with Time Stratification (4, 8, and 12th week).\n\n"))
  })
}
```

# Model Results

To evaluate differences in mean milestone measures between treatments over time, we conducted repeated measures ANOVA (or logistic regression for a cross-sectional categorical milestone measure), incorporating animals as a random effect.

We fitted 2 models as follows in order: 

- Fit the 3-way interaction ANOVA model: Begin with a 3-way interaction model with the main effects of time, treatment (trt), and sex, their 2-way interactions and the three-way interaction of all three variables. This model assumes sex as an effect modifier. 

- If the 3-way interaction term in ANOVA is not statistically significant (i.e, if sex is not an effect modifier), we drop the 3-way interaction term and fit the 2-way interaction model that assumes sex as a confounding variable. 

Additionally, for binary outcomes (Startle), we use ordinary logistic regression to evaluate the effect on the 12th week, this is because no animal has Startle at the 4th and 8th week (this is non-repeated measure). Forelimb Grasp is no evaluated because no animal has Startle at any time point.

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report’s delivery format.

## ANOVA Table

Type III ANOVA table displays the anova p-values for each variable in each model, and below is the explanation of each continuous outcome model:

- Three interaction model shows the p-value on three way interaction: y = time(PND) * trt * sex

- Two-way interaction model shows the p-value without the three way interaction: y = time(PND) * trt + Sex + time(PND) : Sex


**OR** for binary outcome models:

- Three interaction model shows the p-value on three way interaction: y = trt * sex

- Two-way interaction model shows the p-value without the three way interaction: y = trt + sex


The models also account for within-subjects variation attributed to animals (i.e, inter-animal variability within the same treatment group), and all p-values are calculated through Type-III sum of squares of residuals, using the R “rstatix” package.

Additionally, we use boxplots for intuitive visualizations for cross-sectional pairwise group comparisons at each time point tested by t-test with a combination of Bonferroni adjustment method (more * means more significance of the pairwise t-test).

- *Some Righting Reflex are excluded (n = 4): At PND 4, 9_42_2 has 55.8, at PND 8, 9_42_7 has 27.4, and at PND 12, 9_15_8 and 9_45_1 have 6.06 and 5.95.*

- *The summary tables for "Righting Reflex", "Circle Traverse", "Negative Geotaxis" and "Cliff Avoidance" are not displayed because they are not normal distributed, providing the results will be misleading.*

## Post-hoc Tests

Post-hoc tests are multiple pairwise comparisons between groups to identify which groups are different. Based on our hypothesis, we designed two set of post-hoc tests for each of variable. The first set is all comparison between the control group and each treatment group, and the second set is across all treatment group. The p-values are adjusted by Bonferroni method on each set of comparison and on each time period.

## Survival Analysis

For numeric outcomes with a maximum measurement cap of 30 seconds, we conducted Cox proportional hazards regression analyses. In each model, animals that did not successfully complete the task within 30 seconds were treated as censored observations.

The analyses were stratified by postnatal day (PND) to account for developmental differences across time points.

Additionally, we visualized the estimated hazard ratios along with their 95% confidence intervals across PND levels for each treatment group.

## Modeling Result

```{r anova-tb, results='asis'}
set_flextable_defaults(
  font.family = "Arial",  # Optional: Set default font
  font.size = 10,         # Optional: Set default font size
  theme_fun = theme_vanilla # Optional: Set default theme
)
for (i in c(vars_numeric,vars_binary,vars_survival)) {
  cat("\n\n### ---", vars[i], "---\n\n")
  # Create flextable summary table
  if (!(i%in%c(5))){
      merged_summary_list[[i]] |>
      as.data.frame(row.names = NULL) |>
      flextable::flextable() |>
      flextable::set_table_properties(layout = "autofit") |>
      flextable::flextable_to_rmd()
    tab_n = tab_n + 1
    if (i %in% vars_survival){
    cat(paste0("\nTable ",tab_n, 
               ": Cox Hazard Regression Results stratified by postnatal day (PND) on ", 
               vars[i] ,"\n\n"))
    } else{
    cat(paste0("\nTable ",tab_n, 
               ": Cross-sectional Comparison of ANOVA Results for Two Model Specifications of ", 
               vars[i] ,"\n\n"))
    }

  }
  
  print(plot_list[[i]])
  fig_n = fig_n + 1
  if(i %in% vars_numeric){
      cat(paste0("\nFigure ",fig_n, 
                 ": Boxplot showing the distribution of ", vars[i] ,
                 " by Treatment and Sex across Postnatal Days (PND) with error bars depict ±2 standard errors (2SE)\n\n"))
    } else{
      cat(paste0("\nFigure ",fig_n, 
                 ": Barplot showing the distribution of ", vars[i] ,
                 " by Treatment and Sex across Postnatal Days (PND) with error bars depict ±2 standard errors (2SE)\n\n"))
    }

  if (i %in% c(6,7,8)){
    print(merged_forest_plot[[i]])
    fig_n = fig_n + 1
    cat(paste0("\nFigure ",fig_n, 
           ": Forest Plot shows Hazard Ratios with 95% Confidence Intervals of ", vars[i] ,
           " at each Treatment Group\n\n"))

    print(plot_list_km_curve[[i]])
    fig_n = fig_n + 1
    cat(paste0("\nFigure ",fig_n, 
           ": Kaplan-Meier curves with Log-rank Test result",
           "showing probabilities of turning over ",
           "within 30 seconds for each Treatment group ",
           "at postnatal days (PND) 4, 8, and 12 of **", 
           vars[i],
           "** \n\n"))
  }
  
  if (!(i%in%c(5,6,7,8))){
    post_hoc_h1_list[[i]]|>
      flextable::flextable_to_rmd()
    tab_n = tab_n + 1
    cat(paste0("\nTable ",tab_n, 
               ": Post-hoc test for comparing the control null group (Adjuvant+Saline) to treatments with a control 95% family-wise confidence level - ", 
               vars[i] ,"\n"))
  
    post_hoc_h2_list[[i]]|>
      flextable::flextable_to_rmd()
    tab_n = tab_n + 1
    cat(paste0("\nTable ",
               tab_n, ": Post-hoc test for comparing across treatments with a control 95% family-wise confidence level - ",
               vars[i] ,"\n\n"))
  }

  cat(paste0("\n\n- - - - - - - - - - - - - - - - - - - - - -\n\n"))
}



```



## Principle Component Analysis (PCA)

This PCA analysis using all raw scores at each time-point, separately. 

- Plot PC1 vs PC2, potentially include loading

- Examine importance scores for PC1, PC2, and potentially PC3

### PCA Result at Week 4

```{r PCA-4, results='asis'}
# Getting the data ready for PCA
data_pca <- data |>
  filter(PND == "4") |>    # PCA at 4th week
  select(all_of(vars[1:9]),Treatment)
non_constant_vars <- names(data_pca %>% select(-Treatment))[sapply(data_pca %>% select(-Treatment), function(col) {
  if (all(is.na(col))) return(FALSE)  
  var(col, na.rm = TRUE) > 0  
})]
data_pca <- data_pca |>
  select(all_of(c(non_constant_vars,"Treatment"))) |>
  na.omit()
data_pca_scaled <- data_pca |> 
  select(-Treatment) |> 
  scale()

# Extract PCA Importance Scores
pca_result <- prcomp(data_pca_scaled, center = TRUE, scale. = TRUE)
pca_summary <- summary(pca_result)
importance_scores <- as.data.frame(pca_summary$importance)
colnames(importance_scores) <- paste0("PC", seq_len(ncol(importance_scores)))
loadings <- as.data.frame(pca_result$rotation)  # Extract PC1 & PC2 loadings

# PCA Biplot
p <- ggbiplot(pca_result, 
              choices = c(1,2),
              ellipse = TRUE,
              var.scale = 1,
              obs.scale = 1,
              varname.size = 5,  
              groups = data_pca$Treatment, 
              repel = TRUE) +
  geom_point(size = 2, alpha = 0.8, aes(color = data_pca$Treatment)) +  # Color points
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    text = element_text(size = 16)  # Ensures readable variable labels
  ) +
  xlab("Principal Component 1 (PC1)") +
  ylab("Principal Component 2 (PC2)") +
  ylim(-5, 5)  # Adjust Y-axis range
print(p)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Biplot (PC1 vs PC2) for Milestone Data at PND = 4 \n\n"))

# Scree Plot to Show Variance Explained
fviz_eig(pca_result)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Scree plot for Milestone Data at PND = 4 \n\n"))

loadings |>
  round(4) |>
  as.data.frame() |>
  tibble::rownames_to_column(var = "Variable") |>
  flextable::flextable() |>
  width(width = 1.15) |>
  width(j = 1, width = 1.16) |>
  autofit()
tab_n = tab_n + 1
cat(paste0("\nTable ",tab_n, ": Factor Loading for PCA Analysis at PND = 4 \n\n"))
```




### PCA Result at Week 8

```{r PCA-8, results='asis'}
# Getting the data ready for PCA
data_pca <- data |>
  filter(PND == "8") |>    # PCA at 8th week
  select(all_of(vars[1:9]),Treatment)
non_constant_vars <- names(data_pca %>% select(-Treatment))[sapply(data_pca %>% select(-Treatment), function(col) {
  if (all(is.na(col))) return(FALSE)  
  var(col, na.rm = TRUE) > 0  
})]
data_pca <- data_pca |>
  select(all_of(c(non_constant_vars,"Treatment"))) |>
  na.omit()
data_pca_scaled <- data_pca |> 
  select(-Treatment) |> 
  scale()

# Extract PCA Importance Scores
pca_result <- prcomp(data_pca_scaled, center = TRUE, scale. = TRUE)
pca_summary <- summary(pca_result)
importance_scores <- as.data.frame(pca_summary$importance)
colnames(importance_scores) <- paste0("PC", seq_len(ncol(importance_scores)))
loadings <- as.data.frame(pca_result$rotation)  # Extract PC1 & PC2 loadings

# PCA Biplot
p <- ggbiplot(pca_result, 
              choices = c(1,2),
              ellipse = TRUE,
              var.scale = 1,
              obs.scale = 1,
              varname.size = 5,  
              groups = data_pca$Treatment, 
              repel = TRUE) +
  geom_point(size = 2, alpha = 0.8, aes(color = data_pca$Treatment)) +  # Color points
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    text = element_text(size = 16)  # Ensures readable variable labels
  ) +
  xlab("Principal Component 1 (PC1)") +
  ylab("Principal Component 2 (PC2)") +
  ylim(-5, 5)  # Adjust Y-axis range
print(p)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Biplot (PC1 vs PC2) for Milestone Data at PND = 8 \n\n"))

# Scree Plot to Show Variance Explained
fviz_eig(pca_result)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Scree plot for Milestone Data at PND = 8 \n\n"))

loadings |>
  round(4) |>
  as.data.frame() |>
  tibble::rownames_to_column(var = "Variable") |>
  flextable::flextable() |>
  width(width = 1.15) |>
  width(j = 1, width = 1.16) |>
  autofit()
tab_n = tab_n + 1
cat(paste0("\nTable ",tab_n, ": Factor Loading for PCA Analysis at PND = 8 \n\n"))
```


### PCA Result at Week 12

```{r PCA-12, results='asis'}
# Getting the data ready for PCA
data_pca <- data |>
  filter(PND == "12") |>    # PCA at 12th week
  select(all_of(vars[1:9]),Treatment)
non_constant_vars <- names(data_pca %>% select(-Treatment))[sapply(data_pca %>% select(-Treatment), function(col) {
  if (all(is.na(col))) return(FALSE)  
  var(col, na.rm = TRUE) > 0  
})]
data_pca <- data_pca |>
  select(all_of(c(non_constant_vars,"Treatment"))) |>
  na.omit()
data_pca_scaled <- data_pca |> 
  select(-Treatment) |> 
  scale()

# Extract PCA Importance Scores
pca_result <- prcomp(data_pca_scaled, center = TRUE, scale. = TRUE)
pca_summary <- summary(pca_result)
importance_scores <- as.data.frame(pca_summary$importance)
colnames(importance_scores) <- paste0("PC", seq_len(ncol(importance_scores)))
loadings <- as.data.frame(pca_result$rotation)  # Extract PC1 & PC2 loadings

# PCA Biplot
p <- ggbiplot(pca_result, 
              choices = c(1,2),
              ellipse = TRUE,
              var.scale = 1,
              obs.scale = 1,
              varname.size = 5,  
              groups = data_pca$Treatment, 
              repel = TRUE) +
  geom_point(size = 2, alpha = 0.8, aes(color = data_pca$Treatment)) +  # Color points
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    text = element_text(size = 16)  # Ensures readable variable labels
  ) +
  xlab("Principal Component 1 (PC1)") +
  ylab("Principal Component 2 (PC2)") +
  ylim(-5, 5)  # Adjust Y-axis range
print(p)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Biplot (PC1 vs PC2) for Milestone Data at PND = 12 \n\n"))

# Scree Plot to Show Variance Explained
fviz_eig(pca_result)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Scree plot for Milestone Data at PND = 12 \n\n"))

loadings |>
  round(4) |>
  as.data.frame() |>
  tibble::rownames_to_column(var = "Variable") |>
  flextable::flextable() |>
  width(width = 1.15) |>
  width(j = 1, width = 1.16) |>
  autofit()
tab_n = tab_n + 1
cat(paste0("\nTable ",tab_n, ": Factor Loading for PCA Analysis at PND = 12 \n\n"))
```


