---
title: "MAR 4.0 - Milestones"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2025-05-08"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, rstatix, ggpubr, ggbiplot, factoextra, survival, 
               survminer, ggplot2, stringr, mar4.0, patchwork)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
fig_n = 0
tab_n = 0
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glm(as.formula(formula_str), data = data, family = binomial(link = "logit")))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}

# Function to summarize and format model results
format_summary <- function(model_summary, coef_indices, column_name) {
  formatted <- apply(model_summary$coefficients[, coef_indices], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(summary_full, ...) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  
  for (i in 2:length(summaries)) {
    merged_data <- full_join(merged_data, summaries[[i]], by = "Effect")
  }
  
  # Convert numeric columns (p-values) to 4 decimal places (non-scientific)
  numeric_cols <- sapply(merged_data, is.numeric)
  merged_data[numeric_cols] <- lapply(merged_data[numeric_cols], function(x) {
    ifelse(is.na(x), "", as.character(format(round(x, 4), nsmall = 4, scientific = FALSE)))
  })
  
  return(merged_data)
}

# Function to return the gtsummary-table by input dataset and outcome
summary_tbls <- function(
    dataset, outcome
) {

  tbl_male <- dataset |>
    dplyr::filter(Sex == "M") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          ) 
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Males"
    )

  tbl_female <- dataset |>
    dplyr::filter(Sex == "F") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          )
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Females"
    )

  ans <- list(
    "males" = tbl_male,
    "females" = tbl_female
  )

  return(ans)
}
```

```{r}
###################### Run the model and create plot for numeric variables

# Ensure PND is a factor with reference level set to "4"
data = milestones |>
  mutate(`Animal ID` = ifelse(is.na(`Animal ID`),paste0(`Litter ID`,"_99"),`Animal ID`),
          Startle = ifelse(Startle=="N",0,1),
         `Forelimb Grasp` = ifelse(`Forelimb Grasp`=="N",0,1),
         PND = relevel(factor(PND), ref = "4"),
         `Circle Traverse 30s` = ifelse(`Circle Traverse`<30,1,0) |> as.factor(),
         `Negative Geotaxis 30s` = ifelse(`Negative Geotaxis`<30,1,0)|> as.factor(),
         `Cliff Avoidance 30s` = ifelse(`Cliff Avoidance`<30,1,0)|> as.factor(),
         `Log2 Righting Reflex` = log2(`Righting Reflex`))


# Variables to analyze, including Count
vars = names(data)[11:24]
vars_numeric = c(1:5,14)
vars_survival = c(6:8)
vars_binary = c(9,11,12,13)

# Create ANOVA results table
merged_summary_list <- list()
plot_list <- list()
post_hoc_h1_list <- list()
post_hoc_h2_list <- list()

# Main loop over variables
for (i in vars_numeric) {
  var <- vars[i]
  # Data preparation (ensure complete cases)
  data_fct_compete <- data |>
    dplyr::mutate(
      Treatment = factor(Treatment, levels = c(
        "Adjuvant+Saline", "CRMP1+CRMP2", "CRMP1+GDA",
        "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE"
      )),
      Sex = factor(Sex, levels = c("M", "F")),
      Litter = factor(`Litter ID`),
      Animal = factor(`Animal ID`)
    ) |>
    dplyr::filter(dplyr::if_all(c(Animal, Litter, Treatment, Sex, PND, !!sym(var)), ~ !is.na(.x)))
  
  # If working on the variable with 30s max recording, trimming those equal to 30s
  # Also check low counting PND and exclude those
  if (var %in% c("Circle Traverse","Negative Geotaxis","Cliff Avoidance")){
    data_fct_compete <- data_fct_compete |>
      filter(!is.na(!!sym(var)) & !!sym(var) < 30)
    
    exclude_pnd <- data_fct_compete |>
      group_by(PND, Treatment, Sex) |>
      summarise(n = n(), .groups = "drop") |>
      filter(n <= 1) |>
      pull(PND) |>
      unique()

    data_fct_compete <- data_fct_compete |>
      filter(!PND %in% exclude_pnd)
  }

  
  # Perform 3-way ANOVA (Treatment * Sex * PND)
  mix_aov_3way <- afex::aov_car(
  formula = as.formula(paste0("`", var, "` ~ Treatment * Sex * PND + Error(Animal/PND)")),
  data = data_fct_compete,
  type = 3
  )
  
  # Perform 2-way ANOVA (Treatment * PND)
  mix_aov_2way <- afex::aov_car(
  formula = as.formula(paste0("`", var, "` ~ Treatment * PND + Sex * PND + Error(Animal/PND)")),
  data = data_fct_compete,
  type = 3
  )
  # mix_aov_3way <- rstatix::anova_test(
  #   dv = !!sym(var), 
  #   wid = Animal, 
  #   between = c(Treatment, Sex), 
  #   within = PND, 
  #   data = data_fct_compete, #|> filter(PND != "12"), 
  #   type = 3
  # )
  # 
  # # Perform 2-way ANOVA (Treatment * PND)
  # mix_aov_2way <- rstatix::anova_test(
  #   dv = `Cliff Avoidance`,
  #   wid = Animal,
  #   between = c(Treatment),
  #   within = PND,
  #   covariate = Sex,
  #   data = data_fct_compete,
  #   type = 3
  # )

  # Extract p-values from ANOVA tests and merge them into one table
  if (!is.null(mix_aov_3way$ANOVA)) {
    p_values_3way <- mix_aov_3way$ANOVA[, c("Effect", "p.value")] |>
      dplyr::rename(`P-Value (3-Way Model)` = p.value)
    
    p_values_2way <- mix_aov_2way$ANOVA[, c("Effect", "p.value")] |>
      dplyr::rename(`P-Value (2-Way Model)` = p.value)
  } else {
    p_values_3way <- mix_aov_3way$anova_table |>
      as.data.frame() |>
      tibble::rownames_to_column(var = "Effect") |>
      dplyr::select(`Effect`, `Pr(>F)`) |>
      dplyr::rename(`P-Value (3-Way Model)` = `Pr(>F)`)
    
    p_values_2way <- mix_aov_2way$anova_table |>
      as.data.frame() |>
      tibble::rownames_to_column(var = "Effect") |>
      dplyr::select(`Effect`, `Pr(>F)`) |>
      dplyr::rename(`P-Value (2-Way Model)` = `Pr(>F)`)
  }
  combined_anova_results <- create_combined_estimate(p_values_3way, p_values_2way)
  merged_summary_list[[i]] <- combined_anova_results
  
  ######## Create Boxplots for numeric variable
  # Compute summary statistics (Mean & SE)
  data_summary <- data_fct_compete %>%
    group_by(Treatment, Sex, PND) %>%
    summarise(
      Mean = mean(.data[[var]], na.rm = TRUE),
      SE = sd(.data[[var]], na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )

  # Create the plot with SE
  plot_list[[i]] <- ggboxplot(data_fct_compete, x = "Treatment", y = var, color = "Sex", fill = "Sex", palette = "Set1") +
    geom_errorbar(data = data_summary, 
                  aes(x = Treatment, y = Mean, ymin = Mean - 2*SE, ymax = Mean + 2*SE, group = interaction(Treatment, Sex)),  
                  position = position_dodge(width = 0.8), 
                  width = 0.2, 
                  color = "black") +  
    geom_point(data = data_summary, 
               aes(x = Treatment, y = Mean, color = Sex, group = interaction(Treatment, Sex)), 
               position = position_dodge(width = 0.8), 
               size = 1, 
               shape = 21) +  
    facet_wrap(~PND, scales = "free_x", nrow = 1, labeller = label_both) +
    theme_minimal() +
    labs(title = paste0("Distribution of ", var, " by Treatment & Sex"),
         x = "Treatment",
         y = paste0(var, " Value")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          strip.text = element_text(size = 12, face = "bold"))

  ######## Running post-hoc test on continuous variable
  p_list_3 <- emmeans::contrast(emmeans::emmeans(mix_aov_3way, ~ Treatment|PND), 
                                method = "pairwise", 
                                adjust = "none") |>
    as.data.frame() |> 
    mutate(PND = gsub("\\D", "", PND))
  p_list_2 <- emmeans::contrast(emmeans::emmeans(mix_aov_2way, ~ Treatment|PND), 
                                method = "pairwise", 
                                adjust = "none") |>
    as.data.frame() |> 
    mutate(PND = gsub("\\D", "", PND))
    
  post_hoc_h1_list[[i]] <- run_posthoc_test_num_h1(data = data_fct_compete, 
                                                   var = var,
                                                   p_list_3 = p_list_3[,c("contrast", "p.value", "PND")],
                                                   p_list_2 = p_list_2[,c("contrast", "p.value", "PND")])
  post_hoc_h1_list[[i]] <- post_hoc_h1_list[[i]] |>
      flextable() |>
      set_table_properties(layout = "autofit")
  post_hoc_h2_list[[i]] <- run_posthoc_test_num_h2(data = data_fct_compete, 
                                                   var = var,
                                                   p_list_3[,c("contrast", "p.value", "PND")],
                                                   p_list_2[,c("contrast", "p.value", "PND")])
  post_hoc_h2_list[[i]] <- post_hoc_h2_list[[i]] |>
      flextable() |>
      set_table_properties(layout = "autofit")
  
  
  create_barplot(data_fct_compete, 
               variable = var, 
               contrast_p = p_list_3 |> filter(PND == 4))
  
  create_boxplot(data_fct_compete, 
               variable = var, 
               contrast_p = p_list_3 |> filter(PND == 4))
  create_meanplot(data_fct_compete, 
               variable = var, 
               contrast_p = p_list_3 |> filter(PND == 4))
}

```

```{r}
###################### Run the model and create plot for binary variables
merged_estimate_list = list()

for (i in vars_binary) {
  var <- vars[i]
  model_type = ifelse(i == 9, "glm","glmer")

  data_fct_compete = data |>
    dplyr::mutate(
      Treatment = factor(Treatment, levels = c(
        "Adjuvant+Saline", "CRMP1+CRMP2", "CRMP1+GDA",
        "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE"
      )),
      Sex = factor(Sex, levels = c("M", "F")),
      Litter = factor(`Litter ID`),
      Animal = factor(`Animal ID`)
    ) |>
    dplyr::filter(dplyr::if_all(c(Animal, Litter, Treatment, Sex, PND, !!sym(var)), ~ !is.na(.x)))
  
  if (model_type == "glm"){
    data_fct_compete = data_fct_compete |> filter(PND == 12)
    # Fit models
    formula1 <-  paste0("`", var, "` ~ `Treatment` + `Sex` + `Treatment`*`Sex` ")
    formula2 <-  paste0("`", var, "` ~ `Treatment` + `Sex` ")
    model1 <- glm(as.formula(formula1), data = data_fct_compete, family = binomial(link = "logit"))
    model2 <- glm(as.formula(formula2), data = data_fct_compete, family = binomial(link = "logit"))
    
    # Summarize and format results
    coef_indices <- if (model_type == "glm") c(1, 2, 4) else c(1, 2, 5)
    summary1 <- format_summary(summary(model1), coef_indices, "Two Vars Model") |> tibble::rownames_to_column()
    names(summary1) <- c("Effect", names(summary1)[2])
    summary2 <- format_summary(summary(model2), coef_indices, "Main Effect Model")|> tibble::rownames_to_column()
    names(summary2) <- c("Effect", names(summary2)[2])
    merged_estimate <- create_combined_estimate(summary_full = summary1, summary2)
    merged_estimate_list[[i]] <- merged_estimate
    
    # Perform ANOVA
    anova_result_1 = Anova(model1, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_1 = anova_result_1[,c(1,4)]
    names(anova_result_1) <- c("Effect", "Two Vars Model")
    anova_result_2 = Anova(model2, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_2 = anova_result_2[,c(1,4)]
    names(anova_result_2) <- c("Effect", "Main Effect Model")
    combined_anova_results <- create_combined_estimate(summary_full = anova_result_1, anova_result_2)
    merged_summary_list[[i]] <- combined_anova_results
    
    p_list_3 <- emmeans::contrast(emmeans::emmeans(model1, ~ Treatment), 
                                  method = "pairwise", 
                                  adjust = "none") |>
      as.data.frame() |>
      mutate(PND = 12)
    p_list_2 <- emmeans::contrast(emmeans::emmeans(model2, ~ Treatment), 
                                  method = "pairwise", 
                                  adjust = "none") |>
      as.data.frame() |>
      mutate(PND = 12)
  } else {
    formula1 <-  paste0("`", var, "` ~ `Treatment` * Sex * `PND` + (1|`Animal ID`)")
    formula2 <-  paste0("`", var, "` ~ `Treatment` + Sex + PND + `Treatment`:PND + Sex:PND + (1|`Animal ID`)")
    model1 <- glmer(as.formula(formula1), data = data_fct_compete, family = binomial(link = "logit"),
                    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))
    model2 <- glmer(as.formula(formula2), data = data_fct_compete, family = binomial(link = "logit"),
                    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)))
    anova_result_1 = Anova(model1, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_1 = anova_result_1[,c(1,4)]
    names(anova_result_1) <- c("Effect", "P-Value (3-Way Model)")
    anova_result_2 = Anova(model2, type = "III") |> as.data.frame() |> tibble::rownames_to_column("Effect")
    anova_result_2 = anova_result_2[,c(1,4)]
    names(anova_result_2) <- c("Effect", "P-Value (2-Way Model)")
    combined_anova_results <- create_combined_estimate(summary_full = anova_result_1, anova_result_2)
    merged_summary_list[[i]] <- combined_anova_results
    
    p_list_3 <- emmeans::contrast(emmeans::emmeans(model1, ~ Treatment|PND), 
                                  method = "pairwise", 
                                  adjust = "none") |>
      as.data.frame() |> 
      mutate(PND = gsub("\\D", "", PND))
    p_list_2 <- emmeans::contrast(emmeans::emmeans(model2, ~ Treatment|PND), 
                                  method = "pairwise", 
                                  adjust = "none") |>
      as.data.frame() |> 
      mutate(PND = gsub("\\D", "", PND))
  }
  
  # Create and store bar plot
  plot_summary <- data_fct_compete |> 
    select(Treatment, PND, Sex, !!sym(var)) |> 
    mutate(Treatment = as.factor(Treatment),
           Sex = as.factor(Sex),
           PND = as.factor(PND),
           !!sym(var) := ifelse(!!sym(var) == 1, 1, 0)) |> 
    filter(!is.na(!!sym(var))) |> 
    group_by(Treatment, Sex, PND) |> 
    summarise(
      Proportion_Yes = mean(!!sym(var)),
      n = n(),
      SE = sqrt((Proportion_Yes * (1 - Proportion_Yes)) / n),  # Standard Error for Proportions
      .groups = "drop"
    ) |> 
    mutate(
      Lower = pmax(Proportion_Yes - 2 * SE, 0),  # Ensure lower bound is not negative
      Upper = pmin(Proportion_Yes + 2 * SE, 1)   # Ensure upper bound does not exceed 1
    )

  # Create bar plot with 2SE error bars
  plot_list[[i]] <- ggplot(plot_summary, aes(x = Treatment, y = Proportion_Yes, fill = Sex, color = Sex)) +
    geom_bar(stat = "identity", position = position_dodge(0.8)) + 
    geom_errorbar(aes(ymin = Lower, ymax = Upper), color = "black",
                  position = position_dodge(0.8), width = 0.2) +  
    facet_wrap(~PND,labeller = label_both) +
    scale_fill_brewer(palette = "Set1") +  
    scale_color_brewer(palette = "Set1") +
    theme_minimal() +
    labs(title = paste0("Proportion of 'Yes' in ", var, " by Treatment & Sex"),
         x = "Treatment",
         y = paste0(var, " Proportion")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          panel.grid.major = element_line(color = "gray85"),
          panel.grid.minor = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(face = "bold"))

  
  post_hoc_h1_list[[i]] <- run_posthoc_test_cat_h1(data = data_fct_compete, 
                                                   var = var,
                                                   p_list_3 = p_list_3[,c("contrast", "p.value", "PND")],
                                                   p_list_2 = p_list_2[,c("contrast", "p.value", "PND")])
  post_hoc_h1_list[[i]] <- post_hoc_h1_list[[i]] |>
      flextable() |>
      set_table_properties(layout = "autofit")
  post_hoc_h2_list[[i]] <- run_posthoc_test_cat_h2(data = data_fct_compete,
                                                   var = var,
                                                   p_list_3 = p_list_3[,c("contrast", "p.value", "PND")],
                                                   p_list_2 = p_list_2[,c("contrast", "p.value", "PND")])
  post_hoc_h2_list[[i]] <- post_hoc_h2_list[[i]] |>
      flextable() |>
      set_table_properties(layout = "autofit")
}
```


```{r}
###################### Run the model and create plot for survival variables
merged_forest_plot = list()
plot_list_km_curve = list()

for (i in vars_survival) {
  var <- vars[i]

  data_survival = data[,c("Animal ID","Litter ID","PND", var, "Treatment","Sex")] |>
    mutate(
      censor = ifelse(!!sym(var) >= 30, 0, 1) |> as.character() |> as.numeric(),
      PND = factor(as.character(PND), levels = c("4", "8", "12")),
      Treatment = factor(Treatment,
                         levels = c("Adjuvant+Saline", "CRMP1+CRMP2",
                                    "CRMP1+GDA", "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE")),
      Sex = factor(Sex, levels = c("M", "F")),
      `Litter ID` = as.factor(`Litter ID`)
      ) |> 
    na.omit()

  # Run Cox model and extract tidy summary with exponentiation coefficients (hazard ratios)
  pnd_results <- list()
  pnd_km_plots <- list()
  for (pnd_val in c("4", "8", "12")) {
    data_subset <- data_survival |> 
      filter(PND == pnd_val) |> 
      mutate(
        Treatment = forcats::fct_recode(
          Treatment,
          "4+Drug+Combo" = "LDHA+LDHB+CRMP1+STIP1"
          )
        )
    if (nrow(data_subset) > 0) {  # Only fit if there's data
      formula_fit <- paste0("Surv(`", var, "`, censor) ~ Treatment + Sex + frailty(`Litter ID`)") |>
        as.formula()
      cox_fit <- coxph(formula_fit, data = data_subset)
      cox_tidy <- tidy(cox_fit, 
                       exponentiate = TRUE, 
                       conf.int = TRUE) |>
        filter(!str_detect(term, "Litter ID"))  # remove frailty term

      # Rename term nicely
      cox_tidy$term <- c("CRMP1+CRMP2 vs Adjuvant+Saline",
                         "CRMP1+GDA vs Adjuvant+Saline",
                         "LDHA+LDHB+CRMP1+STIP1 vs Adjuvant+Saline",
                         "STIP1+NSE vs Adjuvant+Saline",
                         "F vs M")

      # Add PND column
      cox_tidy <- cox_tidy |>
        transmute(
          PND = pnd_val,
          Term = term,
          HR = round(estimate, 3),
          `Std. Error` = round(std.error, 3),
          `P-value` = ifelse(p.value < 0.001, "<0.001", round(p.value, 3)),
          `95% CI Lower` = round(conf.low, 3),
          `95% CI Upper` = round(conf.high, 3)
        )

      pnd_results[[pnd_val]] <- cox_tidy
      
      # Fit KM curve
      surv_object <- Surv(data_subset[[var]], data_subset$censor)
      km_fit <- survfit(surv_object ~ Treatment, data = data_subset)
      pnd_km_plots[[pnd_val]] <- ggsurvplot(
        fit = km_fit,
        data = data_subset |> as.data.frame(),
        conf.int = TRUE,
        pval = TRUE,
        risk.table = TRUE,
        xlab = "Time (Seconds)",
        ylab = "Turn-Over Probability",
        title = paste0("PND ", pnd_val),
        ggtheme = theme_minimal()
      )
    }
  }
  
  # combine PND 4/8/12 results into one table
  plot_list_km_curve[[i]] <- pnd_km_plots[["4"]]$plot + 
                              pnd_km_plots[["8"]]$plot + 
                              pnd_km_plots[["12"]]$plot +
                              plot_layout(ncol = 1) +
                              plot_layout(guides = "collect") & 
                              theme(
                                legend.position = "right",
                                legend.direction = "vertical"
                              )

  merged_summary_list[[i]] <- bind_rows(pnd_results)

  # Draw forest plot ("Hazard Ratios with 95% Confidence Intervals")
  cox_plot_data <- merged_summary_list[[i]] |>
    mutate(lower = `95% CI Lower`,
           upper = `95% CI Upper`,
           PND_label = factor(
             paste0(PND, "th Week"),
             levels = paste0(c(4, 8, 12), "th Week"
                             ))
           )|>
    filter(Term != "F vs M")
  merged_forest_plot[[i]] = ggplot(cox_plot_data, aes(x = HR, y = reorder(Term, HR))) +
  geom_point(size = 3, color = "darkblue") +
  geom_errorbarh(aes(xmin = `95% CI Lower`, xmax = `95% CI Upper`), height = 0.2, color = "darkblue") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  scale_x_log10() +
  labs(
    x = "Hazard Ratio (log scale)",
    y = ""
  ) +
  facet_wrap(~ PND_label, ncol = 1, scales = "free_y") +  # <--- facet by PND vertically
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    strip.text = element_text(face = "bold", size = 14)
  )

  
  # Create and store bar plot
  plot_summary <- data_survival |>
    group_by(Treatment, Sex, PND) |> 
    summarise(
      Proportion_Yes = mean(censor),
      n = n(),
      SE = sqrt((Proportion_Yes * (1 - Proportion_Yes)) / n),  # Standard Error for Proportions
      .groups = "drop"
    ) |> 
    mutate(
      Lower = pmax(Proportion_Yes - 2 * SE, 0),  # Ensure lower bound is not negative
      Upper = pmin(Proportion_Yes + 2 * SE, 1)   # Ensure upper bound does not exceed 1
    )

  # Create bar plot with 2SE error bars
  plot_list[[i]] <- ggplot(plot_summary, aes(x = Treatment, y = Proportion_Yes, fill = Sex, color = Sex)) +
    geom_bar(stat = "identity", position = position_dodge(0.8)) + 
    geom_errorbar(aes(ymin = Lower, ymax = Upper), color = "black",
                  position = position_dodge(0.8), width = 0.2) +  
    facet_wrap(~PND,labeller = label_both) +
    scale_fill_brewer(palette = "Set1") +  
    scale_color_brewer(palette = "Set1") +
    theme_minimal() +
    labs(title = paste0("Proportion of Animals First Record Achieves 30s Mark in ", var, " by Treatment & Sex"),
         x = "Treatment",
         y = paste0(var, " Proportion")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          panel.grid.major = element_line(color = "gray85"),
          panel.grid.minor = element_blank(),
          strip.background = element_blank(), 
          strip.text = element_text(face = "bold"))
  
  #post_hoc_h1_list[[i]] <- run_posthoc_test_cat_h1(data = data_fct_compete, var = var)
  #post_hoc_h2_list[[i]] <- run_posthoc_test_cat_h2(data = data_fct_compete, var = var)
}
```
# Background

In MAR 4.0 project, this section we will conduct analysis on the milestone data and report analysis results.

We are going to run the analysis using R-package structure within github. R 4.4.3 is used for the analysis, and we are wrapping all the materials into a R package.

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with total observations of `r milestones |> dplyr::pull(dplyr::any_of('Animal ID')) |> length()`, and total number of animals `r milestones |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. 

The interested variables at milestones data are: `r paste(vars, collapse = ", ")`.

The `Startle` and `Forelimb Grasp` variables were converted to binary (yes/no of performance) for analysis.

We also dichotomized three variables (`Circle Traverse`, `Negative Geotaxis`, and `Cliff Avoidance`) as if the animal achieved the milestone within 30 seconds. Those three variables are evaluated in both numeric and binary variable version.

# Summary statistics

The all interested variables were summarized by Treatment group overall, stratified by Time and Sex.

```{r summary-tbl, results='asis'}
for (i in seq_along(vars)) {
  cat("\n\n### ---", vars[i], "---\n\n")
  summary_list <- summary_tbls(data, vars[i])
  
  purrr::walk2(summary_list, vars[i], function(tbl, var_name) {
    tbl |>
      gtsummary::as_flex_table() |>
      flextable::width(width = 1.15) |>
      flextable::width(j = 1, width = 1.16) |>
      flextable::flextable_to_rmd()
    
    tab_n <<- tab_n + 1  
    cat(paste0("\nTable ", tab_n, 
               ": Summary of ", 
               var_name, 
               " by Treatment Groups with Time Stratification (4, 8, and 12th week).\n\n"))
  })
}
```

# Model Results

To evaluate differences in mean milestone measures between treatments over time, we conducted repeated measures ANOVA (or logistic regression for a cross-sectional categorical milestone measure), incorporating animals as a random effect.

We fitted 2 models as follows in order: 

- Fit the 3-way interaction ANOVA model: Begin with a 3-way interaction model with the main effects of time, treatment (trt), and sex, their 2-way interactions and the three-way interaction of all three variables. This model assumes sex as an effect modifier. 

- If the 3-way interaction term in ANOVA is not statistically significant (i.e, if sex is not an effect modifier), we drop the 3-way interaction term and fit the 2-way interaction model that assumes sex as a confounding variable. 

Additionally, for binary outcomes (Startle), we use ordinary logistic regression to evaluate the effect on the 12th week, this is because no animal has Startle at the 4th and 8th week (this is non-repeated measure). Forelimb Grasp is no evaluated because no animal has Startle at any time point.

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report’s delivery format.

## ANOVA Table

Type III ANOVA table displays the anova p-values for each variable in each model, and below is the explanation of each continuous outcome model:

- Three interaction model shows the p-value on three way interaction: y = time(PND) * trt * sex

- Two-way interaction model shows the p-value without the three way interaction: y = time(PND) * trt + Sex + time(PND) : Sex


**OR** for binary outcome models:

- Three interaction model shows the p-value on three way interaction: y = trt * sex

- Two-way interaction model shows the p-value without the three way interaction: y = trt + sex


The models also account for within-subjects variation attributed to animals (i.e, inter-animal variability within the same treatment group), and all p-values are calculated through Type-III sum of squares of residuals, using the R “rstatix” package.

Additionally, we use boxplots for intuitive visualizations for cross-sectional pairwise group comparisons at each time point tested by t-test with a combination of Bonferroni adjustment method (more * means more significance of the pairwise t-test).

- *Some Righting Reflex are excluded (n = 4): At PND 4, 9_42_2 has 55.8, at PND 8, 9_42_7 has 27.4, and at PND 12, 9_15_8 and 9_45_1 have 6.06 and 5.95.*

- *The summary tables for "Righting Reflex", "Circle Traverse", "Negative Geotaxis" and "Cliff Avoidance" are not displayed because they are not normal distributed, providing the results will be misleading.*

## Post-hoc Tests

Post-hoc tests are multiple pairwise comparisons between groups to identify which groups are different. Based on our hypothesis, we designed two set of post-hoc tests for each of variable. The first set is all comparison between the control group and each treatment group, and the second set is across all treatment group. The p-values are adjusted by Bonferroni method on each set of comparison and on each time period.

## Survival Analysis

For numeric outcomes with a maximum measurement cap of 30 seconds, we conducted Cox proportional hazards regression analyses. In each model, animals that did not successfully complete the task within 30 seconds were treated as censored observations.

The analyses were stratified by postnatal day (PND) to account for developmental differences across time points.

Additionally, we visualized the estimated hazard ratios along with their 95% confidence intervals across PND levels for each treatment group.

There is one animal with `Negative Geotaxis` 30.22 seconds, we will just treated it as a censored data point as other 30 seconds.

## Post Hoc Analysis and P-Value Reporting

To evaluate pairwise differences between treatment groups across postnatal day (PND) levels, we performed post hoc tests tailored to the data type. The resulting tables report p-values from both the interaction model (inter) and the additive model (add). For each model, both raw and adjusted p-values were provided, with adjustment performed using the Hochberg method. The p-value adjustment was applied separately within each time period (PND = 4, 8, 12) and within each hypothesis type (primary or secondary).

For categorical outcomes, the proportion difference between groups was computed along with its standard error using the pooled standard error formula for two proportions. A 95% confidence interval for the difference was calculated using the normal approximation.

A custom helper function was used to format and annotate p-values: values less than 0.05 were marked with an asterisk (*), and all values were rounded to three decimal places to enhance readability.

**Note:** Post hoc analysis p-values are derived from the `emmeans` package, which calculates *estimated marginal means* (least-squares means) based on the fitted model. These p-values reflect pairwise comparisons between adjusted group means, accounting for other covariates or factors included in the model. Therefore, they may differ from the p-values reported in the main model summary (e.g., from ANOVA tables or survival model estimates), which test overall effects (e.g., for the entire Treatment factor) rather than specific comparisons. A non-significant main effect does not rule out significant pairwise comparisons between individual groups, and vice versa. Post hoc p-values should be interpreted accordingly, in the context of the specific contrasts being tested.

## Modeling Result


How to Interpret This ANOVA P-Value Table
This table presents p-values for the main effects and interaction terms from two factorial ANOVA models: a 3-way model, which includes all main effects and interactions up to the 3-way level (e.g., Treatment × Sex × Time), and a 2-way model, which includes only main effects and 2-way interactions. Each row corresponds to a specific effect or interaction (e.g., Treatment, Treatment × Time), and the p-values indicate whether that term has a statistically significant influence on the outcome variable.

Main effects (e.g., Treatment, Sex, Time) test whether each factor has an overall effect on the outcome, regardless of the levels of other factors.

Interaction terms test whether the effect of one factor changes depending on the level of another. For example, a significant Treatment × Time interaction suggests the treatment effect differs across time points.

3-way interactions test whether a 2-way interaction (e.g., Treatment × Time) itself varies across the third factor (e.g., Sex). A non-significant 3-way interaction suggests the simpler 2-way model may be sufficient.

### EXAMPLE - How to interpret the ANOVA table

The main effect of Treatment has a p-value of 0.001 in both the 3-way and 2-way models, indicating a highly statistically significant effect. This means that, regardless of Sex or PND (Postnatal Day), the outcome variable differs significantly across the treatment groups. In other words, the choice of treatment alone has a meaningful impact on the outcome, and this finding is robust across both models. 
### EXAMPLE - How to interpret the Posthoc table

Post hoc pairwise comparisons were conducted to examine differences between treatment groups at each postnatal day (PND) using chi-squared or Fisher's exact tests, depending on cell sizes. The table reports the proportion differences, standard errors, and 95% confidence intervals for each comparison. For example, at PND 4, the comparison between CRMP1+CRMP2 and the control group (Adjuvant+Saline) yielded a proportion difference of 0.215, with a 95% confidence interval of [0.048, 0.382] and an adjusted p-value of 0.028*, indicating a statistically significant increase in the response for the treatment group. P-values were adjusted using the Bonferroni method to control for multiple comparisons within each time point and hypothesis set. Asterisks next to p-values denote statistical significance at the 0.05 level.

### EXAMPLE - How to interpret the survival table

At postnatal day 12, animals in the LDHA+LDHB+CRMP1+STIP1 group had a hazard ratio of 0.569 compared to the Adjuvant+Saline group. This indicates a 43.1% reduction in the hazard of the event (e.g., avoidance failure, mortality, or whichever survival endpoint you're using), suggesting a potential protective effect of the treatment. The 95% confidence interval ranges from 0.325 to 0.996, and the p-value is 0.048, which is statistically significant at the 0.05 level. This suggests the observed difference is unlikely due to chance.

```{r anova-tb, results='asis'}
set_flextable_defaults(
  font.family = "Arial",  # Optional: Set default font
  font.size = 10,         # Optional: Set default font size
  theme_fun = theme_vanilla # Optional: Set default theme
)
for (i in c(vars_numeric,vars_binary,vars_survival)) {
  cat("\n\n### ---", vars[i], "---\n\n")
  # Create flextable summary table
  if (!(i%in%c(5))){
      merged_summary_list[[i]] |>
      as.data.frame(row.names = NULL) |>
      flextable::flextable() |>
      flextable::set_table_properties(layout = "autofit") |>
      flextable::flextable_to_rmd()
    tab_n = tab_n + 1
    if (i %in% vars_survival){
    cat(paste0("\nTable ",tab_n, 
               ": Cox Hazard Regression Results stratified by postnatal day (PND) on ", 
               vars[i] ,"\n\n"))
    } else{
    cat(paste0("\nTable ",tab_n, 
               ": Cross-sectional Comparison of ANOVA Results for Two Model Specifications of ", 
               vars[i] ,"\n\n"))
    }

  }
  
  print(plot_list[[i]])
  fig_n = fig_n + 1
  if(i %in% vars_numeric){
      cat(paste0("\nFigure ",fig_n, 
                 ": Boxplot showing the distribution of ", vars[i] ,
                 " by Treatment and Sex across Postnatal Days (PND) with error bars depict ±2 standard errors (2SE)\n\n"))
    } else{
      cat(paste0("\nFigure ",fig_n, 
                 ": Barplot showing the distribution of ", vars[i] ,
                 " by Treatment and Sex across Postnatal Days (PND) with error bars depict ±2 standard errors (2SE)\n\n"))
    }

  if (i %in% c(6,7,8)){
    print(merged_forest_plot[[i]])
    fig_n = fig_n + 1
    cat(paste0("\nFigure ",fig_n, 
           ": Forest Plot shows Hazard Ratios with 95% Confidence Intervals of ", vars[i] ,
           " at each Treatment Group\n\n"))

    print(plot_list_km_curve[[i]])
    fig_n = fig_n + 1
    cat(paste0("\nFigure ",fig_n, 
           ": Kaplan-Meier curves with Log-rank Test result",
           "showing probabilities of turning over ",
           "within 30 seconds for each Treatment group ",
           "at postnatal days (PND) 4, 8, and 12 of **", 
           vars[i],
           "** \n\n"))
  }
  
  if (!(i%in%c(5,6,7,8))){
    post_hoc_h1_list[[i]]|>
      flextable::flextable_to_rmd()
    tab_n = tab_n + 1
    cat(paste0("\nTable ",tab_n, 
               ": Post-hoc test for comparing the control null group (Adjuvant+Saline) to treatments with a control 95% family-wise confidence level - ", 
               vars[i] ,"\n"))
  
    post_hoc_h2_list[[i]]|>
      flextable::flextable_to_rmd()
    tab_n = tab_n + 1
    cat(paste0("\nTable ",
               tab_n, ": Post-hoc test for comparing across treatments with a control 95% family-wise confidence level - ",
               vars[i] ,"\n\n"))
  }

  cat(paste0("\n\n- - - - - - - - - - - - - - - - - - - - - -\n\n"))
}



```



## Principle Component Analysis (PCA)

This PCA analysis using all raw scores at each time-point, separately. 

- Plot PC1 vs PC2, potentially include loading

- Examine importance scores for PC1, PC2, and potentially PC3

### PCA Result at Week 4

```{r PCA-4, results='asis'}
# Getting the data ready for PCA
data_pca <- data |>
  filter(PND == "4") |>    # PCA at 4th week
  select(all_of(vars[1:9]),Treatment)
non_constant_vars <- names(data_pca %>% select(-Treatment))[sapply(data_pca %>% select(-Treatment), function(col) {
  if (all(is.na(col))) return(FALSE)  
  var(col, na.rm = TRUE) > 0  
})]
data_pca <- data_pca |>
  select(all_of(c(non_constant_vars,"Treatment"))) |>
  na.omit()
data_pca_scaled <- data_pca |> 
  select(-Treatment) |> 
  scale()

# Extract PCA Importance Scores
pca_result <- prcomp(data_pca_scaled, center = TRUE, scale. = TRUE)
pca_summary <- summary(pca_result)
importance_scores <- as.data.frame(pca_summary$importance)
colnames(importance_scores) <- paste0("PC", seq_len(ncol(importance_scores)))
loadings <- as.data.frame(pca_result$rotation)  # Extract PC1 & PC2 loadings

# PCA Biplot
p <- ggbiplot(pca_result, 
              choices = c(1,2),
              ellipse = TRUE,
              var.scale = 1,
              obs.scale = 1,
              varname.size = 5,  
              groups = data_pca$Treatment, 
              repel = TRUE) +
  geom_point(size = 2, alpha = 0.8, aes(color = data_pca$Treatment)) +  # Color points
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    text = element_text(size = 16)  # Ensures readable variable labels
  ) +
  xlab("Principal Component 1 (PC1)") +
  ylab("Principal Component 2 (PC2)") +
  ylim(-5, 5)  # Adjust Y-axis range
print(p)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Biplot (PC1 vs PC2) for Milestone Data at PND = 4 \n\n"))

# Scree Plot to Show Variance Explained
fviz_eig(pca_result)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Scree plot for Milestone Data at PND = 4 \n\n"))

loadings |>
  round(4) |>
  as.data.frame() |>
  tibble::rownames_to_column(var = "Variable") |>
  flextable::flextable() |>
  width(width = 1.15) |>
  width(j = 1, width = 1.16) |>
  autofit()
tab_n = tab_n + 1
cat(paste0("\nTable ",tab_n, ": Factor Loading for PCA Analysis at PND = 4 \n\n"))
```




### PCA Result at Week 8

```{r PCA-8, results='asis'}
# Getting the data ready for PCA
data_pca <- data |>
  filter(PND == "8") |>    # PCA at 8th week
  select(all_of(vars[1:9]),Treatment)
non_constant_vars <- names(data_pca %>% select(-Treatment))[sapply(data_pca %>% select(-Treatment), function(col) {
  if (all(is.na(col))) return(FALSE)  
  var(col, na.rm = TRUE) > 0  
})]
data_pca <- data_pca |>
  select(all_of(c(non_constant_vars,"Treatment"))) |>
  na.omit()
data_pca_scaled <- data_pca |> 
  select(-Treatment) |> 
  scale()

# Extract PCA Importance Scores
pca_result <- prcomp(data_pca_scaled, center = TRUE, scale. = TRUE)
pca_summary <- summary(pca_result)
importance_scores <- as.data.frame(pca_summary$importance)
colnames(importance_scores) <- paste0("PC", seq_len(ncol(importance_scores)))
loadings <- as.data.frame(pca_result$rotation)  # Extract PC1 & PC2 loadings

# PCA Biplot
p <- ggbiplot(pca_result, 
              choices = c(1,2),
              ellipse = TRUE,
              var.scale = 1,
              obs.scale = 1,
              varname.size = 5,  
              groups = data_pca$Treatment, 
              repel = TRUE) +
  geom_point(size = 2, alpha = 0.8, aes(color = data_pca$Treatment)) +  # Color points
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    text = element_text(size = 16)  # Ensures readable variable labels
  ) +
  xlab("Principal Component 1 (PC1)") +
  ylab("Principal Component 2 (PC2)") +
  ylim(-5, 5)  # Adjust Y-axis range
print(p)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Biplot (PC1 vs PC2) for Milestone Data at PND = 8 \n\n"))

# Scree Plot to Show Variance Explained
fviz_eig(pca_result)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Scree plot for Milestone Data at PND = 8 \n\n"))

loadings |>
  round(4) |>
  as.data.frame() |>
  tibble::rownames_to_column(var = "Variable") |>
  flextable::flextable() |>
  width(width = 1.15) |>
  width(j = 1, width = 1.16) |>
  autofit()
tab_n = tab_n + 1
cat(paste0("\nTable ",tab_n, ": Factor Loading for PCA Analysis at PND = 8 \n\n"))
```


### PCA Result at Week 12

```{r PCA-12, results='asis'}
# Getting the data ready for PCA
data_pca <- data |>
  filter(PND == "12") |>    # PCA at 12th week
  select(all_of(vars[1:9]),Treatment)
non_constant_vars <- names(data_pca %>% select(-Treatment))[sapply(data_pca %>% select(-Treatment), function(col) {
  if (all(is.na(col))) return(FALSE)  
  var(col, na.rm = TRUE) > 0  
})]
data_pca <- data_pca |>
  select(all_of(c(non_constant_vars,"Treatment"))) |>
  na.omit()
data_pca_scaled <- data_pca |> 
  select(-Treatment) |> 
  scale()

# Extract PCA Importance Scores
pca_result <- prcomp(data_pca_scaled, center = TRUE, scale. = TRUE)
pca_summary <- summary(pca_result)
importance_scores <- as.data.frame(pca_summary$importance)
colnames(importance_scores) <- paste0("PC", seq_len(ncol(importance_scores)))
loadings <- as.data.frame(pca_result$rotation)  # Extract PC1 & PC2 loadings

# PCA Biplot
p <- ggbiplot(pca_result, 
              choices = c(1,2),
              ellipse = TRUE,
              var.scale = 1,
              obs.scale = 1,
              varname.size = 5,  
              groups = data_pca$Treatment, 
              repel = TRUE) +
  geom_point(size = 2, alpha = 0.8, aes(color = data_pca$Treatment)) +  # Color points
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    text = element_text(size = 16)  # Ensures readable variable labels
  ) +
  xlab("Principal Component 1 (PC1)") +
  ylab("Principal Component 2 (PC2)") +
  ylim(-5, 5)  # Adjust Y-axis range
print(p)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Biplot (PC1 vs PC2) for Milestone Data at PND = 12 \n\n"))

# Scree Plot to Show Variance Explained
fviz_eig(pca_result)
fig_n = fig_n + 1
cat(paste0("\nFigure ",fig_n, 
           ": PCA Scree plot for Milestone Data at PND = 12 \n\n"))

loadings |>
  round(4) |>
  as.data.frame() |>
  tibble::rownames_to_column(var = "Variable") |>
  flextable::flextable() |>
  width(width = 1.15) |>
  width(j = 1, width = 1.16) |>
  autofit()
tab_n = tab_n + 1
cat(paste0("\nTable ",tab_n, ": Factor Loading for PCA Analysis at PND = 12 \n\n"))
```


