---
title: "NOR Analysis"
author: "Yunyi Ren, Matthew Dominic Ponzini"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
    df_print: paged
  pdf_document: default
  word_document: default
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE
)
```

```{r setup, include=FALSE}
library(mar4.0)

my_packages <- c("tidyverse", "kableExtra", "knitr", "readxl","lubridate", 
                "arsenal","tibbletime","dplyr","ggplot2", "ggpubr", 
                "sjPlot","car","DT")
lapply(my_packages, library, character.only = TRUE)
```

# Background

In MAR 4.0 project, investigator wants to know pre-lim analysis result on NOR with variable Discrimination Index. 

We are going to run the analysis use R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.

# Summary Statistics

```{r}
data = nor
vars <- c("Discirmination Index")

formula = as.formula("`Treat` ~ `Discirmination Index`")

tab = tableby(formula,
    data=data, numeric.stats=c("Nmiss","meansd","medianq1q3", "range"),digits=2,num.test="anova",test=F)

#fdr = as.data.frame(summary(padjust(tab, method = "fdr")))$'p value'

tab = as.data.frame(summary(tab))

#tab$'FDR' = fdr

kable(tab,format="markdown", caption = "Summary Table of Discirmination Index by Treatment")
```

# Regression Analysis

We will start with the interaction model and reduce it step by step to the null model. The mixed effect linear regression is used on estimating the effect.

## Interaction (Full) Model

To explore whether gender influences the results, we fit the following model:

The model we fit is: `Discrimination Index` =  `Treatment` + `Sex` + `Treatment * Sex` + `(1 | Litter ID)`

The results suggest a Sex is significant under the interaction model, but the sex-treatment is not significant.

```{r results='markup',warning = FALSE}
NOR_tb_anova <- data.frame(Variables = c("Treatment","Sex","Interaction"))

# Fit the mixed effects model
lmer_model <- lmerTest::lmer(as.formula(paste0("`",vars[1],"`", "~ `Treat` + `Sex` + `Treat` * `Sex` + (1 | `Litter ID`)")), data = data) 
lmer_model2 <- lmerTest::lmer(
  as.formula(
    paste0("`",vars[1],"`", "~ `Treat` + `Sex` + (1 | `Litter ID`)")
    ), 
  data = data) 
lmer_summary <- lmer_model |> summary()
# Convert the result into a data frame for clarity
formatted_summary <- apply(lmer_summary$coefficients[, c(1, 2, 5)], 1, function(row) {
  estimate <- round(row[1], 3)
  se <- round(row[2], 3)
  p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
  formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
  return(formatted_row)
})
formatted_summary_df <- as.data.frame(t(formatted_summary))
colnames(formatted_summary_df) <- c("Estimate (SE) - Interaction", "P-Value")
# Extract the ANOVA Type III p-value
NOR_tb_anova[1:3, "ANOVA III P-value"] = Anova(lmer_model, type = "III")$`Pr(>Chisq)`[2:4]

p_list_3 <- 
  emmeans::contrast(emmeans::emmeans(lmer_model, ~ Treat), 
                    method = "pairwise", 
                    adjust = "none") |>
  as.data.frame() |> 
  mutate(PND = 4) |>
  select(`contrast`, `p.value`, `PND`)
p_list_2 <- 
  emmeans::contrast(emmeans::emmeans(lmer_model2, ~ Treat), 
                    method = "pairwise", 
                    adjust = "none") |>
  as.data.frame() |> 
  mutate(PND = 4) |>
  select(`contrast`, `p.value`, `PND`)

post_hoc_h1_list <- 
  run_posthoc_test_num_h1(
    data = data |> 
      mutate(PND = 4, Animal = as.factor(seq_len(n())),
             Treatment = Treat), 
    var = vars[1], 
    p_list_3 = p_list_3,
    p_list_2 = p_list_2) |>
  select(-PND)

post_hoc_h2_list <- 
  run_posthoc_test_num_h2(
    data = data |> 
      mutate(PND = 4, Animal = as.factor(seq_len(n())),
             Treatment = Treat), 
    var = vars[1], 
    p_list_3 = p_list_3,
    p_list_2 = p_list_2) |>
  select(-PND)
  
formatted_summary_df |> 
  kable(format = "markdown",
        style = "ccc", 
        caption = "Summary Estiamted table on regression analysis")

merged_summary_df = formatted_summary_df
```

### Type III ANOVA 

```{r summary-3, results='asis'}
NOR_tb_anova %>% 
  kable(format='markdown',align="ccc",caption = "Type III ANOVA P-value of Treatment and Sex Interaction Effect on Discrimination Index")
```

```{r ,include=F}
performance::check_model(lmer_model, re_formula = NA)
```


## Sex Adjustment Model

To explore whether gender influences the results, we fit the following model without considering interaction:

The model we fit is: `Discrimination Index` =  `Treatment` + `Sex` + `(1 | Litter ID)`

The results suggest Sex is not significant under the sex adjustment model.

```{r results='markup',warning = FALSE}
NOR_tb_anova <- data.frame(Variables = c("Treatment","Sex"))

# Loop through each variable
#for (i in 1:length(vars)) {
  # Fit the mixed effects model
  lmer_model <- lmerTest::lmer(as.formula(paste0("`",vars[1],"`", "~ `Treat` + `Sex` + (1 | `Litter ID`)")), data = data) 
  lmer_summary <- lmer_model |> summary()
  # Convert the result into a data frame for clarity
  formatted_summary <- apply(lmer_summary$coefficients[, c(1, 2, 5)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
    formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
    return(formatted_row)
  })
  formatted_summary_df <- as.data.frame(t(formatted_summary))
  colnames(formatted_summary_df) <- c("Estimate (SE)", "P-Value")
  # Extract the ANOVA Type III p-value
  NOR_tb_anova[1:2, "ANOVA III P-value"] = Anova(lmer_model, type = "III")$`Pr(>Chisq)`[2:3]
  # NOR_tb_anova[i, "False Discovery Rate (FDR)"] = NA
  # emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treat`,adjust = "none")
  # 
  # # Obtaining Contrasts
  # for (j in 1:length(contrast_index)){
  #   NOR_tb_anova[i,contrast_label[j]] = emm_s.t$contrasts[contrast_index[j]] %>% 
  #     as.data.frame() %>% 
  #     select(p.value) %>% 
  #     as.numeric %>%
  #     round(4)
  # }
#}
# NOR_tb_anova$`False Discovery Rate (FDR)` <- round(p.adjust(NOR_tb_anova$`ANOVA III P-value`, method = "BH"),6)
# NOR_tb_anova$`ANOVA III P-value` <- round(NOR_tb_anova$`ANOVA III P-value`,6)
  
formatted_summary_df |> kable(format = "markdown",style = "ccc", caption = "Summary Estiamted table on regression analysis")

merged_summary_df$`Estimate (SE) Add Sex` = c(formatted_summary_df$`Estimate (SE)`,rep("",4))
merged_summary_df$`P-Value ` = c(formatted_summary_df$`P-Value` ,rep("",4))
```

### Type III ANOVA 

```{r summary-2, results='asis'}
NOR_tb_anova %>% 
  kable(format='markdown',align="ccc",caption = "Type III ANOVA P-value of Treatment Effect on Discrimination Index")
```

```{r ,include=F}
performance::check_model(lmer_model, re_formula = NA)
```

## Null Model

The model we fit is: `discrimination Index` =  `Treatment` + `(1 | Litter ID)`

The results suggest treatment is not significantly associated to Discrimination Index.

```{r results='markup'}
NOR_tb_anova <- data.frame(Variables = "Treatment")

# Loop through each variable
#for (i in 1:length(vars)) {
  # Fit the mixed effects model
  lmer_model <- lmerTest::lmer(as.formula(paste0("`",vars[1],"`", "~ `Treat` + (1 | `Litter ID`)")), data = data) 
  lmer_summary <- lmer_model |> summary()
  # Convert the result into a data frame for clarity
  formatted_summary <- apply(lmer_summary$coefficients[, c(1, 2, 5)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
    formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
    return(formatted_row)
  })
  formatted_summary_df <- as.data.frame(t(formatted_summary))
  colnames(formatted_summary_df) <- c("Estimate (SE)", "P-Value")

  
  # Extract the ANOVA Type III p-value
  NOR_tb_anova[1, "ANOVA III P-value"] = Anova(lmer_model, type = "III")$`Pr(>Chisq)`[2]
  
  # NOR_tb_anova[i, "False Discovery Rate (FDR)"] = NA
  # emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treat`,adjust = "none")
  # 
  # # Obtaining Contrasts
  # for (j in 1:length(contrast_index)){
  #   NOR_tb_anova[i,contrast_label[j]] = emm_s.t$contrasts[contrast_index[j]] %>% 
  #     as.data.frame() %>% 
  #     select(p.value) %>% 
  #     as.numeric %>%
  #     round(4)
  # }
#}
# NOR_tb_anova$`False Discovery Rate (FDR)` <- round(p.adjust(NOR_tb_anova$`ANOVA III P-value`, method = "BH"),6)
# NOR_tb_anova$`ANOVA III P-value` <- round(NOR_tb_anova$`ANOVA III P-value`,6)
  
formatted_summary_df |> kable(format = "markdown",style = "ccc", caption = "Summary Estiamted table on regression analysis")

merged_summary_df$`Estimate (SE) Basic` = c(formatted_summary_df$`Estimate (SE)`,rep("",5))
merged_summary_df$`P-Value  ` = c(formatted_summary_df$`P-Value` ,rep("",5))
```

### Type III ANOVA 

Type III ANOVA indicates that by considering treatment, there is no benefit of predicting Discrimination Index.

```{r summary-1, results='asis'}
NOR_tb_anova %>% 
  kable(format='markdown',align="ccc",caption = "Type III ANOVA P-value of Treatment Effect on discrimination Index")
```

```{r include=F}
# Also, see the linear model diagnostic criteria below.

performance::check_model(lmer_model, re_formula = NA)
```


## Table of Estimate Comparison

The overall estimation table also displayed here for all the models for an easier comparison.

```{r summary-all, results='asis'}
merged_summary_df |> kable(format = "markdown",style = "ccc", caption = "Summary Estiamted table on regression analysis for All the Models")
```

## Boxplot by Treatment and Sex on Discrimination Index

Here we running the boxplot for discrimination index by sex and treatment to show an insignificant result.

The data shows a higher discrimination index in female group within the placebo group and higher DI in male group in CRMP1+2 group, and this results the significance in the interaction model.

```{r}
ggplot(data, aes(x = Treat, y = `Discirmination Index`, fill = Sex)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 3, color = "black", position = position_dodge(width = 0.75)) +
  labs(
    title = "Grouped Boxplot of discrimination Index by Treat and Sex",
    x = "Treatment",
    y = "discrimination Index"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  scale_fill_brewer(palette = "Set2", name = "Sex") + 
  geom_jitter()
```

## Post-Hoc Comparison on Treatment

```{r results='asis'}
post_hoc_h1_list |> kable(format = "markdown",style = "ccc", caption = "Post Hoc Comparison on regression analysis result - Primary Hypothesis")
```

```{r results='asis'}
post_hoc_h2_list |> kable(format = "markdown",style = "ccc", caption = "Post Hoc Comparison on regression analysis result - Secondary Hypothesis")
```
