---
title: "NOR-Report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NOR-Report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE
)
```

```{r setup, include=FALSE}
library(mar4.0)

my_packages <- c("tidyverse", "kableExtra", "knitr", "readxl","lubridate", 
                "arsenal","tibbletime","dplyr","ggplot2", "ggpubr", 
                "sjPlot","car","DT")
lapply(my_packages, library, character.only = TRUE)
```

# Background

In MAR4.0 project, investigator wants to know pre-lim analysis result on NOR with variable discrimination index. 

We are going to run the analysis use R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.

# Summary Statistics

```{r}
data = nor
vars <- c("Discirmination Index")

formula = as.formula("`Treat` ~ `Discirmination Index`")

tab = tableby(formula,
    data=data, numeric.stats=c("Nmiss","meansd","medianq1q3", "range"),digits=2,num.test="anova",test=F)

#fdr = as.data.frame(summary(padjust(tab, method = "fdr")))$'p value'

tab = as.data.frame(summary(tab))

#tab$'FDR' = fdr

kable(tab,format="markdown", caption = "Summary Table of Discirmination Index by Treatment")
```

# Regression Analysis

The model we fit is: `Discirmination Index` =  `Treatment` + `(1 | Litter ID)`

## Main effect

When running the model, "boundary (singular) fit: see help('isSingular')" shows up as alert.

```{r results='markup'}
NOR_tb_anova <- data.frame(Variables = vars)

# Loop through each variable
#for (i in 1:length(vars)) {
  # Fit the mixed effects model
  lmer_model <- lmerTest::lmer(as.formula(paste0("`",vars[1],"`", "~ `Treat` + (1 | `Litter ID`)")), data = data) 
  lmer_model |> summary()
  # Extract the ANOVA Type III p-value
  NOR_tb_anova[1, "ANOVA III P-value"] = Anova(lmer_model, type = "III")$`Pr(>Chisq)`[2]
  # NOR_tb_anova[i, "False Discovery Rate (FDR)"] = NA
  # emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treat`,adjust = "none")
  # 
  # # Obtaining Contrasts
  # for (j in 1:length(contrast_index)){
  #   NOR_tb_anova[i,contrast_label[j]] = emm_s.t$contrasts[contrast_index[j]] %>% 
  #     as.data.frame() %>% 
  #     select(p.value) %>% 
  #     as.numeric %>%
  #     round(4)
  # }
#}
# NOR_tb_anova$`False Discovery Rate (FDR)` <- round(p.adjust(NOR_tb_anova$`ANOVA III P-value`, method = "BH"),6)
# NOR_tb_anova$`ANOVA III P-value` <- round(NOR_tb_anova$`ANOVA III P-value`,6)
```

## Type III ANOVA 

Type III ANOVA indicates that by considering treatment, there is no benefit of predicting discrimination index.

```{r summary-1, results='asis'}
NOR_tb_anova %>% 
  kable(format='markdown',align="ccc",caption = "Type III ANOVA P-value of Treatment Effect on Discrimination Index")
```

## Model Performance Check

Also, see the linear model diagnostic criteria below.

```{r}
performance::check_model(lmer_model, re_formula = NA)
```

