---
title: "NOR Analysis"
author: "Yunyi Ren, Matthew Dominic Ponzini"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
    df_print: paged
  pdf_document: default
  word_document: default
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE
)
```

```{r setup, include=FALSE}
library(mar4.0)

my_packages <- c("tidyverse", "kableExtra", "knitr", "readxl","lubridate", 
                "arsenal","tibbletime","dplyr","ggplot2", "ggpubr", 
                "sjPlot","car","DT")
lapply(my_packages, library, character.only = TRUE)
```

# Background

In MAR4.0 project, investigator wants to know pre-lim analysis result on NOR with variable discrimination index. 

We are going to run the analysis use R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.

# Summary Statistics

```{r}
data = nor
vars <- c("Discirmination Index")

formula = as.formula("`Treat` ~ `Discirmination Index`")

tab = tableby(formula,
    data=data, numeric.stats=c("Nmiss","meansd","medianq1q3", "range"),digits=2,num.test="anova",test=F)

#fdr = as.data.frame(summary(padjust(tab, method = "fdr")))$'p value'

tab = as.data.frame(summary(tab))

#tab$'FDR' = fdr

kable(tab,format="markdown", caption = "Summary Table of Discirmination Index by Treatment")
```

# Regression Analysis

We will start with the interaction model and reduce it step by step to the null model.

## Interaction (Full) Model

To explore whether gender influences the results, we fit the following model:

The model we fit is: `Discirmination Index` =  `Treatment` + `Sex` + `Treatment * Sex` + `(1 | Litter ID)`

The results suggest a significant interaction between Treatment and Sex, indicating that the effect of the treatment varies across sexes.

```{r results='markup',warning = FALSE}
NOR_tb_anova <- data.frame(Variables = c(vars,"Sex","Interaction"))

# Loop through each variable
#for (i in 1:length(vars)) {
  # Fit the mixed effects model
  lmer_model <- lmerTest::lmer(as.formula(paste0("`",vars[1],"`", "~ `Treat` + `Sex` + `Treat` * `Sex` + (1 | `Litter ID`)")), data = data) 
  lmer_summary <- lmer_model |> summary()
  # Convert the result into a data frame for clarity
  formatted_summary <- apply(lmer_summary$coefficients[, c(1, 2, 5)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
    formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
    return(formatted_row)
  })
  formatted_summary_df <- as.data.frame(t(formatted_summary))
  colnames(formatted_summary_df) <- c("Estimate (SE)", "P-Value")
  # Extract the ANOVA Type III p-value
  NOR_tb_anova[1:3, "ANOVA III P-value"] = Anova(lmer_model, type = "III")$`Pr(>Chisq)`[2:4]
  # NOR_tb_anova[i, "False Discovery Rate (FDR)"] = NA
  # emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treat`,adjust = "none")
  # 
  # # Obtaining Contrasts
  # for (j in 1:length(contrast_index)){
  #   NOR_tb_anova[i,contrast_label[j]] = emm_s.t$contrasts[contrast_index[j]] %>% 
  #     as.data.frame() %>% 
  #     select(p.value) %>% 
  #     as.numeric %>%
  #     round(4)
  # }
#}
# NOR_tb_anova$`False Discovery Rate (FDR)` <- round(p.adjust(NOR_tb_anova$`ANOVA III P-value`, method = "BH"),6)
# NOR_tb_anova$`ANOVA III P-value` <- round(NOR_tb_anova$`ANOVA III P-value`,6)
  
formatted_summary_df |> kable(format = "markdown",style = "ccc", caption = "Summary Estiamted table on regression analysis")
```

### Type III ANOVA 

```{r summary-3, results='asis'}
NOR_tb_anova %>% 
  kable(format='markdown',align="ccc",caption = "Type III ANOVA P-value of Treatment Effect on Discrimination Index")
```

```{r ,include=F}
performance::check_model(lmer_model, re_formula = NA)
```


## Sex Adjustment Model

To explore whether gender influences the results, we fit the following model without considering interaction:

The model we fit is: `Discirmination Index` =  `Treatment` + `Sex` + `(1 | Litter ID)`

The results suggest a significant interaction between Treatment and Sex, indicating that the effect of the treatment varies across sexes.

```{r results='markup',warning = FALSE}
NOR_tb_anova <- data.frame(Variables = c(vars,"Sex"))

# Loop through each variable
#for (i in 1:length(vars)) {
  # Fit the mixed effects model
  lmer_model <- lmerTest::lmer(as.formula(paste0("`",vars[1],"`", "~ `Treat` + `Sex` + (1 | `Litter ID`)")), data = data) 
  lmer_summary <- lmer_model |> summary()
  # Convert the result into a data frame for clarity
  formatted_summary <- apply(lmer_summary$coefficients[, c(1, 2, 5)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
    formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
    return(formatted_row)
  })
  formatted_summary_df <- as.data.frame(t(formatted_summary))
  colnames(formatted_summary_df) <- c("Estimate (SE)", "P-Value")
  # Extract the ANOVA Type III p-value
  NOR_tb_anova[1:2, "ANOVA III P-value"] = Anova(lmer_model, type = "III")$`Pr(>Chisq)`[2:3]
  # NOR_tb_anova[i, "False Discovery Rate (FDR)"] = NA
  # emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treat`,adjust = "none")
  # 
  # # Obtaining Contrasts
  # for (j in 1:length(contrast_index)){
  #   NOR_tb_anova[i,contrast_label[j]] = emm_s.t$contrasts[contrast_index[j]] %>% 
  #     as.data.frame() %>% 
  #     select(p.value) %>% 
  #     as.numeric %>%
  #     round(4)
  # }
#}
# NOR_tb_anova$`False Discovery Rate (FDR)` <- round(p.adjust(NOR_tb_anova$`ANOVA III P-value`, method = "BH"),6)
# NOR_tb_anova$`ANOVA III P-value` <- round(NOR_tb_anova$`ANOVA III P-value`,6)
  
formatted_summary_df |> kable(format = "markdown",style = "ccc", caption = "Summary Estiamted table on regression analysis")
```

### Type III ANOVA 

```{r summary-2, results='asis'}
NOR_tb_anova %>% 
  kable(format='markdown',align="ccc",caption = "Type III ANOVA P-value of Treatment Effect on Discrimination Index")
```

```{r ,include=F}
performance::check_model(lmer_model, re_formula = NA)
```

## Null Model

The model we fit is: `Discirmination Index` =  `Treatment` + `(1 | Litter ID)`

```{r results='markup'}
NOR_tb_anova <- data.frame(Variables = vars)

# Loop through each variable
#for (i in 1:length(vars)) {
  # Fit the mixed effects model
  lmer_model <- lmerTest::lmer(as.formula(paste0("`",vars[1],"`", "~ `Treat` + (1 | `Litter ID`)")), data = data) 
  lmer_summary <- lmer_model |> summary()
  # Convert the result into a data frame for clarity
  formatted_summary <- apply(lmer_summary$coefficients[, c(1, 2, 5)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3)<0.001,"<0.001",round(row[3], 3))
    formatted_row <- c(paste0(estimate, " (", se, ")"), p_value)
    return(formatted_row)
  })
  formatted_summary_df <- as.data.frame(t(formatted_summary))
  colnames(formatted_summary_df) <- c("Estimate (SE)", "P-Value")

  
  # Extract the ANOVA Type III p-value
  NOR_tb_anova[1, "ANOVA III P-value"] = Anova(lmer_model, type = "III")$`Pr(>Chisq)`[2]
  
  # NOR_tb_anova[i, "False Discovery Rate (FDR)"] = NA
  # emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treat`,adjust = "none")
  # 
  # # Obtaining Contrasts
  # for (j in 1:length(contrast_index)){
  #   NOR_tb_anova[i,contrast_label[j]] = emm_s.t$contrasts[contrast_index[j]] %>% 
  #     as.data.frame() %>% 
  #     select(p.value) %>% 
  #     as.numeric %>%
  #     round(4)
  # }
#}
# NOR_tb_anova$`False Discovery Rate (FDR)` <- round(p.adjust(NOR_tb_anova$`ANOVA III P-value`, method = "BH"),6)
# NOR_tb_anova$`ANOVA III P-value` <- round(NOR_tb_anova$`ANOVA III P-value`,6)
  
formatted_summary_df |> kable(format = "markdown",style = "ccc", caption = "Summary Estiamted table on regression analysis")
```

### Type III ANOVA 

Type III ANOVA indicates that by considering treatment, there is no benefit of predicting discrimination index.

```{r summary-1, results='asis'}
NOR_tb_anova %>% 
  kable(format='markdown',align="ccc",caption = "Type III ANOVA P-value of Treatment Effect on Discrimination Index")
```

```{r include=F}
# Also, see the linear model diagnostic criteria below.

performance::check_model(lmer_model, re_formula = NA)
```
