---
title: "MAR 4.0 - Social Dyads"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2025-03-21"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE , warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, mar4.0)

sd_data = sd_adult |>
  rename(Treatment = Treat) |>
  left_join(weights |> 
              filter(Week == 15) |> 
              select(`Animal ID`,Weights), 
            by = "Animal ID") |>
  mutate(`Total number Grooming Self` = as.numeric(`Total number Grooming Self`),
         Weights = as.numeric(Weights))
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glmmTMB::glmmTMB(as.formula(formula_str), ziformula = ~ 1, family = glmmTMB::tweedie(link = "log"), data = data))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}

# Function to summarize and format model results
format_summary_lmer <- function(model_summary, column_name) {
  formatted <- apply(model_summary$coefficients[, c(1,2,4)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

format_summary_glmmTMB <- function(model_summary, column_name) {
  formatted <- apply(model_summary$coefficients$cond[, c(1,2,4)], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(..., summary_full) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  for (i in 2:length(summaries)) {
    merged_data <- merge(merged_data, summaries[[i]], by = "row.names", all = TRUE)
    rownames(merged_data) <- merged_data$Row.names
    merged_data <- merged_data[ , !(colnames(merged_data) %in% "Row.names")]
  }
  ordered_data <- merged_data[order(match(rownames(merged_data), rownames(summary_full))), ]
  ordered_data[is.na(ordered_data)] = ""
  return(ordered_data)
}

# Function to perform ANOVA
perform_anova <- function(model, type = "III") {
  return(Anova(model1, type = type)[, 3])
}
```




```{r}
tx_tbl <- sd_data |>
  dplyr::select(
    Treatment, Weights,
    `Total duration Non-social (Total)`, 
    `Total duration Social Investigation (Total)`, 
    `Total duration Play`, 
    `Total duration Grooming Self`,
    `Total number Play`,
    `Total number Grooming Self`
  ) |>
  gtsummary::tbl_summary(
    by = Treatment,
    type = list(
      Weights ~ "continuous2",
      `Total duration Non-social (Total)` ~ "continuous2",
      `Total duration Social Investigation (Total)` ~ "continuous2",
      `Total duration Play` ~ "continuous2",
      `Total duration Grooming Self` ~ "continuous2",
      `Total number Play` ~ "continuous2",
      `Total number Grooming Self` ~ "continuous2"
    ),
    statistic = list(
      gtsummary::all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} [{p25}, {p75}]",
        "{min}, {max}"
      )
    )
  ) |>
  gtsummary::add_overall()


tx_sex_tbl <- sd_data |>
  dplyr::select(
    Treatment, Weights, Sex,
    `Total duration Non-social (Total)`, 
    `Total duration Social Investigation (Total)`, 
    `Total duration Play`, 
    `Total duration Grooming Self`,
    `Total number Play`,
    `Total number Grooming Self`
  ) |>
  gtsummary::tbl_strata(
    strata = Sex,
    .tbl_fun = 
      ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = list(
            Weights ~ "continuous2",
            `Total duration Non-social (Total)` ~ "continuous2",
            `Total duration Social Investigation (Total)` ~ "continuous2",
            `Total duration Play` ~ "continuous2",
            `Total duration Grooming Self` ~ "continuous2",
            `Total number Play` ~ "continuous2",
            `Total number Grooming Self` ~ "continuous2"
          ),
          statistic = list(
            gtsummary::all_continuous2() ~ c(
              "{mean} ({sd})",
              "{median} [{p25}, {p75}]",
              "{min}, {max}"
            )
          )
        ) |>
        gtsummary::add_overall() |>
        gtsummary::modify_header(
          gtsummary::all_stat_cols() ~ "**{level}**"
        ),
    .combine_with = 'tbl_stack'
  )
```


```{r warning=FALSE}
# Comment Code to examine the distribution and model fitting
# hist(sd_data$`Total duration Non-social (Total)`)
# hist(sd_data$`Total duration Play`)
# hist(sd_data$`Total duration Social Investigation (Total)`)
# hist(sd_data$`Total duration Grooming Self`)
# hist(sd_data$`Total number Play`)
# hist(sd_data$`Total number Grooming Self`)
# sim_res <- DHARMa::simulateResiduals(fittedModel = model1)
# hist(sim_res$fittedResiduals)
# Plot residuals vs fitted values
# DHARMa::plotResiduals(sim_res$fittedPredicted, sim_res$fittedResiduals,
#               main = "Residuals vs Fitted Values",
#               xlab = "Fitted Values",
#               ylab = "Scaled Residuals")

data = sd_data
vars = c("Total duration Non-social (Total)", 
         "Total duration Play", 
         "Total duration Social Investigation (Total)", 
         "Total duration Grooming Self",
         "Total number Play",   # THIS IS LIKELY TO BE 0 INFLATED
         "Total number Grooming Self")

usv_anova_tb <- data.frame(Variables = c("Interaction - Treatment",
                                         "Interaction - Sex",
                                         "Interaction - Treatment:Sex", 
                                         "No Interaction - Treatment",
                                         "No Interaction - Sex"))
merged_summary_list <- list()

# Main loop
for (i in 1:4) {
  var <- vars[i]
  model_type <- if (var %in% c("Total duration Play", "Total duration Grooming Self")) "glmer" else "lmer"
  
  # Define formulas
  formula1 <-  paste0("`", var, "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex` + Weights + (1 | `Litter ID`)")
  formula2 <-  paste0("`", var, "` ~ `Treatment` + `Sex` +  Weights + (1 | `Litter ID`)")

  # Fit models
  model1 <- fit_model(var, formula1, data, model_type)
  model2 <- fit_model(var, formula2, data, model_type)
  
  # Summarize and format results
  if (model_type == "lmer") {
    summary1 <- format_summary_lmer(summary(model1), "Three Interac")
    summary2 <- format_summary_lmer(summary(model2), "Two - Full")
  } else {
    summary1 <- format_summary_glmmTMB(summary(model1), "Three Interac")
    summary2 <- format_summary_glmmTMB(summary(model2), "Two - Full")
  }

  merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
  merged_summary_list[[i]] <- merged_summary_df
  
  # Perform ANOVA
  usv_anova_tb[1:3, var] <- perform_anova(model1)[2:4]
  usv_anova_tb[4:5, var] <- perform_anova(model2)[2:3]

#   emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treatment`,adjust = "none")
#   usv_tb_contrast = emm_s.t$contrasts %>%
#     as.data.frame() %>%
#     select(estimate, p.value) %>%
#     mutate(across(everything(), ~ round(.x, 3)))
#   rownames(usv_tb_contrast) <- rownames(emm_s.t$contrasts)
}

# #5 Total number Play - THIS IS a 0 INFLATED VARIABLE

formula1 <-  paste0("`", vars[5], "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex` + Weights + (1 | `Litter ID`)")
formula2 <-  paste0("`", vars[5], "` ~ `Treatment` + `Sex` +  Weights + (1 | `Litter ID`)")
model1 <- glmmTMB::glmmTMB(as.formula(formula1), family = glmmTMB::nbinom2(link = "log"), data = data |> filter(!is.na(!!sym(vars[5]))))
model2 <- glmmTMB::glmmTMB(as.formula(formula2), family = glmmTMB::nbinom2(link = "log"), data = data |> filter(!is.na(!!sym(vars[5]))))
summary1 <- format_summary_glmmTMB(summary(model1), "Three Interac")
summary2 <- format_summary_glmmTMB(summary(model2), "Two - Full")
merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
merged_summary_list[[5]] <- merged_summary_df
usv_anova_tb[1:3, vars[5]] <- perform_anova(model1)[2:4]
usv_anova_tb[4:5, vars[5]] <- perform_anova(model2)[2:3]
  

# #6 Total number Grooming Self
formula1 <-  paste0("`", vars[6], "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex` + Weights + (1 | `Litter ID`)")
formula2 <-  paste0("`", vars[6], "` ~ `Treatment` + `Sex` +  Weights + (1 | `Litter ID`)")
model1 <- glmmTMB::glmmTMB(as.formula(formula1), family = poisson(link = "log"), data = data |> filter(!is.na(!!sym(vars[6]))))
model2 <- glmmTMB::glmmTMB(as.formula(formula2), family = poisson(link = "log"), data = data |> filter(!is.na(!!sym(vars[6]))))
summary1 <- format_summary_glmmTMB(summary(model1), "Three Interac")
summary2 <- format_summary_glmmTMB(summary(model2), "Two - Full")
merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
merged_summary_list[[6]] <- merged_summary_df
usv_anova_tb[1:3, vars[6]] <- perform_anova(model1)[2:4]
usv_anova_tb[4:5, vars[6]] <- perform_anova(model2)[2:3]

usv_anova_tb = usv_anova_tb |> 
  column_to_rownames(var = "Variables") |>
  t() |>
  round(3)
```

# Background

In MAR 4.0 project, this section we will wants to know analysis result on the	Social Dyads data. 

We are going to run the analysis use R-package structure within github. R 4.4.3 is used for the analysis, and we are wrapping all the materials into a R package.

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with a final sample size of `r usv |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. 

The interested variables at USVs data are: `r paste(vars, collapse = ", ")`.

Weight for adult use the data at the 15 month.

# Summary statistics

The all interested variables were summarized by Treatment group overall (Table \@ref(tab:summary-tbl)), and by treatment groups stratified by sex (Table \@ref(tab:summary-tbl-sex)).

```{r summary-tbl}
tx_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of USV by Treatment Groups."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 1.16
  )
```

\

```{r summary-tbl-sex}
tx_sex_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of USV by Treatment Groups, stratified",
      "by Sex."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 0.7
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```

# Model Results

The relationship between treatment groups and variables of interest was assessed using linear mixed-effects models (Total duration Non-social and Total duration Social Investigation) or generalized regression with *Tweedie* distribution (Total duration Play and Total duration Grooming Self), incorporating a random litter effect. Negative binomial regression also examined but the fitting performance is worse than Tweedie.

Negative binomial or Poisson Regression are fitted to count variables.

The regression fitting process followed these steps:

- Fit the interaction: Begin with a model interactions.

- Evaluate the interaction: If the interaction is not significant, proceed with a model without interaction.

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report's delivery format.

## ANOVA Table

Type III ANOVA table (Table \@ref(tab:anova-tb)) display the anova p-values for each variables at each models, and below is the explanation of each model:

Final two-way interaction model (last column) show the p-value on the addition two-way interaction: y = trt + sex + *time(PND) x trt*

**The outcome "Startle" and "Forelimb Grasp" are fitted by logistic regression, and they are all or mostly 1 on p-values due to not enough incidence on the outcomes.**


```{r anova-tb}
usv_anova_tb |>
  as.data.frame() |>
  tibble::rownames_to_column("Variable") |>
  flextable::flextable() |>
  flextable::set_caption(
    caption = paste0(
      "ANOVA P-Value Table of USV Object in the Three Different Model ",
      "(Interaction, Additive, and Basic)"
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 1.5
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```

## Estimation Table

The overall estimation tables also displayed here for all the models for an easier comparison. The no interaction model is also displayed for examine purpose.

```{r results='asis'}
# Loop to generate flextables
for (i in 1:length(vars)) {
  merged_summary_list[[i]] |>
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    flextable::flextable() |>
    flextable::set_caption(
      caption = paste0(
        "Model Estimation Result in the Model - ", vars[i]
      )
    ) |>
    flextable::width(
      width = 1.15
    ) |>
    flextable::width(
      j = 1,
      width = 1.5
    ) |>
    flextable::flextable_to_rmd()

}
```
