---
title: "MAR 4.0 - Social Dyads"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2025-01-09"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, mar4.0)
sd_data = sd_adult |>
  rename(Treatment = Treat)
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glmmTMB::glmmTMB(as.formula(formula_str), ziformula = ~ 1, family = glmmTMB::tweedie(link = "log"), data = data))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}



# Function to summarize and format model results
format_summary <- function(model_summary, coef_indices, column_name) {
  formatted <- apply(model_summary$coefficients[, coef_indices], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(..., summary_full) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  for (i in 2:length(summaries)) {
    merged_data <- merge(merged_data, summaries[[i]], by = "row.names", all = TRUE)
    rownames(merged_data) <- merged_data$Row.names
    merged_data <- merged_data[ , !(colnames(merged_data) %in% "Row.names")]
  }
  ordered_data <- merged_data[order(match(rownames(merged_data), rownames(summary_full))), ]
  ordered_data[is.na(ordered_data)] = ""
  return(ordered_data)
}

# Function to perform ANOVA
perform_anova <- function(model, type = "III") {
  return(Anova(model1, type = type)[, 3])
}
```




```{r}
tx_tbl <- sd_data |>
  dplyr::select(
    Treatment, `Total duration Non-social (Total)`, `Total duration Social Investigation (Total)`, `Total duration Play`, `Total duration Grooming Self`) |>
  gtsummary::tbl_summary(
    by = Treatment,
    statistic = list(
      gtsummary::all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} [{p25}, {p75}]",
        "{min}, {max}"
      )
    )
  ) |>
  gtsummary::add_overall()

tx_sex_tbl <- sd_data |>
  dplyr::select(
    Treatment, Sex, `Total duration Non-social (Total)`, `Total duration Social Investigation (Total)`, `Total duration Play`, `Total duration Grooming Self`) |>
  gtsummary::tbl_strata(
    strata = Sex,
    .tbl_fun = 
      ~ .x |>
      gtsummary::tbl_summary(
        by = Treatment,
        statistic = list(
          gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})",
            "{median} [{p25}, {p75}]",
            "{min}, {max}"
          )
        )
      ) |>
      gtsummary::add_overall() |>
      gtsummary::modify_header(
        gtsummary::all_stat_cols() ~ "**{level}**"
      ),
    .combine_with = 'tbl_stack'
  )

hist(sd_data$`Total duration Non-social (Total)`)
hist(sd_data$`Total duration Play`)
hist(sd_data$`Total duration Social Investigation (Total)`)
hist(sd_data$`Total duration Grooming Self`)

```


```{r}
data = sd_data
vars = c("Total duration Non-social (Total)", 
         "Total duration Play", 
         "Total duration Social Investigation (Total)", 
         "Total duration Grooming Self")

usv_anova_tb <- data.frame(Variables = c("Interaction - Treatment",
                                         "Interaction - Sex",
                                         "Interaction - Treatment:Sex", 
                                         "No Interaction - Treatment",
                                         "No Interaction - Sex"))
merged_summary_list <- list()

# Main loop
for (i in 1:length(vars)) {
  var <- vars[i]
  model_type <- if (var %in% c("Total duration Play", "Total duration Grooming Self")) "glmer" else "lmer"
  
  # Define formulas
  formula1 <-  paste0("`", var, "` ~ `Treatment` + `Sex` +  `Treatment`:`Sex`  + (1 | `Litter ID`)")
  formula2 <-  paste0("`", var, "` ~ `Treatment` + `Sex` +  (1 | `Litter ID`)")

  # Fit models
  model1 <- fit_model(var, formula1, data, model_type)
  model2 <- fit_model(var, formula2, data, model_type)
  
  # Summarize and format results
  coef_indices <- if (model_type == "glmer") c(1, 2, 4) else c(1, 2, 5)
  summary1 <- format_summary(summary(model1), coef_indices, "Three Interac")
  summary2 <- format_summary(summary(model2), coef_indices, "Two - Full")
  merged_summary_df <- create_combined_estimate(summary_full = summary1, summary2)
  merged_summary_list[[i]] <- merged_summary_df
  
  # Perform ANOVA
  usv_anova_tb[1:3, var] <- perform_anova(model1)[2:4]
  usv_anova_tb[4:5, var] <- perform_anova(model2)[2:3]

#   emm_s.t <- emmeans::emmeans(lmer_model, pairwise ~  `Treatment`,adjust = "none")
#   usv_tb_contrast = emm_s.t$contrasts %>%
#     as.data.frame() %>%
#     select(estimate, p.value) %>%
#     mutate(across(everything(), ~ round(.x, 3)))
#   rownames(usv_tb_contrast) <- rownames(emm_s.t$contrasts)
}

usv_anova_tb = usv_anova_tb |> 
  column_to_rownames(var = "Variables") |>
  t() |>
  round(3)
```


# Background

In MAR 4.0 project, this section we will wants to know analysis result on the USV data. 

We are going to run the analysis use R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with a final sample size of `r usv |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. 

The interested variables at USVs data are: `r paste(vars, collapse = ", ")`.

# Summary statistics

The all interested variables were summarized by Treatment group overall (Table \@ref(tab:summary-tbl)), and by Treatment groups stratified by sex (Table \@ref(tab:summary-tbl-sex)).

```{r summary-tbl}
tx_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of USV by Treatment Groups."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 1.16
  )
```

\

```{r summary-tbl-sex}
tx_sex_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of USV by Treatment Groups, stratified",
      "by Sex."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 0.7
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```

# Model Results

The relationship between treatment groups and variables of interest was assessed using linear mixed-effects models or logistic regression, incorporating a random litter effect.

The regression fitting process followed these steps:

- Fit the 3-way interaction: Begin with a model including all three-way interactions.

- Evaluate the 3-way interaction: If the three-way interaction is not significant, proceed with a model including all two-way interactions.

- Simplify further if needed: If the model with all two-way interactions shows no significant interaction terms involving sex, remove those terms. The final model in this case is: y = time(PND) + trt + sex + PND*trt.

- Fit a no-interaction model if applicable: If the interaction between PND and treatment is not significant, fit a model without interaction terms.

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report's delivery format.

## ANOVA Table

Type III ANOVA table (Table \@ref(tab:anova-tb)) display the anova p-values for each variables at each models, and below is the explanation of each model:

Three interaction model show the p-value on three way interaction y = time(PND) + trt + sex + time(PND) x trt + time(PND) x sex + trt x sex + *time(PND) x trt x sex*

Two-way interaction full model show the p-value on each two-way interaction y = time(PND) + trt + sex + *time(PND) x trt* + *time(PND) x sex* + *trt x sex*

Two-way interaction models show the p-value on each addition two-way interaction:

- y = time(PND) + trt + sex + *time(PND) x trt* + *time(PND) x sex*

- y = time(PND) + trt + sex + *time(PND) x trt* + *trt x sex*

Final two-way interaction model (last column) show the p-value on the addition two-way interaction: y = time(PND) + trt + sex + *time(PND) x trt*

**The outcome "Startle" and "Forelimb Grasp" are fitted by logistic regression, and they are all or mostly 1 on p-values due to not enough incidence on the outcomes.**


```{r anova-tb}
usv_anova_tb |>
  as.data.frame() |>
  tibble::rownames_to_column("Variable") |>
  flextable::flextable() |>
  flextable::set_caption(
    caption = paste0(
      "ANOVA P-Value Table of USV Object in the Three Different Model ",
      "(Interaction, Additive, and Basic)"
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 0.7
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```

## Estimation Table

The overall estimation tables also displayed here for all the models for an easier comparison. The no interaction model is also displayed for examine purpose.

```{r results='asis'}
# Loop to generate flextables
for (i in 1:length(vars)) {
  merged_summary_list[[i]] |>
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    flextable::flextable() |>
    flextable::set_caption(
      caption = paste0(
        "Model Estimation Result in the Model - ", vars[i]
      )
    ) |>
    flextable::width(
      width = 1.15
    ) |>
    flextable::width(
      j = 1,
      width = 1.5
    ) |>
    flextable::flextable_to_rmd()

}
```
