---
title: "MAR 4.0 - Novel Object Recognition"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2024-10-29"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, mar4.0)
```

```{r}
nor_analytic <- nor |> dplyr::filter(!is.na(`Discirmination Index`)) |>
  dplyr::rename(Litter = `Litter ID`)
```

```{r}
discrim_tx_tbl <- nor |>
  dplyr::select(
    Treat, `Discirmination Index`
  ) |>
  gtsummary::tbl_summary(
    by = Treat,
    type = list(
      `Discirmination Index` ~ "continuous2"
    ),
    statistic = list(
      gtsummary::all_continuous2() ~ c(
        "{mean} ({sd})",
        "{median} [{p25}, {p75}]",
        "{min}, {max}"
      )
    )
  ) |>
  gtsummary::add_overall()

discrim_tx_sex_tbl <- nor_analytic |>
  dplyr::select(
    Treat, `Discirmination Index`, Sex
  ) |>
  gtsummary::tbl_strata(
    strata = Sex,
    .tbl_fun = 
      ~ .x |>
      gtsummary::tbl_summary(
        by = Treat,
        type = list(
          `Discirmination Index` ~ "continuous2"
        ),
        statistic = list(
          gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})",
            "{median} [{p25}, {p75}]",
            "{min}, {max}"
          )
        )
      ) |>
      gtsummary::add_overall() |>
      gtsummary::modify_header(
        gtsummary::all_stat_cols() ~ "**{level}**"
      ),
    .combine_with = 'tbl_stack'
  )
```

```{r fit-models}
# fit model with Tx*Sex
fit_sex_intrxn <- lmerTest::lmer(
  `Discirmination Index` ~ Treat*Sex + (1|Litter),
  data = nor_analytic
)

fit_sex_intrxn_type3 <- car::Anova(fit_sex_intrxn, type = 3)

# fit model with Tx + Sex
fit_sex <- lmerTest::lmer(
  `Discirmination Index` ~ Treat + Sex + (1|Litter),
  data = nor_analytic
)

fit_sex_type3 <- car::Anova(fit_sex, type = 3)

# fit model with Tx
fit <- lmerTest::lmer(
  `Discirmination Index` ~ Treat + (1|Litter),
  data = nor_analytic
)

fit_type3 <- car::Anova(fit, type = 3)
```

```{r process-models}
fit_sex_intrxn_tbl <- broom.mixed::tidy(fit_sex_intrxn) |>
  # restrcit to fixed effects
  dplyr::filter(
    effect == 'fixed'
  ) |>
  # select term, estimate std.error, and p.value
  dplyr::select(
    term, estimate, std.error, p.value
  ) |>
  # round estimates std.error, and p.value to 3 digits
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  # combine estimate and std.error into single column
  tidyr::unite(col = 'Estimate (SE)', estimate, std.error, sep = " (") |>
  # add missing `)` to `Estimate (SE)`
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ')'),
    p.value = dplyr::if_else(
      p.value == '0.000',
      true = "<0.001",
      false = p.value
    )
  )

fit_sex_tbl <- broom.mixed::tidy(fit_sex) |>
  # restrcit to fixed effects
  dplyr::filter(
    effect == 'fixed'
  ) |>
  # select term, estimate std.error, and p.value
  dplyr::select(
    term, estimate, std.error, p.value
  ) |>
  # round estimates std.error, and p.value to 3 digits
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  # combine estimate and std.error into single column
  tidyr::unite(col = 'Estimate (SE)', estimate, std.error, sep = " (") |>
  # add missing `)` to `Estimate (SE)`
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ')'),
    p.value = dplyr::if_else(
      p.value == '0.000',
      true = "<0.001",
      false = p.value
    )
  )

fit_tbl <- broom.mixed::tidy(fit) |>
  # restrcit to fixed effects
  dplyr::filter(
    effect == 'fixed'
  ) |>
  # select term, estimate std.error, and p.value
  dplyr::select(
    term, estimate, std.error, p.value
  ) |>
  # round estimates std.error, and p.value to 3 digits
  dplyr::mutate(
    dplyr::across(
      .cols = c(estimate, std.error, p.value),
      ~ format(round(.x, 3), nsmall = 3)
    )
  ) |>
  # combine estimate and std.error into single column
  tidyr::unite(col = 'Estimate (SE)', estimate, std.error, sep = " (") |>
  # add missing `)` to `Estimate (SE)`
  dplyr::mutate(
    `Estimate (SE)` = paste0(`Estimate (SE)`, ')'),
    p.value = dplyr::if_else(
      p.value == '0.000',
      true = "<0.001",
      false = p.value
    )
  )

# combine into single table
tmp <- dplyr::right_join(
  x = fit_tbl,
  y = fit_sex_tbl,
  by = 'term'
) |>
  dplyr::right_join(
    y = fit_sex_intrxn_tbl,
    by = 'term'
  )
```

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with a final sample size of `r nor |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()` Finally, the `discrimination index` variable was converted to numeric for analysis.

# Summary statistics

The discrimination index was summarized by Treatment group overall (Table \@ref(tab:summary-tbl)), and by Treatment groups stratified by sex (Table \@ref(tab:summary-tbl-sex)).


```{r summary-tbl}
discrim_tx_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of Discrimination Index by Treatment Groups."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 1.16
  )
```

\

```{r summary-tbl-sex}
discrim_tx_sex_tbl |>
  gtsummary::as_flex_table() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of Discrimination Index by Treatment Groups, stratified ",
      "by Sex."
    )
  ) |>
  flextable::width(
    width = 1.15
  ) |>
  flextable::width(
    j = 1,
    width = 0.7
  ) |>
  flextable::width(
    j = 2,
    width = 1.16
  )
```


```{r, include=FALSE}
nor_plot_data <- dplyr::bind_rows(
  nor_analytic, 
  nor_analytic |> dplyr::mutate(Treat = "Overall")
) |>
  dplyr::mutate(
    Treat = factor(
      Treat,
      levels = c(
        "Adjuvant+Saline", "CRMP1+CRMP2",
        "CRMP1+GDA", "LDHA+LDHB+CRMP1+STIP1",
        "STIP1+NSE", "Overall"
      )
    )
  )

cowplot::plot_grid(
  ggplot2::ggplot(
    data = nor_analytic,
    ggplot2::aes(x = `Discirmination Index`)
  ) +
  ggplot2::geom_histogram() +
  ggplot2::theme_bw(),
  ggplot2::ggplot(
    data = nor_plot_data,
    ggplot2::aes(x = `Discirmination Index`)
  ) +
  ggplot2::geom_histogram() +
  ggplot2::facet_wrap(
    .~Treat
  ) +
  ggplot2::theme_bw(),
  nrow = 2
)

```

\
\

# Model results

The relationship between treatment groups and discrimination index was assessed using linear mixed effects models with a random litter effect. First we fit a model with treatment groups, sex, and the interaction between them to assess for potential sex differences across the groups. The model results are shown in Table \@ref(tab:lme-tbl). \

There was not sufficient evidence to conclude that there are sex differences between the groups (p=`r round(fit_sex_intrxn_type3['Treat:Sex', 'Pr(>Chisq)'], 3)`), thus we can look at the main effects only models. Since the estimates of the effect of Treatment group on discrimination index are largely the same, we can focus on the model with only Treatment group as our final model. \

The discrimination index between the treatment groups was not statistically significant (Type 3 p-value = `r round(fit_type3['Treat', 'Pr(>Chisq)'], 3)`).

\

```{r lme-tbl}
tmp |>
  flextable::flextable() |>
  flextable::set_caption(
    caption = paste0(
      "Summary of model results assessing the relationship between Treatment ",
      "groups and discrimination index"
    )
  ) |>
  flextable::set_header_labels(
    values = list(
      term = 'Predictors',
      `Estimate (SE).x` = "Estimate (SE)",
      `p.value.x` = 'p-value',
      `Estimate (SE).y` = "Estimate (SE)",
      `p.value.y` = 'p-value',
      `Estimate (SE)` = "Estimate (SE)",
      `p.value` = 'p-value'
    )
  ) |>
  flextable::width(
    j = ,
    width = 1.1
  )
```
