---
title: "MAR 4.0 - USV"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2025-03-10"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, rstatix, ggpubr, mar4.0)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
fig_n = 0
tab_n = 0
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glmer(as.formula(formula_str), data = data, family = binomial(link = "logit")))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}

# Function to summarize and format model results
format_summary <- function(model_summary, coef_indices, column_name) {
  formatted <- apply(model_summary$coefficients[, coef_indices], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(summary_full, ...) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  
  for (i in 2:length(summaries)) {
    merged_data <- full_join(merged_data, summaries[[i]], by = "Effect")
  }
  
  # Convert numeric columns (p-values) to 4 decimal places (non-scientific)
  numeric_cols <- sapply(merged_data, is.numeric)
  merged_data[numeric_cols] <- lapply(merged_data[numeric_cols], function(x) {
    ifelse(is.na(x), "", as.character(format(round(x, 4), nsmall = 4, scientific = FALSE)))
  })
  
  return(merged_data)
}

# Function to perform ANOVA
perform_anova <- function(model, type = "III") {
  return(Anova(model, type = type)[, 3])
}


# Function to return the gtsummary-table by input dataset and outcome
summary_tbls <- function(
    dataset, outcome
) {

  tbl_male <- dataset |>
    dplyr::filter(Sex == "M") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          ) 
        ) |> gtsummary::add_overall(),,
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Males"
    )

  tbl_female <- dataset |>
    dplyr::filter(Sex == "F") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          )
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Females"
    )

  ans <- list(
    "males" = tbl_male,
    "females" = tbl_female
  )

  return(ans)
}
```

```{r}
# Ensure PND is a factor with reference level set to "4"
usv <- usv |>
  mutate(PND = factor(PND, levels = unique(PND)), PND = relevel(PND, ref = "4")) |>
  mutate(Slope = ifelse(Slope>=-1000,Slope,NA))

# Variables to analyze, including Count
vars <- names(usv)[11:15] 

# Create ANOVA results table
merged_summary_list <- list()
boxplot_list <- list()
post_hoc_h1_list <- list()
post_hoc_h2_list <- list()

# Main loop over variables
for (i in seq_along(vars)) {
  var <- vars[i]
  
  # Data preparation (ensure complete cases)
  usv_fct_compete <- usv |>
    dplyr::mutate(
      Treatment = factor(Treatment, levels = c(
        "Adjuvant+Saline", "CRMP1+CRMP2", "CRMP1+GDA",
        "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE"
      )),
      Sex = factor(Sex, levels = c("M", "F")),
      Litter = factor(`Litter ID`),
      Animal = factor(`Animal ID`)
    ) |>
    dplyr::filter(dplyr::if_all(c(Animal, Litter, Treatment, Sex, PND, !!sym(var)), ~ !is.na(.x)))

  # Perform 3-way ANOVA (Treatment * Sex * PND)
  mix_aov_3way <- rstatix::anova_test(
    dv = !!sym(var), 
    wid = Animal, 
    between = c(Treatment, Sex), 
    within = PND, 
    data = usv_fct_compete, 
    type = 3
  )

  # Perform 2-way ANOVA (Treatment * PND)
  mix_aov_2way <- rstatix::anova_test(
    dv = !!sym(var), 
    wid = Animal, 
    between = c(Treatment), 
    within = PND,
    covariate = Sex,
    data = usv_fct_compete, 
    type = 3
  )

    # Extract p-values from ANOVA tests and merge them into one table
  p_values_3way <- mix_aov_3way$ANOVA[,c("Effect","p")] |> dplyr::rename(`P-Value (3-Way Model; Sex as Effect Modifier)` = p)
  p_values_2way <- mix_aov_2way$ANOVA[,c("Effect","p")] |> dplyr::rename(`P-Value (2-Way Model; Sex as Confounder)` = p)
  combined_anova_results <- create_combined_estimate(p_values_3way, p_values_2way)
  
  # Store results in list format for further use
  merged_summary_list[[var]] <- combined_anova_results
  
  ######## Create Boxplots for numeric variable
  data_summary <- usv_fct_compete %>%
    group_by(Treatment, Sex, PND) %>%
    summarise(
      Mean = mean(.data[[var]], na.rm = TRUE),
      SE = sd(.data[[var]], na.rm = TRUE) / sqrt(n()),
      .groups = "drop"
    )

  # Create the plot with SE
  boxplot_list[[i]] <- ggboxplot(usv_fct_compete, x = "Treatment", y = var, color = "Sex", fill = "Sex", palette = "Set1") +
    geom_errorbar(data = data_summary, 
                  aes(x = Treatment, y = Mean, ymin = Mean - 2*SE, ymax = Mean + 2*SE, group = interaction(Treatment, Sex)),  
                  position = position_dodge(width = 0.8), 
                  width = 0.2, 
                  color = "black") +  
    geom_point(data = data_summary, 
               aes(x = Treatment, y = Mean, color = Sex, group = interaction(Treatment, Sex)), 
               position = position_dodge(width = 0.8), 
               size = 1, 
               shape = 21) +  
    facet_wrap(~PND, scales = "free_x", nrow = 1, labeller = label_both) +
    theme_minimal() +
    labs(title = paste0("Distribution of ", var, " by Treatment & Sex"),
         x = "Treatment",
         y = paste0(var, " Value")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "bottom",
          strip.text = element_text(size = 12, face = "bold"))
  
  ######## Running post-hoc test on continuous variable
  post_hoc_h1_list[[i]] <- run_posthoc_test_num_h1(data = usv_fct_compete, vars = var)
  post_hoc_h2_list[[i]] <- run_posthoc_test_num_h2(data = usv_fct_compete, vars = var)
}

```


# Background

In MAR 4.0 project, this section we will conduct analysis on the milestone data and report analysis results.

We are going to run the analysis using R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.


# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with total observations `r usv |> dplyr::pull(dplyr::any_of('Animal ID')) |> length()`, and total number of animals `r usv |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. Within the data, the "Slope" in one observation is set to be ignored because its extreme value (~= -2000).

The interested variables at USVs data are: `r paste(vars, collapse = ", ")`.

# Summary statistics

The all interested variables were summarized by Treatment group overall, stratified by Time and Sex.

```{r summary-tbl, results='asis'}
for (i in seq_along(vars)) {
  summary_list <- summary_tbls(usv, vars[i])
  
  purrr::walk2(summary_list, vars[i], function(tbl, var_name) {
    tbl |>
      gtsummary::as_flex_table() |>
      flextable::width(width = 1.15) |>
      flextable::width(j = 1, width = 1.16) |>
      flextable::flextable_to_rmd()
    
    tab_n <<- tab_n + 1  
    cat(paste0("\nTable ", tab_n, 
               ": Summary of ", 
               var_name, 
               " by Treatment Groups with Time Stratification (4, 8, and 12th week).\n\n"))
  })
}

```

# Model Results

To evaluate differences in mean USV measures between treatments over time, we conducted repeated measures ANOVA (or logistic regression for a cross-sectional categorical USV measure), incorporating animals as a random effect.

We fitted 2 models as follows in order: 

- Fit the 3-way interaction ANOVA model: Begin with a 3-way interaction model with the main effects of time, treatment (trt), and sex, their 2-way interactions and the three-way interaction of all three variables. This model assumes sex as an effect modifier. 

- If the 3-way interaction term in ANOVA is not statistically significant (i.e, if sex is not an effect modifier), we drop the 3-way interaction term and fit the 2-way interaction model that assumes sex as a confounding variable. 

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report’s delivery format.

## ANOVA Table

Type III ANOVA table displays the anova p-values for each variable in each model, and below is the explanation of each model:

- Three interaction model shows the p-value on three way interaction: y = time(PND) * trt * sex

- Two-way interaction model shows the p-value without the three way intereaction: y = time(PND) * trt + Sex + time(PND) : Sex

The models also account for within-subjects variation attributed to animals (i.e, inter-animal variability within the same treatment group), and all p-values are calculated through Type-III sum of squares of residuals, using the R “rstatix” package.

Additionally, we use boxplots for intuitive visualizations for cross-sectional pairwise group comparisons at each time point tested by t-test with a combination of Bonferroni adjustment method (more * means more significance of the pairwise t-test).


## Post-hoc Tests

Post-hoc tests are multiple pairwise comparisons between groups to identify which groups are different. Based on our hypothesis, we designed two sets of post-hoc tests for each variable. The first set is all comparisons between the control group and each treatment group, and the second set is pairwise comparisons across all treatment groups. The p-values are adjusted by Bonferroni method on each set of comparison and on each time period.


```{r anova-tb, results='asis'}
set_flextable_defaults(
  font.family = "Arial",  
  font.size = 10,      
  theme_fun = theme_vanilla 
)

# Loop and show all the table & figures
for (i in seq_along(vars)) {
  merged_summary_list[[i]] |>
    as.data.frame(row.names = NULL) |>
    flextable::flextable() |>
    flextable::set_table_properties(layout = "autofit") |>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, ": Cross-sectional Comparison of ANOVA Results for Two Model Specifications of ", vars[i] ,"\n\n"))
  
  # Print the boxplot
  print(boxplot_list[[i]])

  # Another page break before the next iteration
  fig_n = fig_n + 1
  cat(paste0("\nFigure ",fig_n, ": Boxplot showing the distribution of ", 
             vars[i] ," by Treatment and Sex across Postnatal Days (PND) with error bars depict ±2 standard errors (2SE)\n\n"))

  post_hoc_h1_list[[i]]|>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, 
             ": Post-hoc test for comparing the control null group (Adjuvant+Saline) to treatments with a control 95% family-wise confidence level - ", 
             vars[i] ,"\n"))

  post_hoc_h2_list[[i]]|>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, ": Post-hoc test for comparing across treatments with a control 95% family-wise confidence level - ", vars[i] ,"\n\n"))
  cat(paste0("\n- - - - - - - - - - - - - - - - - - - - - -\n"))
  cat(paste0("\n- - - - - - - - - - - - - - - - - - - - - -\n"))
}
```
