---
title: "MAR 4.0 - USV"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "2025-02-28"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, rstatix, ggpubr, mar4.0)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
fig_n = 0
tab_n = 0
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glmer(as.formula(formula_str), data = data, family = binomial(link = "logit")))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}

# Function to summarize and format model results
format_summary <- function(model_summary, coef_indices, column_name) {
  formatted <- apply(model_summary$coefficients[, coef_indices], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(summary_full, ...) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  
  for (i in 2:length(summaries)) {
    merged_data <- full_join(merged_data, summaries[[i]], by = "Effect")
  }
  
  # Convert numeric columns (p-values) to 4 decimal places (non-scientific)
  numeric_cols <- sapply(merged_data, is.numeric)
  merged_data[numeric_cols] <- lapply(merged_data[numeric_cols], function(x) {
    ifelse(is.na(x), "", as.character(format(round(x, 4), nsmall = 4, scientific = FALSE)))
  })
  
  return(merged_data)
}

# Function to perform ANOVA
perform_anova <- function(model, type = "III") {
  return(Anova(model, type = type)[, 3])
}


# Function to return the gtsummary-table by input dataset and outcome
summary_tbls <- function(
    dataset, outcome
) {

  tbl_male <- dataset |>
    dplyr::filter(Sex == "M") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          ) 
        ) |> gtsummary::add_overall(),,
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Males"
    )

  tbl_female <- dataset |>
    dplyr::filter(Sex == "F") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          )
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Females"
    )

  ans <- list(
    "males" = tbl_male,
    "females" = tbl_female
  )

  return(ans)
}
```

```{r}
# Ensure PND is a factor with reference level set to "4"
usv <- usv |>
  mutate(PND = factor(PND, levels = unique(PND)), PND = relevel(PND, ref = "4")) |>
  mutate(Slope = ifelse(Slope>=-1000,Slope,NA))

# Variables to analyze, including Count
vars <- names(usv)[11:15] 

# Create ANOVA results table
merged_summary_list <- list()
boxplot_list <- list()

# Main loop over variables
for (i in seq_along(vars)) {
  var <- vars[i]
  
  # Data preparation (ensure complete cases)
  usv_fct_compete <- usv |>
    dplyr::mutate(
      Treatment = factor(Treatment, levels = c(
        "Adjuvant+Saline", "CRMP1+CRMP2", "CRMP1+GDA",
        "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE"
      )),
      Sex = factor(Sex, levels = c("M", "F")),
      Litter = factor(`Litter ID`),
      Animal = factor(`Animal ID`)
    ) |>
    dplyr::filter(dplyr::if_all(c(Animal, Litter, Treatment, Sex, PND, !!sym(var)), ~ !is.na(.x)))

  # Perform 3-way ANOVA (Treatment * Sex * PND)
  mix_aov_3way <- rstatix::anova_test(
    dv = !!sym(var), 
    wid = Animal, 
    between = c(Treatment, Sex), 
    within = PND, 
    data = usv_fct_compete, 
    type = 3
  )

  # Perform 2-way ANOVA (Treatment * PND)
  mix_aov_2way <- rstatix::anova_test(
    dv = !!sym(var), 
    wid = Animal, 
    between = c(Treatment), 
    within = PND,
    covariate = Sex,
    data = usv_fct_compete, 
    type = 3
  )

  # Extract p-values from ANOVA tests
  p_values_3way <- mix_aov_3way$ANOVA[,c("Effect","p")] |> dplyr::rename(`P-Value (3-Way Model)` = p)  # Adjust indices based on output structure
  p_values_2way <- mix_aov_2way$ANOVA[,c("Effect","p")] |> dplyr::rename(`P-Value (2-Way Model)` = p) # Adjust indices based on output structure

  combined_anova_results <- create_combined_estimate(p_values_3way, p_values_2way)
  
  # Store results in list format for further use
  merged_summary_list[[var]] <- combined_anova_results
  
  # Perform T-test comparisons and merge the result into the boxplot (non-paired t-tests with Bonferroni correction)
  pwc <- usv_fct_compete %>%
    group_by(PND) %>%
    rstatix::pairwise_t_test(as.formula(paste0("`", var, "` ~ Treatment")), 
                             paired = FALSE, p.adjust.method = "bonferroni") %>%
    rstatix::add_xy_position(x = "PND")

  
  boxplot_list[[var]] <- ggpubr::ggboxplot(
    usv_fct_compete, x = "PND", y = var,
    color = "Treatment", palette = "Sex"
    ) +
    stat_pvalue_manual(pwc, tip.length = 0, hide.ns = TRUE) +
    labs(
      subtitle = get_test_label(mix_aov_2way, detailed = TRUE),
      caption = get_pwc_label(pwc)
    )
}

```


# Background

In MAR 4.0 project, this section we will wants to know analysis result on the USV data. 

We are going to run the analysis use R-package structure within github. R 4.4.1 is used for the analysis, and we are wrapping all the materials into a R package.

# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with a total observation `r usv |> dplyr::pull(dplyr::any_of('Animal ID')) |> length()`, and total number of animals `r usv |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. Within the data, the "Slope" in one observation is set to be ignored because its extreme value (~= -2000).

The interested variables at USVs data are: `r paste(vars, collapse = ", ")`.

# Summary statistics

The all interested variables were summarized by Treatment group overall, stratified by Time and Sex (Table \@ref(tab:summary-tbl)).

```{r summary-tbl, results='asis'}
for (i in seq_along(vars)) {
  summary_list <- summary_tbls(usv, vars[i])
  
  purrr::walk2(summary_list, vars[i], function(tbl, var_name) {
    tbl |>
      gtsummary::as_flex_table() |>
      flextable::width(width = 1.15) |>
      flextable::width(j = 1, width = 1.16) |>
      flextable::flextable_to_rmd()
    
    tab_n <<- tab_n + 1  
    cat(paste0("\nTable ", tab_n, 
               ": Summary of ", 
               var_name, 
               " by Treatment Groups with Time Stratification (4, 8, and 12th week).\n\n"))
  })
}

```

# Model Results

To evaluate differences in mean USV measures between treatments over time, we conducted repeated measures ANOVA (or logistic regression for a cross-sectional categorical USV measure), incorporating animals as a random effect.

We fitted 2 models as follows in order: 

- Fit the 3-way interaction ANOVA model: Begin with a 3-way interaction model with the main effects of time, treatment (trt), and sex, their 2-way interactions and the three-way interaction of all three variables. This model assumes sex as an effect modifier. 

- If the 3-way interaction term in ANOVA is not statistically significant (i.e, if sex is not an effect modifier), we drop the 3-way interaction term and fit the 2-way interaction model that assumes sex as a confounding variable. 

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report’s delivery format.

## ANOVA Table

Type III ANOVA table (Table \@ref(tab:anova-tb)) displays the anova p-values for each variable in each model, and below is the explanation of each model:

- Three interaction model shows the p-value on three way interaction: y = time(PND) * trt * sex

- Two-way interaction model shows the p-value without the three way intereaction: y = time(PND) * trt + Sex + time(PND) : Sex

The models also account for within-subjects variation attributed to animals (i.e, inter-animal variability within the same treatment group), and all p-values are calculated through Type-III sum of squares of residuals, using the R “rstatix” package.

Additionally, we use boxplots for intuitive visualizations for cross-sectional pairwise group comparisons at each time point tested by t-test with a combination of Bonferroni adjustment method (more * means more significance of the pairwise t-test).

```{r anova-tb, results='asis'}
set_flextable_defaults(
  font.family = "Arial",  # Optional: Set default font
  font.size = 10,         # Optional: Set default font size
  theme_fun = theme_vanilla # Optional: Set default theme
)

for (i in seq_along(vars)) {
  # Create flextable summary table
  merged_summary_list[[i]] |>
    as.data.frame(row.names = NULL) |>
    flextable::flextable() |>
    flextable::set_table_properties(layout = "autofit") |>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, ": Model Estimation Result - ", vars[i] ,"\n\n"))
  
  # Print the boxplot
  print(boxplot_list[[i]])

  # Another page break before the next iteration
  fig_n = fig_n + 1
  cat(paste0("\nFigure ",fig_n, ": Boxplot of Response Variable by Treatment, Time, and Sex\n\n"))

  # DescTools::DunnettTest(x = usv[!is.na(usv[vars[i]]),vars[i]] |> unlist(),g = usv$Treatment[!is.na(usv[vars[i]])]|> unlist())$`Adjuvant+Saline` |>
  #   as.data.frame() |>
  #   rownames_to_column("Comparison") |>
  #   mutate(
  #     Effect_Size = round(diff, 3),  # Round effect size to 3 decimals
  #     Confidence_Interval = sprintf("[%.3f, %.3f]", round(lwr.ci, 3), round(upr.ci, 3)),  # Round CI
  #     P_Value = formatC(round(pval, 3), format = "f", digits = 3)  # Round p-value and keep decimal format
  #   ) |>
  #   select(
  #     Group_Comparison = Comparison,  
  #     Effect_Size,                    
  #     Confidence_Interval,            
  #     P_Value                         
  #   ) |>
  #   flextable::flextable() |>
  #   flextable::set_table_properties(layout = "autofit") |>
  #   flextable::flextable_to_rmd()
  # tab_n = tab_n + 1
  # cat(paste0("\nTable ",tab_n, ": Dunnett's test for comparing several treatments with a control 95% family-wise confidence level - ", vars[i] ,"\n"))
  # 
  # TukeyHSD(aov(as.formula(paste0("`", vars[i], "` ~ Treatment")), data = usv|>filter(Treatment!="Adjuvant+Saline")))$Treatment |>
  #   as.data.frame() |>
  #   rownames_to_column("Comparison") |>
  #   mutate(
  #     Effect_Size = round(diff, 3),  
  #     Confidence_Interval = sprintf("[%.3f, %.3f]", round(lwr, 3), round(upr, 3)),  # Round CI
  #     P_Value = formatC(round(`p adj`, 3), format = "f", digits = 3)  # Round p-value and keep decimal format
  #   ) |>
  #   select(
  #     Group_Comparison = Comparison, 
  #     Effect_Size,                  
  #     Confidence_Interval,           
  #     P_Value                         
  #   ) |>
  #   flextable::flextable() |>
  #   flextable::set_table_properties(layout = "autofit") |>
  #   flextable::flextable_to_rmd()
  # tab_n = tab_n + 1
  # cat(paste0("\nTable ",tab_n, ": Tukey multiple comparisons of means 95% family-wise confidence level - ", vars[i] ,"\n\n"))
  cat(paste0("\n- - - - - - - - - - - - - - - - - - - - - -\n"))
}
```
