---
title: "MAR 4.0 - USV"
output: 
  bookdown::word_document2:
    number_section: false
    reference_docx: docx_landscape.docx
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pacman::p_load(dplyr, tidyr, tibble, lme4, lmerTest, car, sjPlot, broom.mixed,
               flextable, rstatix, ggpubr, mar4.0)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
set_flextable_defaults(
  font.family = "Arial",
  font.size = 8,
  theme_fun = theme_vanilla
)
fig_n = 0
tab_n = 0
```

```{r}
###################### Functions to generate the table and models

# Loop through each variable
fit_model <- function(var, formula_str, data, model_type = "lmer") {
  if (model_type == "glmer") {
    return(glmer(as.formula(formula_str), data = data, family = binomial(link = "logit")))
  } else {
    return(lmerTest::lmer(as.formula(formula_str), data = data))
  }
}

# Function to summarize and format model results
format_summary <- function(model_summary, coef_indices, column_name) {
  formatted <- apply(model_summary$coefficients[, coef_indices], 1, function(row) {
    estimate <- round(row[1], 3)
    se <- round(row[2], 3)
    p_value <- ifelse(round(row[3], 3) < 0.001, "<0.001", paste0("=", round(row[3], 3)))
    paste0(estimate, " ±", se, " (p", p_value, ")")
  })
  formatted_summary <- as.data.frame(formatted)
  colnames(formatted_summary) <- c(column_name)
  return(formatted_summary)
}

# Function to combine different model estimates into one table and ordered by the summary_fullmodel (for the same variable).
create_combined_estimate <- function(summary_full, ...) {
  summaries <- list(summary_full, ...)
  merged_data <- summaries[[1]]
  
  for (i in 2:length(summaries)) {
    merged_data <- full_join(merged_data, summaries[[i]], by = "Effect")
  }
  
  # Convert numeric columns (p-values) to 4 decimal places (non-scientific)
  numeric_cols <- sapply(merged_data, is.numeric)
  merged_data[numeric_cols] <- lapply(merged_data[numeric_cols], function(x) {
    ifelse(is.na(x), "", as.character(format(round(x, 4), nsmall = 4, scientific = FALSE)))
  })
  
  return(merged_data)
}

# Function to perform ANOVA
perform_anova <- function(model, type = "III") {
  return(Anova(model, type = type)[, 3])
}


# Function to return the gtsummary-table by input dataset and outcome
summary_tbls <- function(
    dataset, outcome
) {

  tbl_male <- dataset |>
    dplyr::filter(Sex == "M") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          ) 
        ) |> gtsummary::add_overall(),,
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Males"
    )

  tbl_female <- dataset |>
    dplyr::filter(Sex == "F") |>
    dplyr::select(Treatment, PND, dplyr::all_of(outcome)) |>
    gtsummary::tbl_strata(
      strata = PND,
      .tbl_fun = ~ .x |>
        gtsummary::tbl_summary(
          by = Treatment,
          type = where(is.numeric) ~ "continuous2",
          statistic = gtsummary::all_continuous2() ~ c(
            "{mean} ({sd})", "{median} ({p25}, {p75})", "{min} - {max}"
          )
        ) |> gtsummary::add_overall(),
      .combine_with = "tbl_stack"
    ) |>
    gtsummary::modify_spanning_header(
      gtsummary::all_stat_cols() ~ "Females"
    )

  ans <- list(
    "males" = tbl_male,
    "females" = tbl_female
  )

  return(ans)
}
```

```{r}
data <- usv |>
  mutate(PND = factor(PND, levels = unique(PND)), PND = relevel(PND, ref = "4")) |>
  mutate(Slope = ifelse(Slope>=-1000,Slope,NA))

# Variables to analyze, including Count
vars <- names(data)[11:15] 

# Create ANOVA results table
merged_summary_list <- list()
plot_list <- list()
post_hoc_h1_list <- list()
post_hoc_h2_list <- list()

for (i in 1:length(vars)) {
  var <- vars[i]
  # Data preparation (ensure complete cases)
  data_fct_compete <- data |>
    dplyr::mutate(
      Treatment = factor(Treatment, levels = c(
        "Adjuvant+Saline", "CRMP1+CRMP2", "CRMP1+GDA",
        "LDHA+LDHB+CRMP1+STIP1", "STIP1+NSE"
      )),
      Sex = factor(Sex, levels = c("M", "F")),
      Litter = factor(`Litter ID`),
      Animal = factor(`Animal ID`),
      PND = factor(PND, levels = c(4, 8, 12), ordered = TRUE)
    ) |>
    dplyr::filter(dplyr::if_all(c(Animal, Litter, Treatment, Sex, PND, !!sym(var)), ~ !is.na(.x)))
  
  # If working on the variable with 30s max recording, trimming those equal to 30s
  # Also check low counting PND and exclude those
  if (var %in% c("Circle Traverse","Negative Geotaxis","Cliff Avoidance")){
    data_fct_compete <- data_fct_compete |>
      filter(!is.na(!!sym(var)) & !!sym(var) < 30)
    
    exclude_pnd <- data_fct_compete |>
      group_by(PND, Treatment, Sex) |>
      summarise(n = n(), .groups = "drop") |>
      filter(n <= 1) |>
      pull(PND) |>
      unique()

    data_fct_compete <- data_fct_compete |>
      filter(!PND %in% exclude_pnd)
  }

  
  # Perform 3-way ANOVA (Treatment * Sex * PND)
  mix_aov_3way <- afex::aov_car(
  formula = as.formula(paste0("`", var, "` ~ Treatment * Sex * PND + Error(Animal/PND)")),
  data = data_fct_compete,
  type = 3
  )
  
  # Perform 2-way ANOVA (Treatment * PND)
  mix_aov_2way <- afex::aov_car(
  formula = as.formula(paste0("`", var, "` ~ Treatment * PND + Sex * PND + Error(Animal/PND)")),
  data = data_fct_compete,
  type = 3
  )

  # Extract p-values from ANOVA tests and merge them into one table
  if (!is.null(mix_aov_3way$ANOVA)) {
    p_values_3way <- mix_aov_3way$ANOVA[, c("Effect", "p.value")] |>
      dplyr::rename(`P-Value (3-Way Model)` = p.value)
    
    p_values_2way <- mix_aov_2way$ANOVA[, c("Effect", "p.value")] |>
      dplyr::rename(`P-Value (2-Way Model)` = p.value)
  } else {
    p_values_3way <- mix_aov_3way$anova_table |>
      as.data.frame() |>
      tibble::rownames_to_column(var = "Effect") |>
      dplyr::select(`Effect`, `Pr(>F)`) |>
      dplyr::rename(`P-Value (3-Way Model)` = `Pr(>F)`)
    
    p_values_2way <- mix_aov_2way$anova_table |>
      as.data.frame() |>
      tibble::rownames_to_column(var = "Effect") |>
      dplyr::select(`Effect`, `Pr(>F)`) |>
      dplyr::rename(`P-Value (2-Way Model)` = `Pr(>F)`)
  }
  combined_anova_results <- create_combined_estimate(p_values_3way, p_values_2way)
  merged_summary_list[[i]] <- combined_anova_results
  
  ######## Running post-hoc test on continuous variable
  p_list_3 <- emmeans::contrast(emmeans::emmeans(mix_aov_3way, ~ Treatment|PND), 
                                method = "pairwise", 
                                adjust = "none") |>
    as.data.frame() |> 
    mutate(PND = gsub("\\D", "", PND))
  p_list_2 <- emmeans::contrast(emmeans::emmeans(mix_aov_2way, ~ Treatment|PND), 
                                method = "pairwise", 
                                adjust = "none") |>
    as.data.frame() |> 
    mutate(PND = gsub("\\D", "", PND))
  
  p_list_sex <- emmeans::contrast(emmeans::emmeans(mix_aov_3way, ~ Sex|PND * Treatment), 
                                method = "pairwise", 
                                adjust = "none") |>
    as.data.frame() |> 
    mutate(PND = gsub("\\D", "", PND))
    
  post_hoc_h1_list[[i]] <- run_posthoc_test_num_h1(
    data = data_fct_compete, 
    var = var,
    p_list_3 = p_list_3[,c("contrast", "p.value", "PND")],
    p_list_2 = p_list_2[,c("contrast", "p.value", "PND")]) |> 
    make_posthoc_flextable(metric = "Mean_Diff")

  post_hoc_h2_list[[i]] <- run_posthoc_test_num_h2(
    data = data_fct_compete, 
    var = var,
    p_list_3[,c("contrast", "p.value", "PND")],
    p_list_2[,c("contrast", "p.value", "PND")]) |> 
    make_posthoc_flextable(metric = "Mean_Diff")
  
  plot_list[[i]] <- create_meanplot(
    dataset = data_fct_compete,
    variable = var, 
    contrast_p = p_list_3 |> adjust_p_by_pnd(),
    contrast_p_sex = p_list_sex |> adjust_p_by_pnd())
}
```


# Background

In MAR 4.0 project, this section we will conduct analysis on the USV data and report analysis results.

We are going to run the analysis using R-package structure within github. R 4.4.3 is used for the analysis, and we are wrapping all the materials into a R package.


# Data Processing

The `familiarity` trials were excluded. Any animals that did not explore at least one area/chamber or animals that were cross-fostered were excluded from analysis, leaving us with total observations `r usv |> dplyr::pull(dplyr::any_of('Animal ID')) |> length()`, and total number of animals `r usv |> dplyr::pull(dplyr::any_of('Animal ID')) |> unique() |> length()`. Within the data, the "Slope" in one observation is set to be ignored because its extreme value (~= -2000).

The interested variables at USVs data are: `r paste(vars, collapse = ", ")`.

# Summary statistics

The all interested variables were summarized by Treatment group overall, stratified by Time and Sex.

```{r summary-tbl, results='asis'}
for (i in seq_along(vars)) {
  cat("\n\n### ---", vars[i], "---\n\n")
  summary_list <- summary_tbls(usv, vars[i])
  
  purrr::walk2(summary_list, vars[i], function(tbl, var_name) {
    tbl |>
      gtsummary::as_flex_table() |>
      flextable::width(width = 1.15) |>
      flextable::width(j = 1, width = 1.16) |>
      flextable::flextable_to_rmd()
    
    tab_n <<- tab_n + 1  
    cat(paste0("\nTable ", tab_n, 
               ": Summary of ", 
               var_name, 
               " by Treatment Groups with Time Stratification (4, 8, and 12th week).\n\n"))
  })
}

```

# Model Results


To evaluate differences in mean milestone measures between treatments over time, we conducted repeated measures ANOVA (or logistic regression for a cross-sectional categorical milestone measure), incorporating animals as a random effect. We fitted model as follow:

-   Fit the 3-way interaction ANOVA model: Begin with a 3-way interaction model with the main effects of time, treatment (trt), and sex, their 2-way interactions and the three-way interaction of all three variables. This model assumes sex as an effect modifier.

At this stage, all models ANOVA and estimated results are fitted for internal review. However, adjustments may be made based on decisions about the report’s delivery format.

## Post Hoc Analysis and P-Value Reporting

Post-hoc tests are multiple pairwise comparisons between groups to identify which groups are different. Based on our hypothesis, we designed two set of post-hoc tests for each of variable. The first set is all comparison between the control group and each treatment group, and the second set is across all treatment group. The third set of comparison is between male and female within each PND and Treatments, and this is shown only in the figure if available.

## ANOVA Table

Type III ANOVA table displays the anova p-values for each variable in each model, and below is the explanation of the model:

- Three interaction model shows the p-value on three way interaction: y = time(PND) * trt * sex

The models also account for within-subjects variation attributed to animals (i.e, inter-animal variability within the same treatment group), and all p-values are calculated through Type-III sum of squares of residuals, using the R “rstatix” package.

## Post-hoc Tests

Post-hoc tests are multiple pairwise comparisons between groups to identify which groups are different. Based on our hypothesis, we designed two sets of post-hoc tests for each variable. The first set is all comparisons between the control group and each treatment group, and the second set is pairwise comparisons across all treatment groups. The p-values are adjusted by Bonferroni method on each set of comparison and on each time period.

## Post Hoc Analysis and P-Value Reporting

To evaluate pairwise differences between treatment groups across postnatal day (PND) levels, we performed post hoc tests tailored to the data type. The resulting tables report p-values from both the interaction model (inter) and the additive model (add). For each model, both raw and adjusted p-values were provided, with adjustment performed using the Hochberg method. The p-value adjustment was applied separately within each time period (PND = 4, 8, 12) and within each hypothesis type (primary or secondary).

For categorical outcomes, the proportion difference between groups was computed along with its standard error using the pooled standard error formula for two proportions. A 95% confidence interval for the difference was calculated using the normal approximation.

A custom helper function was used to format and annotate p-values: values less than 0.05 were marked with an asterisk (\*), and all values were rounded to three decimal places to enhance readability.

**Note:** Post hoc analysis p-values are derived from the `emmeans` package for *estimated marginal means* and adjusted by Hochberg method on each set of comparison and on each time period. These p-values reflect pairwise comparisons between adjusted group means, accounting for other covariates or factors included in the model. A non-significant main effect does not rule out significant pairwise comparisons between individual groups, and vice versa. Post hoc p-values should be interpreted accordingly, in the context of the specific contrasts being tested. The direct mean and standard error are also provided.

## Modeling Result

### EXAMPLE - How to interpret the Posthoc table

The post hoc table presents pairwise group comparisons for the outcome variable (here, "Belly Temp") across different postnatal days (PND = 4, 8, 12). Each row corresponds to a comparison between a treatment group and the control group (Adjuvant+Saline).

### EXAMPLE - How to interpret errorbar plot

- Each colored dot represents the mean belly temperature for a group (blue = males, yellow = females).

- The vertical lines ("error bars") show the 95% confidence interval (CI) for each group mean, indicating the precision of the estimate.

- The x-axis shows different treatment groups. The plot is divided into three panels, one for each time point (PND).

- Horizontal black brackets with asterisks indicate significant differences between the connected groups AFTER p.value adjustment:

    - * means p < 0.05 (statistically significant)
    
    - ** means p < 0.01 (very significant)
    
    - *** means p < 0.001 (highly significant)

**How to interpret:**

- **Compare Means:** Look at the position of each colored dot. Higher dots mean higher average belly temperatures. For example, at PND 8, both males and females have higher temperatures compared to PND 4.

- **Confidence Intervals:** If error bars for two groups do not overlap much, this suggests a real difference in their means. However, always rely on the asterisks for statistical significance.

- **Significance Brackets:** A black bracket with an asterisk means the difference between those two groups is statistically significant after correcting for multiple comparisons.


```{r anova-tb, results='asis'}
set_flextable_defaults(
  font.family = "Arial",  
  font.size = 10,      
  theme_fun = theme_vanilla 
)

# Loop and show all the table & figures
for (i in seq_along(vars)) {
  cat("\n\n### ---", vars[i], "---\n\n")
  merged_summary_list[[i]] |>
    as.data.frame(row.names = NULL) |>
    flextable::flextable() |>
    flextable::set_table_properties(layout = "autofit") |>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, ": Cross-sectional Comparison of ANOVA Results for Two Model Specifications of ", vars[i] ,"\n\n"))
  
  # Print the boxplot
  print(plot_list[[i]])

  # Another page break before the next iteration
  fig_n = fig_n + 1
  cat(paste0(
    "\nFigure ", fig_n, ": ",
    "Errorbar plot showing the estimated proportion and 95% confidence intervals for ", vars[i],
    " by Treatment group and Sex across Postnatal Days (PND). ",
    "P-values for between-treatment and between-sex comparisons are reported separately and adjusted using the Hochberg method.\n\n"
  ))
  post_hoc_h1_list[[i]]|>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, 
             ": Post-hoc test for comparing the control null group (Adjuvant+Saline) to treatments with a control 95% family-wise confidence level - ", 
             vars[i] ,"\n"))

  post_hoc_h2_list[[i]]|>
    flextable::flextable_to_rmd()
  tab_n = tab_n + 1
  cat(paste0("\nTable ",tab_n, ": Post-hoc test for comparing across treatments with a control 95% family-wise confidence level - ", vars[i] ,"\n\n"))
  cat(paste0("\n- - - - - - - - - - - - - - - - - - - - - -\n"))
  cat(paste0("\n- - - - - - - - - - - - - - - - - - - - - -\n"))
}
```
